---
title: "Assess Simulated PTSD Symptoms"
author:
- "<h5 style=\"font-style:italic\"> Laura Weidmann"
- "<h5 style=\"font-style:italic\"> Tobias R. Spiller"
date: "<h5 style=\"font-style:roman\"> `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
  pdf_document:
    toc: true
    toc_depth: '5'
subtitle: Version 0.0.2
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)
```

```{r Load Libraries, message=FALSE, warning=TRUE, include=FALSE}
# Data handling
library(tidyverse)
library(readr)

# Demographics
library(table1)
library(gtsummary)

# Analysis
#library("devtools")
#devtools::install_github("orduek/PsychPower")
library(PsychPower)
```

## 1. Import and prepare data

```{r PCL import data, message=FALSE, warning=TRUE, include=FALSE}
# Import data
simulated_ptsd_data <- read_csv("Data/simulated_skewed_ptsd_data.csv")
```

## 2. Sample descriptive

#### 2.1 Total score & Diagnosis

```{r PCL Table 1, echo=FALSE, message=FALSE}
# Create PCL total score
simulated_ptsd_total <- simulated_ptsd_data %>% 
  mutate(total = rowSums(select(., symptom_1:symptom_20)))

# # Create Diagnosis
simulated_ptsd_total <- simulated_ptsd_total %>%
  mutate(
    PTSD_Diagnosis = ifelse(
      (symptom_1 > 2 | symptom_2 > 2 | symptom_3 > 2 | symptom_4 > 2 | symptom_5 > 2) &
      (symptom_6 > 2 | symptom_7 > 2) &
      (symptom_8 > 2 | symptom_9 > 2 | symptom_10 > 2 | symptom_11 > 2 | symptom_12 > 2 | symptom_13 > 2 | symptom_14 > 2) &
      (symptom_15 > 2 | symptom_16 > 2 | symptom_17 > 2 | symptom_18 > 2 | symptom_19 > 2 | symptom_20 > 2),
      TRUE, FALSE
    )
  )

simulated_ptsd_total %>%
  summarise(
    mean_total = mean(total),
    sd_total = sd(total),
    n_diagnosed = sum(PTSD_Diagnosis)
  )
```

#### 2.2 Cronbach's alpha

Selected sample

```{r PCL cronbach, echo=FALSE, message=FALSE}
cronbach <- psych::alpha(subset(simulated_ptsd_total, select = (-total)))
cronbach$total
```

Summary of items & histogram

```{r PCL summary, echo=FALSE, message=FALSE}
summary(simulated_ptsd_total)
hist(simulated_ptsd_total$total)
```

## 3. Binarizing & create datax

Symptoms rated 2 or 3 will be considered as present. [not shown]

```{r PCL Binarizing, echo=FALSE, message=FALSE, warning=FALSE}
## set cut-off
cut_off <- 1 #will be used with <= // 

## Binarize
data1_binarized <- simulated_ptsd_total
n_cols <- ncol(data1_binarized)
data1_binarized <- data1_binarized %>%
  rename_with(~ paste0("Q", seq_len(n_cols - 2)), 1:(n_cols -2 ))

for (i in 1:(ncol(data1_binarized)-2)){
  orig <- paste("q", i, sep = "")
  bin <- paste("Q", i, sep = "")
  data1_binarized[orig] <- dplyr::case_when(data1_binarized[bin]<= cut_off ~ 0, data1_binarized[bin]>cut_off ~ 1)  #0 = "Symptom absent", 1 = "Symptom present"

}

# Create new data frame
data2 <- data1_binarized %>% 
  select(PTSD_Diagnosis:ncol(data1_binarized)) %>% 
  select(-PTSD_Diagnosis) %>% 
  tibble()


## Count frequency of profiles
data2_counted <- plyr::count(data2[, ]) # 510

# Create sum score of endorsed symptoms
data2_counted <- data2_counted %>% 
  mutate(total_bin = rowSums(data2_counted)-freq)


## Create datax 
# Create full dataset
datax <- dplyr::left_join(data1_binarized, data2_counted)


## Create dataframes for PsychPower Package 
data_PTSD_PsychPower <- read_csv("Data/simulated_skewed_ptsd_data.csv") %>%  select (symptom_1:symptom_20)
data_binarized_PsychPower <- PsychPower::binarize(data_PTSD_PsychPower, cut_off = 1)

data_frequency_PsychPower <- PsychPower::pheno_frequency(
  data_binarized_PsychPower, target_columns = tidyselect::starts_with("v_bin"))
```

## 4. Description of combinations

Most common combination, its frequency & median frequency

```{r PCL Describe Number, echo=FALSE, message=FALSE}
desc_pheno_PsychPower <- PsychPower::describe_pheno(data_frequency_PsychPower, frequency = "freq") 
desc_pheno_PsychPower
```

10 Most common Phenotype

```{r PCL Describe 10 most, echo=FALSE, message=FALSE}
PsychPower::common_pheno(data_frequency_PsychPower, frequency = "freq", n_phenotypes = 10)
```

Profiles (N, %) reported by less or equal than in 1% of the individuals

```{r PCL Having less then 1 percent, echo=FALSE, message=FALSE}
one_percent_sample = round(nrow(data_binarized_PsychPower)*0.01, 0)
data2_counted %>% filter(freq<=one_percent_sample) %>% nrow()

round((data2_counted %>% filter(freq<=one_percent_sample) %>% nrow()) / nrow(data2_counted), 3)
```

Individuals reporting one of the ten most common combinations

```{r PCL Having 10 most, echo=FALSE, message=FALSE}
data2_counted <- data2_counted %>% 
  arrange(desc(freq))
data2_counted[1:10,] %>% summarise(sum(freq))
```

Individuals reporting NOT one of the ten most common combinations

```{r PCL  Not Having 10 most, echo=FALSE, message=FALSE}
data2_counted[11:nrow(data2_counted),] %>% summarise(sum(freq))
```

Individuals reporting 1% most common combinations

```{r PCL Describe 1 percent, echo=FALSE, message=FALSE}
one_percent =  round(nrow(data2_counted)*0.01, 0)

data2_counted[1:one_percent,] %>% summarise(sum(freq))
```

Individuals reporting one of the 50% least common combinations

```{r PCL Describe 50 percent, echo=FALSE, message=FALSE}
fifty_percent =  round(nrow(data2_counted)*0.5, 0)

data2_counted[fifty_percent:nrow(data2_counted),] %>% summarise(sum(freq))

```

Plot 100 most common profiles

```{r Plot 100 most common, echo=FALSE, message=FALSE}
fig1 <- PsychPower::plot_pheno(data_frequency_PsychPower, frequency = "freq",
                               n_phenotypes = 100, color = "grey26")
fig1
```

## 5. Session Info

```{r Session Info, echo=FALSE, message=FALSE}
sessionInfo()
```
