---
title: "Assess Simulated PTSD Symptoms"
author:
- "<h5 style=\"font-style:italic\"> Laura Weidmann"
- "<h5 style=\"font-style:italic\"> Tobias R. Spiller"
date: "<h5 style=\"font-style:roman\"> `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
  pdf_document:
    toc: true
    toc_depth: '5'
subtitle: Version 0.0.2
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)
```

```{r Load Libraries, message=FALSE, warning=TRUE, include=FALSE}
# Data handling
library(tidyverse)
library(readr)
library(data.table)

# Demographics
library(table1)
library(gtsummary)

# Analysis
#library("devtools")
#devtools::install_github("orduek/PsychPower")
library(PsychPower)
```

## 1. Import and prepare data

```{r PCL import data, message=FALSE, warning=TRUE, include=FALSE}
# Import data
simulated_ptsd_data <- read_csv("Data/simulated_skewed_ptsd_data.csv")
```

## 2. Sample descriptive

#### 2.1 Total score & Diagnosis

```{r PCL Table 1, echo=FALSE, message=FALSE}
# Create PCL total score
simulated_ptsd_total <- simulated_ptsd_data %>% 
  mutate(total = rowSums(select(., symptom_1:symptom_20)))

# # Create Diagnosis
#simulated_ptsd_total <- simulated_ptsd_total %>%
  #mutate(
    #PTSD_Diagnosis = ifelse(
      #(symptom_1 > 2 | symptom_2 > 2 | symptom_3 > 2 | symptom_4 > 2 | symptom_5 > 2) &
      #(symptom_6 > 2 | symptom_7 > 2) &
      #(symptom_8 > 2 | symptom_9 > 2 | symptom_10 > 2 | symptom_11 > 2 | symptom_12 > 2 | symptom_13 > 2 | symptom_14 #> 2) &
      #(symptom_15 > 2 | symptom_16 > 2 | symptom_17 > 2 | symptom_18 > 2 | symptom_19 > 2 | symptom_20 > 2),
      #TRUE, FALSE
    #)
  #)

# Create function that implements PTSD diagnostic criteria
  create_ptsd_diagnosis <- function(data) {
    
    # Create new column initially set to FALSE
    data$PTSD_Diagnosis <- FALSE
    
    # Check total score criterion
    total_criterion <- data$total >= 33
    
    # Check symptom criteria
    criterion_B <- rowSums(data[, paste0("symptom_", 1:5)] >= 2) >= 1  # At least 1 from symptoms 1-5
    criterion_C <- rowSums(data[, paste0("symptom_", 6:7)] >= 2) >= 1  # At least 1 from symptoms 6-7
    criterion_D <- rowSums(data[, paste0("symptom_", 8:14)] >= 2) >= 2 # At least 2 from symptoms 8-14
    criterion_E <- rowSums(data[, paste0("symptom_", 15:20)] >= 2) >= 2 # At least 2 from symptoms 15-20
    
    # All symptom criteria met
    all_symptom_criteria <- criterion_B & criterion_C & criterion_D & criterion_E
    
    # Combined criteria: Either total >= 33 OR all symptom criteria met OR both conditions together
    data$PTSD_Diagnosis <- total_criterion | 
      all_symptom_criteria |
      (total_criterion & all_symptom_criteria)
    
    return(data)
  }

#Apply the function to our dataframe
  simulated_ptsd_total <- create_ptsd_diagnosis (simulated_ptsd_total)

#Summarize
simulated_ptsd_total %>%
  summarise(
    mean_total = mean(total),
    sd_total = sd(total),
    n_diagnosed = sum(PTSD_Diagnosis)
  )
```

#### 2.2 Cronbach's alpha

Selected sample

```{r PCL cronbach, echo=FALSE, message=FALSE}
cronbach <- psych::alpha(subset(simulated_ptsd_total, select = (-total)))
cronbach$total
```

### 2.3 Summary of items & histogram

```{r PCL summary, echo=FALSE, message=FALSE}
summary(simulated_ptsd_total)
hist(simulated_ptsd_total$total)
```

## 3. Binarizing & create datax

Symptoms rated 2 or 3 will be considered as present. [not shown]

```{r PCL Binarizing, echo=FALSE, message=FALSE, warning=FALSE}
#selecting only the "TRUE" Diagnosis
simulated_ptsd_total_selected <- simulated_ptsd_total %>%
  filter(PTSD_Diagnosis=="TRUE")

## set cut-off
cut_off <- 1 #will be used with <= // 

##rename variables to "Q1-QN" for binarization
n_cols <- ncol(simulated_ptsd_total)
simulated_ptsd_selected_rename <- simulated_ptsd_total_selected %>%
  rename_with(~ paste0("Q", seq_len(n_cols - 2)), 1:(n_cols -2 ))

## Binarize
data_binarized <- simulated_ptsd_selected_rename

for (i in 1:(ncol(data_binarized)-2)){
  orig <- paste("q", i, sep = "")
  bin <- paste("Q", i, sep = "")
  data_binarized[orig] <- dplyr::case_when(data_binarized[bin]<= cut_off ~ 0, data_binarized[bin]>cut_off ~ 1)  #0 = "Symptom absent", 1 = "Symptom present"
}

# Create new data frame
data2 <- data_binarized %>% 
  select(PTSD_Diagnosis:ncol(data_binarized)) %>% 
  select(-PTSD_Diagnosis) %>% 
  tibble()

## Count frequency of profiles
data2_counted <- plyr::count(data2[, ]) 

# Create sum score of endorsed symptoms
data2_counted <- data2_counted %>% 
  mutate(total_bin = rowSums(data2_counted)-freq)

## Create datax 
# Create full dataset
datax <- dplyr::left_join(data_binarized, data2_counted)


## Create dataframes for PsychPower Package 
data_PTSD_PsychPower <- simulated_ptsd_total_selected %>%  
  	select (symptom_1:symptom_20)
#binarize
data_binarized_PsychPower <- PsychPower::binarize(data_PTSD_PsychPower, cut_off = 1)
#frequency
data_frequency_PsychPower <- PsychPower::pheno_frequency(
  data_binarized_PsychPower, target_columns = tidyselect::starts_with("v_bin"))
```

## 4. Description of combinations

Most common combination, its frequency & median frequency

```{r PCL Describe Number, echo=FALSE, message=FALSE}
desc_pheno_PsychPower <- PsychPower::describe_pheno(data_frequency_PsychPower, frequency = "freq") 
desc_pheno_PsychPower
```

10 Most common Phenotype

```{r PCL Describe 10 most, echo=FALSE, message=FALSE}
PsychPower::common_pheno(data_frequency_PsychPower, frequency = "freq", n_phenotypes = 10)
```

Profiles (N, %) reported by less or equal than in 1% of the individuals

```{r PCL Having less then 1 percent, echo=FALSE, message=FALSE}
one_percent_sample = round(nrow(data_binarized_PsychPower)*0.01, 0)
data2_counted %>% filter(freq<=one_percent_sample) %>% nrow()

round((data2_counted %>% filter(freq<=one_percent_sample) %>% nrow()) / nrow(data2_counted), 3)
```

Individuals reporting one of the ten most common combinations

```{r PCL Having 10 most, echo=FALSE, message=FALSE}
data2_counted <- data2_counted %>% 
  arrange(desc(freq))
data2_counted[1:10,] %>% summarise(sum(freq))
```

Individuals reporting NOT one of the ten most common combinations

```{r PCL  Not Having 10 most, echo=FALSE, message=FALSE}
data2_counted[11:nrow(data2_counted),] %>% summarise(sum(freq))
```

Individuals reporting 1% most common combinations

```{r PCL Describe 1 percent, echo=FALSE, message=FALSE}
one_percent =  round(nrow(data2_counted)*0.01, 0)

data2_counted[1:one_percent,] %>% summarise(sum(freq))
```

Individuals reporting one of the 50% least common combinations

```{r PCL Describe 50 percent, echo=FALSE, message=FALSE}
fifty_percent =  round(nrow(data2_counted)*0.5, 0)

data2_counted[fifty_percent:nrow(data2_counted),] %>% summarise(sum(freq))

```

Plot 100 most common profiles

```{r Plot 100 most common, echo=FALSE, message=FALSE}
fig1 <- PsychPower::plot_pheno(data_frequency_PsychPower, frequency = "freq",
                               n_phenotypes = 100, color = "grey26")
fig1
```

## 5. Theoretically possible combinations of binarized symptoms fulfilling diagnosis (without consideration of the total score criterion) and their frequency in real-world data
```{r Theoretically possible combinations of binarized symptoms fulfilling diagnosis (without consideration of the total score criterion) and their frequency in real-world data, echo=FALSE, message=FALSE}
 ## Create function
create_possible_combinedwith_real_binarized_without_total <- function(data) {
  # First rename the columns in the input data (real-world) to match the expected format
  renamed_data <- data %>%
    rename_with(~ paste0("symptom_", 1:20))
  
  # Generate new dataframe (all_combinations_binarized_without_total) with all possible combinations of symptom combination (binarized) without respecting the criteria of totalscore
  symptom_cols_binarized_without_total <- paste0("symptom_", 1:20) # Total number of symptoms
  symptom_values_binarized_without_total <- 0:1  # 0: absent, 1: present
  all_combinations_binarized_without_total <- expand.grid(lapply(1:20, function(i) symptom_values_binarized_without_total))
  colnames(all_combinations_binarized_without_total) <- symptom_cols_binarized_without_total
  
  # Define criteria for filtering
  criteria_1 <- (
    rowSums(all_combinations_binarized_without_total[, 1:5]) >= 1 &          # At least 1 symptom from 1-5
      rowSums(all_combinations_binarized_without_total[, 6:7]) >= 1 &          # At least 1 symptom from 6-7
      rowSums(all_combinations_binarized_without_total[, 8:14]) >= 2 &         # At least 2 symptoms from 8-14
      rowSums(all_combinations_binarized_without_total[, 15:20]) >= 2          # At least 2 symptoms from 15-20
  )
  
  # Filter combinations based on criteria
  all_combinations_binarized_without_total_selectedPTSD <- filter(all_combinations_binarized_without_total, criteria_1=TRUE)
  
  # Count frequency of binarized profiles in real-world data
  binarized_data_counted <- renamed_data %>%
    group_by_all() %>%
    summarise(freq = n(), .groups = 'drop')
  
  # Merge theoretically possible combinations with their frequency in real-world data
  merged_binarized_possible_real_without_total <- left_join(all_combinations_binarized_without_total_selectedPTSD, binarized_data_counted, by = symptom_cols_binarized_without_total)
  merged_binarized_possible_real_without_total$freq <- coalesce(merged_binarized_possible_real_without_total$freq, 0)
 
   return(merged_binarized_possible_real_without_total)
  }
  ## Applying function to our data
possible_combinedwith_real_binarized_without_total <- create_possible_combinedwith_real_binarized_without_total(data2)

```

## x. Session Info

```{r Session Info, echo=FALSE, message=FALSE}
sessionInfo()
```
