---
title: "Assess Simulated PTSD Symptoms"
author:
- "<h5 style=\"font-style:italic\"> Laura Weidmann"
- "<h5 style=\"font-style:italic\"> Tobias R. Spiller"
date: "<h5 style=\"font-style:roman\"> `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
  pdf_document:
    toc: true
    toc_depth: '5'
subtitle: Version 0.0.4
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)
```

```{r Load Libraries, message=FALSE, warning=TRUE, include=FALSE}
# Data handling
library(tidyverse)
library(readr)
library(data.table)
library(scales)
library(DT)

# Demographics
library(table1)
library(gtsummary)

# Analysis
#library("devtools")
#devtools::install_github("orduek/PsychPower")
library(PsychPower)
```

## 1. Import and prepare data

```{r Import data, message=FALSE, warning=TRUE, include=FALSE}
# Import data
simulated_ptsd_data <- read_csv("Data/simulated_skewed_ptsd_data.csv")

# Renaming the columns in the input data (20 columns with non-binarized symptoms) to match the expected format
simulated_ptsd_data_renamed <- simulated_ptsd_data %>%
  rename_with(~ paste0("symptom_", 1:20))
```

## 2. Sample descriptive

### 2.1. Total score & Diagnosis

```{r Total score & Diagnosis, echo=FALSE, message=FALSE}
## Create PCL total score
simulated_ptsd_total <- simulated_ptsd_data_renamed %>% 
  mutate(total = rowSums(select(., symptom_1:symptom_20)))

## Create function that implements PTSD diagnostic criteria
  create_ptsd_diagnosis <- function(data) {
    
    # Create new column initially set to FALSE
    data$PTSD_Diagnosis <- FALSE
   
    # Check symptom criteria
    criterion_A <- rowSums(data[, paste0("symptom_", 1:5)] >= 2) >= 1  # At least 1 from symptoms 1-5
    criterion_B <- rowSums(data[, paste0("symptom_", 6:7)] >= 2) >= 1  # At least 1 from symptoms 6-7
    criterion_C <- rowSums(data[, paste0("symptom_", 8:14)] >= 2) >= 2 # At least 2 from symptoms 8-14
    criterion_D <- rowSums(data[, paste0("symptom_", 15:20)] >= 2) >= 2 # At least 2 from symptoms 15-20
    
    # All symptom criteria met
    all_symptom_criteria <- criterion_A & criterion_B & criterion_C & criterion_D
    
    data$PTSD_Diagnosis <- all_symptom_criteria
    
    return(data)
  }

# Apply the function to our dataframe
  simulated_ptsd_total <- create_ptsd_diagnosis (simulated_ptsd_total)

## Summarize
simulated_ptsd_total %>%
  summarise(
    mean_total = mean(total),
    sd_total = sd(total),
    n_diagnosed = sum(PTSD_Diagnosis)
  )
```

### 2.2. Cronbach's alpha

Selected sample

```{r PCL cronbach, echo=FALSE, message=FALSE}
cronbach <- psych::alpha(subset(simulated_ptsd_total, select = (-total)))
cronbach$total
```

### 2.3. Summary of items & histogram

```{r PCL summary, echo=FALSE, message=FALSE}
summary(simulated_ptsd_total)
hist(simulated_ptsd_total$total)
```

## 3. Binarization and selection of subset

```{r Binarization and selection of subset, echo=FALSE, message=FALSE}
## Binarization of all cases
  # Create copy of dataframe
  binarized_all <- simulated_ptsd_data_renamed
  # Binarize values (0,1 -> 0; 2,3,4 -> 1)
  binarized_all[binarized_all <= 1] <- 0
  binarized_all[binarized_all >= 2] <- 1

## Selection and binarization of diagnosed cases
  # Selecting diagnosed cases
  binarized_diagnosed <- simulated_ptsd_total %>%
    filter (PTSD_Diagnosis == TRUE) %>%
    select(starts_with("symptom_"))
  # Binarize values (0,1 -> 0; 2,3,4 -> 1)
  binarized_diagnosed[binarized_diagnosed <= 1] <- 0
  binarized_diagnosed[binarized_diagnosed >= 2] <- 1
```

## 4. Influence of deleting symptoms on the diagnosis

### 4.1. Function to determine the diagnosis with the original 20 symptoms

```{r Function to determine the diagnosis with the original 20 symptoms, echo=FALSE, message=FALSE}
### Function to analyze original symptoms
  # Input: Data set with all patients and 20 columns with non-binarized symptoms
analyze_ptsd_all <- function(data) {
  ## Function to check PTSD criteria based on symptoms
  check_ptsd_criteria <- function(symptoms) {
    # Check each criterion
    criterion_1 <- any(symptoms[1:5] == 1)
    criterion_2 <- any(symptoms[6:7] == 1)
    criterion_3 <- sum(symptoms[8:14] == 1) >= 2
    criterion_4 <- sum(symptoms[15:20] == 1) >= 2
    
    # All criteria must be met
    return(criterion_1 & criterion_2 & criterion_3 & criterion_4)
  }
  
  ## Binarization
  # Create copy of dataframe
  binarized_all <- data
  # Binarize values (0,1 -> 0; 2,3,4 -> 1)
  binarized_all[binarized_all <= 1] <- 0
  binarized_all[binarized_all >= 2] <- 1
  
  # Check PTSD criteria for each row
  ptsd_results <- apply(binarized_all, 1, check_ptsd_criteria)
  
  # Create new dataframe with results
  return(data.frame(PTSD_all = ptsd_results))
}
```

###  4.2. Function to determine the diagnosis when one or more symptoms are deleted manually

```{r Function to determine the diagnosis when one or more symptoms are deleted manually, echo=FALSE, message=FALSE}
### Function to analyze with one or more symptoms removed 
  # Input: Data: data set with all patients and 20 columns with non-binarized symptoms, symptoms_to_remove: symptom(s) you want to remove (multiple as vector)
analyze_ptsd_without_symptoms <- function(data, symptoms_to_remove) {
  ## Function to check PTSD criteria based on symptoms
  check_ptsd_criteria <- function(symptoms) {
    # Check each criterion
    criterion_1 <- any(symptoms[1:5] == 1)
    criterion_2 <- any(symptoms[6:7] == 1)
    criterion_3 <- sum(symptoms[8:14] == 1) >= 2
    criterion_4 <- sum(symptoms[15:20] == 1) >= 2
    
    # All criteria must be met
    return(criterion_1 & criterion_2 & criterion_3 & criterion_4)
  }
  
  ## Binarization
  # Create copy of dataframe
  binarized_all <- data
  # Binarize values (0,1 -> 0; 2,3,4 -> 1)
  binarized_all[binarized_all <= 1] <- 0
  binarized_all[binarized_all >= 2] <- 1
  
  # Set specified symptoms to 0 (effectively removing them)
  binarized_all[, symptoms_to_remove] <- 0
  
  # Check PTSD criteria for each row
  ptsd_results <- apply(binarized_all, 1, check_ptsd_criteria)
  
  # Create column name based on removed symptoms
  col_name <- paste0("PTSD_", paste(symptoms_to_remove, collapse = "_"))
  
  # Return results as dataframe
  result <- data.frame(matrix(ptsd_results, ncol = 1))
  names(result) <- col_name
  return(result)
}
```

###  4.3. Function to determine the diagnosis when all symptoms are deleted individually

```{r Function to determine the diagnosis when all symptoms are deleted individually, echo=FALSE, message=FALSE}
### Function to perform analysis for all possible single-symptom removals
  # Input: Data set with all patients and 20 columns with non-binarized symptoms
analyze_all_symptom_removals <- function(data) {
  # Get initial analysis with all symptoms
  results <- analyze_ptsd_all(data)
  
  # For each symptom
  for(i in 1:20) {
    # Analyze without this symptom
    result_without_i <- analyze_ptsd_without_symptoms(data, i)
    
    # Add to results dataframe
    results <- cbind(results, result_without_i)
  }
  
  return(results)
}
```
Dataframe with TRUE/FALSE for the diagnosis if each symptom is deleted individually (50 individuals to illustrate)
```{r Dataframe with TRUE/FALSE for the diagnosis if each symptom is deleted individually, echo=FALSE, message=FALSE}
### Applying to our data
PTSD_Diagnosis_individual <- analyze_all_symptom_removals(simulated_ptsd_data_renamed)
datatable(head(PTSD_Diagnosis_individual, 50),
          options = list(scrollX = TRUE))
```

###  4.4. Function to determine the diagnosis when several symptoms are deleted at the same time

```{r Function to determine the diagnosis when several symptoms are deleted at the same time, echo=FALSE, message=FALSE}
### Function to analyze combinations of multiple symptom removals (2 can be replaced by other numbers)
  # Input: data = Data set with all patients and 20 columns with non-binarized symptoms; existing_results: PTSD_Diagnosis_individual (created with analyze_all_symptom_removals(simulated_ptsd_data))
analyze_multiple_symptom_removals <- function(data, existing_results, max_combinations = 2) {
  # Start with existing results dataframe
  results <- existing_results
  
  ## Generate combinations for different numbers of symptoms
  for(n_symptoms in 2:max_combinations) {
    # Get all possible combinations of n symptoms
    symptom_combinations <- combn(1:20, n_symptoms, simplify = TRUE)
    
    # For each combination
    for(i in 1:ncol(symptom_combinations)) {
      # Get current combination of symptoms
      symptoms_to_remove <- symptom_combinations[,i]
      
      # Analyze without these symptoms
      result_without_symptoms <- analyze_ptsd_without_symptoms(data, symptoms_to_remove)
      
      # Add to results dataframe
      results <- cbind(results, result_without_symptoms)
    }
  }
  
  return(results)
}
```
Dataframe with TRUE/FALSE for the diagnosis of each single-symptom removal, two-combined-symptom removal and three-combined-symptom removal (10 individuals to illustrate)
```{r Dataframe with TRUE/FALSE for the diagnosis of each single-symptom removal, two-combined-symptom removal and three-combined-symptom removal, echo=FALSE, message=FALSE}
### Applying to our data for combination of two symptoms
PTSD_Diagnosis_combined_two <- analyze_multiple_symptom_removals(simulated_ptsd_data_renamed, PTSD_Diagnosis_individual, max_combinations = 2)

### Applying to our data for combination of three symptoms
PTSD_Diagnosis_combined_three <- analyze_multiple_symptom_removals(simulated_ptsd_data_renamed, PTSD_Diagnosis_individual, max_combinations = 3)
datatable(head(PTSD_Diagnosis_combined_three, 10),
          options = list(scrollX = TRUE))

# ### Applying to our data for combination of four symptoms
# PTSD_Diagnosis_combined_four <- analyze_multiple_symptom_removals(simulated_ptsd_data_renamed, PTSD_Diagnosis_individual, max_combinations = 4)
# datatable(head(PTSD_Diagnosis_combined_four, 10),
#           options = list(scrollX = TRUE))
```

###  4.5. Function to determine the changes in the PTSD diagnosis when deleting symptoms

#### 4.5.1. Overview of all individual and combined symptom deletions

```{r Overview of all individual and combined symptom deletions, echo=FALSE, message=FALSE}
### Function to summarize PTSD diagnosis changes
  # Input: dataframe PTSD_Diagnosis_individual (created with "analyze_all_symptom_removals(simulated_ptsd_data)") or PTSD_Diagnosis_combined_two/PTSD_Diagnosis_combined_three (created with "analyze_multiple_symptom_removals(simulated_ptsd_data, PTSD_Diagnosis_individual, max_combinations = 2/3)")
summarize_ptsd_changes <- function(data) {
  # Initialize results dataframe
  summary_stats <- data.frame(
    column = names(data),
    diagnosed = colSums(data),
    non_diagnosed = colSums(!data),
    stringsAsFactors = FALSE
  )
  
  # Calculate changes compared to PTSD_all
  baseline <- data$PTSD_all
  
  # For each column
  for(col in names(data)) {
    current <- data[[col]]
    
    # New diagnoses (FALSE in baseline, TRUE in current)
    newly_diagnosed <- sum(!baseline & current)
    
    # Lost diagnoses (TRUE in baseline, FALSE in current)
    newly_nondiagnosed <- sum(baseline & !current)

    # True positives (diagnosed in both)
    true_positive <- sum(baseline & current)
    
    # True negatives (non-diagnosed in both)
    true_negative <- sum(!baseline & !current)

    # Add to summary stats
    summary_stats$newly_diagnosed[summary_stats$column == col] <- newly_diagnosed
    summary_stats$newly_nondiagnosed[summary_stats$column == col] <- newly_nondiagnosed
    summary_stats$true_positive[summary_stats$column == col] <- true_positive
    summary_stats$true_negative[summary_stats$column == col] <- true_negative
    summary_stats$true_cases[summary_stats$column == col] <- true_positive + true_negative
    summary_stats$false_cases[summary_stats$column == col] <- newly_diagnosed + newly_nondiagnosed
  }
  
  # Calculate percentages
  total_cases <- nrow(data)
  summary_stats$diagnosed_percent <- round(summary_stats$diagnosed / total_cases * 100, 2)
  summary_stats$non_diagnosed_percent <- round(summary_stats$non_diagnosed / total_cases * 100, 2)
  
  return(summary_stats)
}

### Applying to our data
summary_PTSD_Diagnosis_individual <- summarize_ptsd_changes(PTSD_Diagnosis_individual)
summary_PTSD_Diagnosis_combined_two <- summarize_ptsd_changes(PTSD_Diagnosis_combined_two)
summary_PTSD_Diagnosis_combined_three <- summarize_ptsd_changes(PTSD_Diagnosis_combined_three)
# summary_PTSD_Diagnosis_combined_four <- summarize_ptsd_changes(PTSD_Diagnosis_combined_four)
```

#### 4.5.2. Overview of selected individual and/or combined symptom deletions

```{r Overview of selected individual and/or combined symptom deletions, echo=FALSE, message=FALSE}
### Function to create a more readable summary for specific columns
  # Input: resulting data frame of the function "summarize_ptsd_changes", columns: which symptom deletions should be displayed
create_readable_summary <- function(summary_stats, columns = NULL) {
  if(is.null(columns)) {
    summary_subset <- summary_stats
  } else {
    summary_subset <- summary_stats[summary_stats$column %in% columns,]
  }
  
  # Create readable summary
  readable_summary <- data.frame(
    Scenario = summary_subset$column,
    `Total Diagnosed` = paste0(summary_subset$diagnosed, 
                               " (", summary_subset$diagnosed_percent, "%)"),
    `Total Non-Diagnosed` = paste0(summary_subset$non_diagnosed,
                                   " (", summary_subset$non_diagnosed_percent, "%)"),
    `True Positive` = summary_subset$true_positive,
    `True Negative` = summary_subset$true_negative,
    `Newly Diagnosed` = summary_subset$newly_diagnosed,
    `Newly Non-Diagnosed` = summary_subset$newly_nondiagnosed,
    `True Cases` = summary_subset$true_cases,
    `False Cases` = summary_subset$false_cases
  )
  return(readable_summary)
}

### Summary for all single-symptom removals
single_removals <- c("PTSD_all", paste0("PTSD_", 1:20))
readable_summary_PTSD_Diagnosis_single <- create_readable_summary(summary_PTSD_Diagnosis_combined_two, single_removals)

### Summary for removal of 2 combined symptoms
two_combined_removals <- c("PTSD_all", summary_PTSD_Diagnosis_combined_two$column[grepl("_.*_", summary_PTSD_Diagnosis_combined_two$column)])
readable_summary_PTSD_Diagnosis_two_combined <- create_readable_summary(summary_PTSD_Diagnosis_combined_two, two_combined_removals)

### Summary for removal of 3 combined symptoms
three_combined_removals <- c("PTSD_all", summary_PTSD_Diagnosis_combined_three$column[grepl("_.*_.*_", summary_PTSD_Diagnosis_combined_three$column)])
readable_summary_PTSD_Diagnosis_three_combined <- create_readable_summary(summary_PTSD_Diagnosis_combined_three, three_combined_removals)
```

#### 4.5.2. Sorting the summary by newly non-diagnosed

Summary for all single-symptom removals sorted by newly non-diagnosed
```{r Summary for all single-symptom removals sorted by newly non-diagnosed, echo=FALSE, message=FALSE}
# Sort by newly non-diagnosed in single-symptom removals
most_lost_diagnoses_single <- summary_PTSD_Diagnosis_individual %>%
  arrange(desc(newly_nondiagnosed))
readable_most_lost_diagnoses_single <- create_readable_summary(most_lost_diagnoses_single)
datatable(readable_most_lost_diagnoses_single,
          options = list(scrollX = TRUE))
```

Summary for two-combined-symptom removals sorted by newly non-diagnosed
```{r Summary for two-combined-symptom removals sorted by newly non-diagnosed, echo=FALSE, message=FALSE}
# Sort by newly non-diagnosed in two-combined-symptom removals
most_lost_diagnoses_two_combined <- summary_PTSD_Diagnosis_combined_two %>%
  # Filter for PTSD_all and combined deletions (containing two underscores)
  filter(column == "PTSD_all" | grepl("_.*_", column)) %>%
  # Arrange by newly_nondiagnosed in descending order (desc())
  arrange(desc(newly_nondiagnosed))
  # Convert to readable format
readable_most_lost_diagnoses_two_combined <- create_readable_summary(most_lost_diagnoses_two_combined)
datatable(readable_most_lost_diagnoses_two_combined,
          options = list(scrollX = TRUE))
```

Summary for three-combined-symptom removals sorted by newly non-diagnosed (100 three-symptoms-combinations with most lost diagnoses)
```{r Summary for three-combined-symptom removals sorted by newly non-diagnosed, echo=FALSE, message=FALSE}
# Sort by newly non-diagnosed in three-combined-symptom removals
most_lost_diagnoses_three_combined <- summary_PTSD_Diagnosis_combined_three %>%
  # Filter for PTSD_all and combined deletions (containing two underscores)
  filter(column == "PTSD_all" | grepl("_.*_.*_", column)) %>%
  # Arrange by newly_nondiagnosed
  arrange(desc(newly_nondiagnosed))
  # Convert to readable format
readable_most_lost_diagnoses_three_combined <- create_readable_summary(most_lost_diagnoses_three_combined)
datatable(head(readable_most_lost_diagnoses_three_combined, 100),
          options = list(scrollX = TRUE))
```

## 5. Identification of the most important symptoms

### 5.1. Identification of the 4 most important symptoms

#### 5.1.1. Find three best four-symptom-combinations overall, regardless of which cluster they belong to

```{r Find three best four-symptom-combinations overall, regardless of which cluster they belong to, echo=FALSE, message=FALSE}
### Function to determine the three best four-symptom-combinations overall, regardless of which cluster they belong to.
  # Depending on the score_by select the best four-symptom-combinations by minimizing the newly_nondiagnosed or by minimizing the false_cases
  # Input: data set with all patients and 20 columns with non-binarized symptoms
analyze_best_four_symptoms <- function(data, score_by = "false_cases") {
  # Validate scoring method
  valid_scoring <- c("false_cases", "newly_nondiagnosed")
  if (!score_by %in% valid_scoring) {
    stop("score_by must be one of: ", paste(valid_scoring, collapse = ", "))
  }
  
  # Helper functions
  binarize_data <- function(data) {
    data[data <= 1] <- 0
    data[data >= 2] <- 1
    return(data == 1) # Convert to logical
  }
  
  # Get baseline results first
  baseline_results <- analyze_ptsd_all(data)$PTSD_all
  baseline_count <- sum(baseline_results)
  
  # Binarize data
  binarized_data <- binarize_data(data)
  
  # Helper function to get diagnosis based on having all specified symptoms
  get_diagnosis <- function(data, symptoms) {
    subset_data <- data[, symptoms]
    return(apply(subset_data, 1, all))
  }
  
  # Helper function to calculate metrics
  get_metrics <- function(current_diagnoses, baseline_results) {
    newly_diagnosed <- sum(current_diagnoses & !baseline_results)
    newly_nondiagnosed <- sum(!current_diagnoses & baseline_results)
    
    list(
      true_positives = sum(current_diagnoses & baseline_results),
      true_negatives = sum(!current_diagnoses & !baseline_results),
      newly_diagnosed = newly_diagnosed,
      newly_nondiagnosed = newly_nondiagnosed,
      false_cases = newly_diagnosed + newly_nondiagnosed,
      retention_rate = sum(current_diagnoses & baseline_results) / sum(baseline_results)
    )
  }
  
  # Initialize top combinations
  top_combinations <- list(
    first = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL),
    second = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL),
    third = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL)
  )
  
  # Get all possible combinations of 4 symptoms
  all_symptoms <- 1:20
  combinations <- combn(all_symptoms, 4, simplify = FALSE)
  
  # Test each combination
  for(combination in combinations) {
    symptom_names <- paste0("symptom_", combination)
    
    # Get diagnoses and calculate metrics
    current_diagnoses <- get_diagnosis(binarized_data, symptom_names)
    metrics <- get_metrics(current_diagnoses, baseline_results)
    
    # Score based on specified method
    score <- if(score_by == "false_cases") {
      -metrics$false_cases
    } else {
      -metrics$newly_nondiagnosed
    }
    
    # Update top combinations if needed
    if(score > top_combinations$first$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- top_combinations$first
      top_combinations$first <- list(
        combination = combination,
        score = score,
        diagnoses = current_diagnoses,
        metrics = metrics
      )
    } else if(score > top_combinations$second$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- list(
        combination = combination,
        score = score,
        diagnoses = current_diagnoses,
        metrics = metrics
      )
    } else if(score > top_combinations$third$score) {
      top_combinations$third <- list(
        combination = combination,
        score = score,
        diagnoses = current_diagnoses,
        metrics = metrics
      )
    }
  }
  
  # Create comparison dataframe
  make_column_name <- function(combination) {
    sprintf("symptom_%d_%d_%d_%d", combination[1], combination[2], combination[3], combination[4])
  }
  
  comparison_df <- data.frame(
    PTSD_all = baseline_results,
    top_combinations$first$diagnoses,
    top_combinations$second$diagnoses,
    top_combinations$third$diagnoses
  )
  names(comparison_df)[2:4] <- c(
    make_column_name(top_combinations$first$combination),
    make_column_name(top_combinations$second$combination),
    make_column_name(top_combinations$third$combination)
  )
  
  return(list(
    best_symptoms = top_combinations$first$combination,
    best_symptom_names = paste0("symptom_", top_combinations$first$combination),
    best_metrics = top_combinations$first$metrics,
    second_best_symptoms = top_combinations$second$combination,
    second_symptom_names = paste0("symptom_", top_combinations$second$combination),
    second_metrics = top_combinations$second$metrics,
    third_best_symptoms = top_combinations$third$combination,
    third_symptom_names = paste0("symptom_", top_combinations$third$combination),
    third_metrics = top_combinations$third$metrics,
    original_diagnoses = sum(baseline_results),
    diagnosis_comparison = comparison_df
  ))
}
```

Summary for top three four-symptom-combinations overall, regardless of which cluster they belong to. Best means fewest newly_nondiagnosed
```{r Summary for top three four-symptom-combinations overall, regardless of which cluster they belong to. Best means fewest newly_nondiagnosed, echo=FALSE, message=FALSE}
### Applying to our data
PTSD_Diagnosis_best_four_symptoms_overall_min_false_neg <- analyze_best_four_symptoms(simulated_ptsd_data_renamed, score_by = "newly_nondiagnosed")
# Summary
summary_PTSD_Diagnosis_best_four_symptoms_overall_min_false_neg <- summarize_ptsd_changes(PTSD_Diagnosis_best_four_symptoms_overall_min_false_neg$diagnosis_comparison)
# Readable summary
readable_summary_PTSD_Diagnosis_best_four_symptoms_overall_min_false_neg <- create_readable_summary(summary_PTSD_Diagnosis_best_four_symptoms_overall_min_false_neg)
datatable(readable_summary_PTSD_Diagnosis_best_four_symptoms_overall_min_false_neg,
          options = list(scrollX = TRUE))
```

Summary for top three four-symptom-combinations overall, regardless of which cluster they belong to. Best means fewest false_cases (newly_nondiagnosed + newly_diagnosed)
```{r Summary for top three four-symptom-combinations overall, regardless of which cluster they belong to. Best means fewest false_cases, echo=FALSE, message=FALSE}
### Applying to our data
PTSD_Diagnosis_best_four_symptoms_overall_min_false <- analyze_best_four_symptoms(simulated_ptsd_data_renamed, score_by = "false_cases")
# Summary
summary_PTSD_Diagnosis_best_four_symptoms_overall_min_false <- summarize_ptsd_changes(PTSD_Diagnosis_best_four_symptoms_overall_min_false$diagnosis_comparison)
# Readable summary
readable_summary_PTSD_Diagnosis_best_four_symptoms_overall_min_false <- create_readable_summary(summary_PTSD_Diagnosis_best_four_symptoms_overall_min_false)
# datatable(readable_summary_PTSD_Diagnosis_best_four_symptoms_overall_min_false,
#           options = list(scrollX = TRUE))
```

#### 5.1.2. Find best four symptom-combination, one symptom from each cluster 
Three different approaches to select from the 20 symptoms these 4 symptoms, one from each cluster, that lead to the fewest newly non-diagnosed cases respectively fewest false cases in total (newly non-diagnosed + newly diagnosed)

##### 5.1.2.1. Variant 1: Analysis of all possible combinations of 4 symptoms from one cluster each

```{r Variant 1: Find best four symptom-combination, one symptom from each cluster, echo=FALSE, message=FALSE}
### Function to analyze all combinations of 4 symptoms from different clusters with adapted diagnostic criteria, output: for each combination and each individual true/false of the PTSD diagnosis
  # Input: Data set with all patients and 20 columns with non-binarized symptoms
variant1_analyze_best_four_symptoms_by_cluster <- function(data) {
  # Define clusters
  cluster_1 <- 1:5   
  cluster_2 <- 6:7   
  cluster_3 <- 8:14  
  cluster_4 <- 15:20 
  
  # Get all possible combinations
  combinations <- expand.grid(
    cluster_1,
    cluster_2,
    cluster_3,
    cluster_4
  )
  
  # Get initial analysis with all symptoms for comparison
  results <- analyze_ptsd_all(data)
  
  # Function to check simplified PTSD criteria
  check_simplified_criteria <- function(symptoms, selected_symptoms) {
    # Check if each selected symptom is present (value of 1 after binarization)
    all(symptoms[selected_symptoms] == 1)
  }
  
  # Process each combination
  for(i in 1:nrow(combinations)) {
    current_symptoms <- as.numeric(combinations[i,])
    col_name <- paste0("PTSD_", paste(current_symptoms, collapse = "_"))
    
    # Create binarized data
    binarized_data <- data
    binarized_data[binarized_data <= 1] <- 0
    binarized_data[binarized_data >= 2] <- 1
    
    # Check criteria for each patient
    ptsd_results <- apply(binarized_data, 1, function(x) 
      check_simplified_criteria(x, current_symptoms))
    
    # Add to results dataframe
    results[[col_name]] <- ptsd_results
  }
  
  return(results)
}
```

##### 5.1.2.2. Variant 2: Find three best combinations of 4 symptoms from one cluster each, by analyzing individual importance, selecting top 2 candidates and testing different combinations

```{r Variant 2: Find best four symptom-combination, one symptom from each cluster, echo=FALSE, message=FALSE}
### Function to analyze three best four-symptom combination by: 1. analyzing individual importance of symptoms, 2. selecting most important symptoms from each cluster as candidates, 3. testing different combinations of these candidates to chose best
  # Input: Data set with all patients and 20 columns with non-binarized symptoms
variant2_analyze_best_four_symptoms_by_cluster <- function(data, score_by = "false_cases") {
  # Validate scoring method
  valid_scoring <- c("false_cases", "newly_nondiagnosed")
  if (!score_by %in% valid_scoring) {
    stop("score_by must be one of: ", paste(valid_scoring, collapse = ", "))
  }
  
  # Initialize top combinations
  top_combinations <- list(
    first = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL),
    second = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL),
    third = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL)
  )
  
  # Helper functions
  binarize_data <- function(data) {
    data[data <= 1] <- 0
    data[data >= 2] <- 1
    return(data == 1) # Convert to logical
  }
  
  # Get baseline results first
  baseline_results <- analyze_ptsd_all(data)$PTSD_all
  baseline_count <- sum(baseline_results)
  
  # Binarize data
  binarized_data <- binarize_data(data)
  
  # Get_candidates function (only considering diagnosed cases)
  get_candidates <- function(data, range, diagnosed_cases) {
    # Subset data to only include rows where PTSD_all is TRUE
    diagnosed_data <- data[diagnosed_cases, ]
    
    # Calculate frequencies only for diagnosed cases
    cluster_data <- diagnosed_data[, paste0("symptom_", range)] 
    symptom_freq <- colSums(cluster_data) # adds up all TRUE values in each column
    
    # Get top 2 most frequent symptoms
    top_indices <- order(symptom_freq, decreasing=TRUE)[1:2] # takes the two symptoms in this range with highest freq 
    return(range[top_indices]) # range converts back the order to original
  }
  
  # Get candidates from each cluster using only diagnosed cases
  candidates <- list(
    cluster1 = get_candidates(binarized_data, 1:5, baseline_results),
    cluster2 = get_candidates(binarized_data, 6:7, baseline_results),
    cluster3 = get_candidates(binarized_data, 8:14, baseline_results),
    cluster4 = get_candidates(binarized_data, 15:20, baseline_results)
  )
  
  get_diagnosis <- function(data, symptoms) {
    subset_data <- data[, symptoms]    # Select only the columns for the specified symptoms
    return(apply(subset_data, 1, all))  # For each row (person), check if ALL symptoms are TRUE
  }
  
  get_metrics <- function(current_diagnoses, baseline_results) {
    newly_diagnosed <- sum(current_diagnoses & !baseline_results)
    newly_nondiagnosed <- sum(!current_diagnoses & baseline_results)
    
    list(
      true_positives = sum(current_diagnoses & baseline_results),
      true_negatives = sum(!current_diagnoses & !baseline_results),
      newly_diagnosed = newly_diagnosed,
      newly_nondiagnosed = newly_nondiagnosed,
      false_cases = newly_diagnosed + newly_nondiagnosed,
      retention_rate = sum(current_diagnoses & baseline_results) / sum(baseline_results)
    )
  }
  
  # update_top_combinations helper function
  update_top_combinations <- function(top_combinations, current_combination, score, diagnoses, metrics) {
    if(score > top_combinations$first$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- top_combinations$first
      top_combinations$first <- list(
        combination = current_combination,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics
      )
    } else if(score > top_combinations$second$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- list(
        combination = current_combination,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics
      )
    } else if(score > top_combinations$third$score) {
      top_combinations$third <- list(
        combination = current_combination,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics
      )
    }
    return(top_combinations)
  }
  
  # Generate and test all combinations
  for(s1 in candidates$cluster1) {
    for(s2 in candidates$cluster2) {
      for(s3 in candidates$cluster3) {
        for(s4 in candidates$cluster4) {
          current_combination <- c(s1, s2, s3, s4)
          symptom_names <- paste0("symptom_", current_combination)
          
          # Get diagnoses and calculate score
          current_diagnoses <- get_diagnosis(binarized_data, symptom_names)
          metrics <- get_metrics(current_diagnoses, baseline_results)
          
          # Score based on specified method
          score <- if(score_by == "false_cases") {
            -(metrics$newly_diagnosed + metrics$newly_nondiagnosed)
          } else {
            -metrics$newly_nondiagnosed
          }
          
          # Update top combinations if needed
          top_combinations <- update_top_combinations(
            top_combinations, 
            current_combination, 
            score, 
            current_diagnoses, 
            metrics
          )
          
        }
      }
    }
  }
  
  # Create comparison dataframe
  make_column_name <- function(combination) {
    sprintf("symptom_%d_%d_%d_%d", combination[1], combination[2], 
            combination[3], combination[4])
  }
  
  comparison_df <- data.frame(
    PTSD_all = baseline_results,
    top_combinations$first$diagnoses,
    top_combinations$second$diagnoses,
    top_combinations$third$diagnoses
  )
  names(comparison_df)[2:4] <- c(
    make_column_name(top_combinations$first$combination),
    make_column_name(top_combinations$second$combination),
    make_column_name(top_combinations$third$combination)
  )
  
  return(list(
    best_symptoms = top_combinations$first$combination,
    best_symptom_names = paste0("symptom_", top_combinations$first$combination),
    best_metrics = top_combinations$first$metrics,
    second_best_symptoms = top_combinations$second$combination,
    second_symptom_names = paste0("symptom_", top_combinations$second$combination),
    second_metrics = top_combinations$second$metrics,
    third_best_symptoms = top_combinations$third$combination,
    third_symptom_names = paste0("symptom_", top_combinations$third$combination),
    third_metrics = top_combinations$third$metrics,
    original_diagnoses = sum(baseline_results),
    diagnosis_comparison = comparison_df
  ))
}
```

##### 5.1.2.3. Variant 3: Find three best combinations of 4 symptoms from one cluster each, by testing each combination individually

```{r Variant 3: Find best four symptom-combination, one symptom from each cluster, echo=FALSE, message=FALSE}
### Function to analyze three best four symptom-combinations from four different clusters by testing each combination individually
  # Depending on the score_by select the best three-symptom-combination by minimizing the newly_nondiagnosed or by minimizing the false_cases
  # Input: Data set with all patients and 20 columns with non-binarized symptoms
variant3_analyze_best_four_symptoms_by_cluster <- function(data, score_by = "false_cases") {
  # Validate scoring method
  valid_scoring <- c("false_cases", "newly_nondiagnosed")
  if (!score_by %in% valid_scoring) {
    stop("score_by must be one of: ", paste(valid_scoring, collapse = ", "))
  }
  
  # Define clusters
  clusters <- list(
    cluster1 = 1:5,    # Criterion A
    cluster2 = 6:7,    # Criterion B
    cluster3 = 8:14,   # Criterion C
    cluster4 = 15:20   # Criterion D
  )
  
  # Helper functions
  binarize_data <- function(data) {
    data[data <= 1] <- 0
    data[data >= 2] <- 1
    return(data == 1)
  }
  
  # Get baseline results first
  baseline_results <- analyze_ptsd_all(data)$PTSD_all
  baseline_count <- sum(baseline_results)
  
  # Binarize data
  binarized_data <- binarize_data(data)
  
  get_diagnosis <- function(data, symptoms) {
    subset_data <- data[, symptoms]
    return(apply(subset_data, 1, all))
  }
  
  get_metrics <- function(current_diagnoses, baseline_results) {
    newly_diagnosed <- sum(current_diagnoses & !baseline_results)
    newly_nondiagnosed <- sum(!current_diagnoses & baseline_results)
    
    list(
      true_positives = sum(current_diagnoses & baseline_results),
      true_negatives = sum(!current_diagnoses & !baseline_results),
      newly_diagnosed = newly_diagnosed,
      newly_nondiagnosed = newly_nondiagnosed,
      false_cases = newly_diagnosed + newly_nondiagnosed,
      retention_rate = sum(current_diagnoses & baseline_results) / sum(baseline_results)
    )
  }

  # Initialize top combinations 
  top_combinations <- list(
    first = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL),
    second = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL),
    third = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL)
  )
  
  # update_top_combinations helper function
  update_top_combinations <- function(top_combinations, current_combination, score, diagnoses, metrics) {
    if(score > top_combinations$first$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- top_combinations$first
      top_combinations$first <- list(
        combination = current_combination,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics
      )
    } else if(score > top_combinations$second$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- list(
        combination = current_combination,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics
      )
    } else if(score > top_combinations$third$score) {
      top_combinations$third <- list(
        combination = current_combination,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics
      )
    }
    return(top_combinations)
  }
  
  
  # Get one symptom from each cluster
  for(symptom1 in clusters$cluster1) {
    for(symptom2 in clusters$cluster2) {
      for(symptom3 in clusters$cluster3) {
        for(symptom4 in clusters$cluster4) { 
          current_combination <- c(symptom1, symptom2, symptom3, symptom4)
          symptom_names <- paste0("symptom_", current_combination)
          
          # Get diagnoses and calculate metrics
          current_diagnoses <- get_diagnosis(binarized_data, symptom_names)
          metrics <- get_metrics(current_diagnoses, baseline_results)
          
          # Score based on specified method
          score <- if(score_by == "false_cases") {
            -metrics$false_cases
          } else {
            -metrics$newly_nondiagnosed
          }
          
         # Update top combinations
          top_combinations <- update_top_combinations(
            top_combinations, 
            current_combination, 
            score, 
            current_diagnoses, 
            metrics
          )
        }
      }
    }
  }
  
  # Create comparison dataframe
  make_column_name <- function(combination) {
    sprintf("symptom_%d_%d_%d_%d", combination[1], combination[2], 
            combination[3], combination[4])
  }
  
  comparison_df <- data.frame(
    PTSD_all = baseline_results,
    top_combinations$first$diagnoses,
    top_combinations$second$diagnoses,
    top_combinations$third$diagnoses
  )
  
  names(comparison_df)[2:4] <- c(
    make_column_name(top_combinations$first$combination),
    make_column_name(top_combinations$second$combination),
    make_column_name(top_combinations$third$combination)
  )
  
  return(list(
    best_symptoms = top_combinations$first$combination,
    best_symptom_names = paste0("symptom_", top_combinations$first$combination),
    best_metrics = top_combinations$first$metrics,
    second_best_symptoms = top_combinations$second$combination,
    second_symptom_names = paste0("symptom_", top_combinations$second$combination),
    second_metrics = top_combinations$second$metrics,
    third_best_symptoms = top_combinations$third$combination,
    third_symptom_names = paste0("symptom_", top_combinations$third$combination),
    third_metrics = top_combinations$third$metrics,
    original_diagnoses = sum(baseline_results),
    diagnosis_comparison = comparison_df
  ))
}
```

Variant 1: Summary of all four-symptom-combinations sorted by newly non-diagnosed identified with variant 1 (10 four-symptoms-combinations with fewest lost diagnoses)
```{r Variant 1: Summary of all four-symptom-combinations sorted by newly non-diagnosed, echo=FALSE, message=FALSE}
### Applying to our data
variant1_PTSD_Diagnosis_best_four_symptoms_by_cluster <- variant1_analyze_best_four_symptoms_by_cluster(simulated_ptsd_data_renamed)
  # Summary
summary_variant1_PTSD_Diagnosis_best_four_symptoms_by_cluster <- summarize_ptsd_changes(variant1_PTSD_Diagnosis_best_four_symptoms_by_cluster)
  # Readable Summary
readable_summary_variant1_PTSD_Diagnosis_best_four_symptoms_by_cluster <- create_readable_summary(summary_variant1_PTSD_Diagnosis_best_four_symptoms_by_cluster) 
 # Sort by newly non-diagnosed 
summary_variant1_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false_neg <- summary_variant1_PTSD_Diagnosis_best_four_symptoms_by_cluster %>%
  arrange(newly_nondiagnosed)
readable_summary_variant1_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false_neg <- create_readable_summary(summary_variant1_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false_neg)
datatable(head(readable_summary_variant1_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false_neg, 11),
          options = list(scrollX = TRUE))
```

Variant 1:  Summary of all four-symptom-combinations sorted by newly false cases (newly diagnosed + newly nondiagnosed) identified with variant 1 (10 four-symptoms-combinations with fewest false cases)
```{r Variant 1: Summary of all four-symptom-combinations sorted by newly false cases, echo=FALSE, message=FALSE}
  # Sort by newly false cases (newly diagnosed + newly nondiagnosed)
summary_variant1_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false <- summary_variant1_PTSD_Diagnosis_best_four_symptoms_by_cluster %>%
  arrange(false_cases)
readable_summary_variant1_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false <- create_readable_summary(summary_variant1_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false)
# datatable(head(readable_summary_variant1_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false, 11),
#           options = list(scrollX = TRUE))
```

Variant 2: Best three four-symptom combinations with minimized newly_nondiagnosed identified with variant 2
```{r Variant 2: Best four-symptom combination with minimized newly_nondiagnosed, echo=FALSE, message=FALSE}
### Applying to our data
variant2_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false_neg <- variant2_analyze_best_four_symptoms_by_cluster(simulated_ptsd_data_renamed, score_by = "newly_nondiagnosed")
# Summary
summary_variant2_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false_neg <- summarize_ptsd_changes(variant2_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false_neg$diagnosis_comparison)
# Readable summary
readable_summary_variant2_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false_neg <- create_readable_summary(summary_variant2_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false_neg)
datatable(readable_summary_variant2_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false_neg,
          options = list(scrollX = TRUE))
```

Variant 2: Best three four-symptom combinations with minimized false_cases (newly_nondiagnosed + newly_diagnosed) identified with variant 2
```{r Variant 2: Best four-symptom combination with minimized false_cases, echo=FALSE, message=FALSE}
### Applying to our data
variant2_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false <- variant2_analyze_best_four_symptoms_by_cluster(simulated_ptsd_data_renamed, score_by = "false_cases")
# Summary
summary_variant2_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false <- summarize_ptsd_changes(variant2_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false$diagnosis_comparison)
# Readable Summary
readable_summary_variant2_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false <- create_readable_summary(summary_variant2_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false)
# datatable(readable_summary_variant2_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false,
#           options = list(scrollX = TRUE))
```

Variant 3: Best three four-symptom combinations with minimized newly_nondiagnosed identified with variant 3
```{r Variant 3: Best four-symptom combination with minimized newly_nondiagnosed, echo=FALSE, message=FALSE}
### Applying to our data
variant3_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false_neg <- variant3_analyze_best_four_symptoms_by_cluster(simulated_ptsd_data_renamed, score_by = "newly_nondiagnosed")
# Summary
summary_variant3_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false_neg <- summarize_ptsd_changes(variant3_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false_neg$diagnosis_comparison)
# Readable summary
readable_summary_variant3_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false_neg <- create_readable_summary(summary_variant3_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false_neg)
datatable(readable_summary_variant3_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false_neg,
          options = list(scrollX = TRUE))
```

Variant 3: Best three four-symptom combinations with minimized false_cases (newly_nondiagnosed + newly_diagnosed) identified with variant 3
```{r Variant 3: Best four-symptom combination with minimized false_cases, echo=FALSE, message=FALSE}
### Applying to our data
variant3_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false <- variant3_analyze_best_four_symptoms_by_cluster(simulated_ptsd_data_renamed, score_by = "false_cases")
# Summary
summary_variant3_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false <- summarize_ptsd_changes(variant3_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false$diagnosis_comparison)
# Readable summary
readable_summary_variant3_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false <- create_readable_summary(summary_variant3_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false)
# datatable(readable_summary_variant3_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false,
#           options = list(scrollX = TRUE))
```

### 5.2. Identification of the 3 most important symptoms

#### 5.2.1. Find three best three-symptom-combinations overall, regardless of which cluster they belong to.

```{r Find three best three-symptom-combinations overall, regardless of which cluster they belong to, echo=FALSE, message=FALSE}
### Function to determine the three best three-symptom-combinations overall, regardless of which cluster they belong to. 
  # Depending on the score_by select the best three-symptom-combination by minimizing the newly_nondiagnosed or by minimizing the false_cases
  # Input: Data set with all patients and 20 columns with non-binarized symptoms
analyze_best_three_symptoms <- function(data, score_by = "false_cases") {
  # Validate scoring method
  valid_scoring <- c("false_cases", "newly_nondiagnosed")
  if (!score_by %in% valid_scoring) {
    stop("score_by must be one of: ", paste(valid_scoring, collapse = ", "))
  }
  
  # Helper functions
  binarize_data <- function(data) {
    data[data <= 1] <- 0
    data[data >= 2] <- 1
    return(data == 1) # Convert to logical
  }
  
  # Get baseline results first
  baseline_results <- analyze_ptsd_all(data)$PTSD_all
  baseline_count <- sum(baseline_results)
  
  # Binarize data
  binarized_data <- binarize_data(data)
  
  # Helper function to get diagnosis based on having all specified symptoms
  get_diagnosis <- function(data, symptoms) {
    subset_data <- data[, symptoms]
    return(apply(subset_data, 1, all))
  }
  
  # Helper function to calculate metrics
  get_metrics <- function(current_diagnoses, baseline_results) {
    newly_diagnosed <- sum(current_diagnoses & !baseline_results)
    newly_nondiagnosed <- sum(!current_diagnoses & baseline_results)
    
    list(
      true_positives = sum(current_diagnoses & baseline_results),
      true_negatives = sum(!current_diagnoses & !baseline_results),
      newly_diagnosed = newly_diagnosed,
      newly_nondiagnosed = newly_nondiagnosed,
      false_cases = newly_diagnosed + newly_nondiagnosed,
      retention_rate = sum(current_diagnoses & baseline_results) / sum(baseline_results)
    )
  }
  
  # Initialize top combinations
  top_combinations <- list(
    first = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL),
    second = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL),
    third = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL)
  )
  
  # Get all possible combinations of 3 symptoms
  all_symptoms <- 1:20
  combinations <- combn(all_symptoms, 3, simplify = FALSE)
  
  # Test each combination
  for(combination in combinations) {
    symptom_names <- paste0("symptom_", combination)
    
    # Get diagnoses and calculate metrics
    current_diagnoses <- get_diagnosis(binarized_data, symptom_names)
    metrics <- get_metrics(current_diagnoses, baseline_results)
    
    # Score based on specified method
    score <- if(score_by == "false_cases") {
      -metrics$false_cases
    } else {
      -metrics$newly_nondiagnosed
    }
    
    # Update top combinations if needed
    if(score > top_combinations$first$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- top_combinations$first
      top_combinations$first <- list(
        combination = combination,
        score = score,
        diagnoses = current_diagnoses,
        metrics = metrics
      )
    } else if(score > top_combinations$second$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- list(
        combination = combination,
        score = score,
        diagnoses = current_diagnoses,
        metrics = metrics
      )
    } else if(score > top_combinations$third$score) {
      top_combinations$third <- list(
        combination = combination,
        score = score,
        diagnoses = current_diagnoses,
        metrics = metrics
      )
    }
  }
  
  # Create comparison dataframe
  make_column_name <- function(combination) {
    sprintf("symptom_%d_%d_%d", combination[1], combination[2], combination[3])
  }
  
  comparison_df <- data.frame(
    PTSD_all = baseline_results,
    top_combinations$first$diagnoses,
    top_combinations$second$diagnoses,
    top_combinations$third$diagnoses
  )
  names(comparison_df)[2:4] <- c(
    make_column_name(top_combinations$first$combination),
    make_column_name(top_combinations$second$combination),
    make_column_name(top_combinations$third$combination)
  )
  
  return(list(
    best_symptoms = top_combinations$first$combination,
    best_symptom_names = paste0("symptom_", top_combinations$first$combination),
    best_metrics = top_combinations$first$metrics,
    second_best_symptoms = top_combinations$second$combination,
    second_symptom_names = paste0("symptom_", top_combinations$second$combination),
    second_metrics = top_combinations$second$metrics,
    third_best_symptoms = top_combinations$third$combination,
    third_symptom_names = paste0("symptom_", top_combinations$third$combination),
    third_metrics = top_combinations$third$metrics,
    original_diagnoses = sum(baseline_results),
    diagnosis_comparison = comparison_df
  ))
}
```

Summary for three best three-symptom-combinations overall, regardless of which cluster they belong to. Best means fewest newly_nondiagnosed
```{r Summary for three best three-symptom-combinations overall, regardless of which cluster they belong to. Best means fewest newly_nondiagnosed, echo=FALSE, message=FALSE}
### Applying to our data
PTSD_Diagnosis_best_three_symptoms_overall_min_false_neg <- analyze_best_three_symptoms(simulated_ptsd_data_renamed, score_by = "newly_nondiagnosed")
# Summary
summary_PTSD_Diagnosis_best_three_symptoms_overall_min_false_neg <- summarize_ptsd_changes(PTSD_Diagnosis_best_three_symptoms_overall_min_false_neg$diagnosis_comparison)
# Readable summary
readable_summary_PTSD_Diagnosis_best_three_symptoms_overall_min_false_neg <- create_readable_summary(summary_PTSD_Diagnosis_best_three_symptoms_overall_min_false_neg)
datatable(readable_summary_PTSD_Diagnosis_best_three_symptoms_overall_min_false_neg,
          options = list(scrollX = TRUE))
```

Summary for three best three-symptom-combinations overall, regardless of which cluster they belong to. Best means fewest false_cases (newly_nondiagnosed + newly_diagnosed)
```{r Summary for three best three-symptom-combinations overall, regardless of which cluster they belong to. Best means fewest false_cases, echo=FALSE, message=FALSE}
### Applying to our data
PTSD_Diagnosis_best_three_symptoms_overall_min_false <- analyze_best_three_symptoms(simulated_ptsd_data_renamed, score_by = "false_cases")
# Summary
summary_PTSD_Diagnosis_best_three_symptoms_overall_min_false <- summarize_ptsd_changes(PTSD_Diagnosis_best_three_symptoms_overall_min_false$diagnosis_comparison)
# Readable summary
readable_summary_PTSD_Diagnosis_best_three_symptoms_overall_min_false <- create_readable_summary(summary_PTSD_Diagnosis_best_three_symptoms_overall_min_false)
# datatable(readable_summary_PTSD_Diagnosis_best_three_symptoms_overall_min_false,
#           options = list(scrollX = TRUE))
```

#### 5.2.2. Find three best three-symptom-combinations, each symptom from a different cluster (excluding one cluster at a time)

```{r Find three best three-symptom-combinations, each symptom from a different cluster (excluding one cluster at a time), echo=FALSE, message=FALSE}
### Function to analyze three best three symptom-combinations from three different clusters (excluding one cluster at a time)
  # Depending on the score_by select the best three-symptom-combination by minimizing the newly_nondiagnosed or by minimizing the false_cases
  # Input: Data set with all patients and 20 columns with non-binarized symptoms
analyze_best_three_symptoms_by_cluster <- function(data, score_by = "false_cases") {
  # Validate scoring method
  valid_scoring <- c("false_cases", "newly_nondiagnosed")
  if (!score_by %in% valid_scoring) {
    stop("score_by must be one of: ", paste(valid_scoring, collapse = ", "))
  }
  
  # Define clusters
  clusters <- list(
    cluster1 = 1:5,    # Criterion A
    cluster2 = 6:7,    # Criterion B
    cluster3 = 8:14,   # Criterion C
    cluster4 = 15:20   # Criterion D
  )
  
  # Helper functions
  binarize_data <- function(data) {
    data[data <= 1] <- 0
    data[data >= 2] <- 1
    return(data == 1)
  }
  
  # Get baseline results first
  baseline_results <- analyze_ptsd_all(data)$PTSD_all
  baseline_count <- sum(baseline_results)
  
  # Binarize data
  binarized_data <- binarize_data(data)
  
  get_diagnosis <- function(data, symptoms) {
    subset_data <- data[, symptoms]
    return(apply(subset_data, 1, all))
  }
  
  get_metrics <- function(current_diagnoses, baseline_results) {
    newly_diagnosed <- sum(current_diagnoses & !baseline_results)
    newly_nondiagnosed <- sum(!current_diagnoses & baseline_results)
    
    list(
      true_positives = sum(current_diagnoses & baseline_results),
      true_negatives = sum(!current_diagnoses & !baseline_results),
      newly_diagnosed = newly_diagnosed,
      newly_nondiagnosed = newly_nondiagnosed,
      false_cases = newly_diagnosed + newly_nondiagnosed,
      retention_rate = sum(current_diagnoses & baseline_results) / sum(baseline_results)
    )
  }
  
  # Helper function to update top 3 combinations
  update_top_combinations <- function(top_combinations, current_combination, score, diagnoses, metrics, excluded_cluster) {
    if(score > top_combinations$first$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- top_combinations$first
      top_combinations$first <- list(
        combination = current_combination,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics,
        excluded_cluster = excluded_cluster
      )
    } else if(score > top_combinations$second$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- list(
        combination = current_combination,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics,
        excluded_cluster = excluded_cluster
      )
    } else if(score > top_combinations$third$score) {
      top_combinations$third <- list(
        combination = current_combination,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics,
        excluded_cluster = excluded_cluster
      )
    }
    return(top_combinations)
  }
  
  # Initialize results list
  results_by_excluded_cluster <- list()
  
  # For each cluster to exclude
  for(exclude_cluster in 1:4) {
    # Get remaining clusters
    remaining_clusters <- clusters[setdiff(1:4, exclude_cluster)]
    
    # Initialize top 3 combinations for this excluded cluster
    top_combinations <- list(
      first = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL, excluded_cluster = exclude_cluster),
      second = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL, excluded_cluster = exclude_cluster),
      third = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL, excluded_cluster = exclude_cluster)
    )
    
    # Generate all possible combinations from remaining clusters
    combinations <- expand.grid(
      remaining_clusters[[1]],
      remaining_clusters[[2]],
      remaining_clusters[[3]]
    )
    
    # Test each combination
    for(i in 1:nrow(combinations)) {
      current_combination <- as.numeric(combinations[i,])
      symptom_names <- paste0("symptom_", current_combination)
      
      # Get diagnoses and calculate metrics
      current_diagnoses <- get_diagnosis(binarized_data, symptom_names)
      metrics <- get_metrics(current_diagnoses, baseline_results)
      
      # Score based on specified method
      score <- if(score_by == "false_cases") {
        -metrics$false_cases
      } else {
        -metrics$newly_nondiagnosed
      }
      
      # Update top combinations
      top_combinations <- update_top_combinations(
        top_combinations,
        current_combination,
        score,
        current_diagnoses,
        metrics,
        exclude_cluster
      )
    }
    
    # Store results for this excluded cluster
    results_by_excluded_cluster[[exclude_cluster]] <- top_combinations
  }
  
  # Create comparison dataframe
  comparison_df <- data.frame(PTSD_all = baseline_results)
  
  # Add all combinations to comparison dataframe
  for(i in 1:4) {
    # Add first combination
    col_name <- sprintf("exclude_cluster%d_%d_%d_%d", 
                        i,
                        results_by_excluded_cluster[[i]]$first$combination[1],
                        results_by_excluded_cluster[[i]]$first$combination[2],
                        results_by_excluded_cluster[[i]]$first$combination[3])
    comparison_df[[col_name]] <- results_by_excluded_cluster[[i]]$first$diagnoses
  
    # Add second combination
    col_name <- sprintf("exclude_cluster%d_second_%d_%d_%d", 
                       i,
                       results_by_excluded_cluster[[i]]$second$combination[1],
                       results_by_excluded_cluster[[i]]$second$combination[2],
                       results_by_excluded_cluster[[i]]$second$combination[3])
    comparison_df[[col_name]] <- results_by_excluded_cluster[[i]]$second$diagnoses
    
    # Add third combination
    col_name <- sprintf("exclude_cluster%d_third_%d_%d_%d", 
                       i,
                       results_by_excluded_cluster[[i]]$third$combination[1],
                       results_by_excluded_cluster[[i]]$third$combination[2],
                       results_by_excluded_cluster[[i]]$third$combination[3])
    comparison_df[[col_name]] <- results_by_excluded_cluster[[i]]$third$diagnoses
  }
  
  return(list(
    results_by_excluded_cluster = results_by_excluded_cluster,
    original_diagnoses = sum(baseline_results),
    diagnosis_comparison = comparison_df
  ))
}
```

Summary for three best three-symptom-combinations, each symptom from a different cluster (excluding one cluster at a time). Best means fewest newly_nondiagnosed
```{r Summary for three best three-symptom-combinations, each symptom from a different cluster (excluding one cluster at a time). Best means fewest newly_nondiagnosed, echo=FALSE, message=FALSE}
### Applying to our data
PTSD_Diagnosis_best_three_symptoms_by_cluster_min_false_neg <- analyze_best_three_symptoms_by_cluster(simulated_ptsd_data_renamed, score_by = "newly_nondiagnosed")
# Summary
summary_PTSD_Diagnosis_best_three_symptoms_by_cluster_min_false_neg <- summarize_ptsd_changes(PTSD_Diagnosis_best_three_symptoms_by_cluster_min_false_neg$diagnosis_comparison)
# Readable summary
readable_summary_PTSD_Diagnosis_best_three_symptoms_by_cluster_min_false_neg <- create_readable_summary(summary_PTSD_Diagnosis_best_three_symptoms_by_cluster_min_false_neg)
datatable(readable_summary_PTSD_Diagnosis_best_three_symptoms_by_cluster_min_false_neg,
          options = list(scrollX = TRUE))
```

Summary for three best three-symptom-combinations, each symptom from a different cluster (excluding one cluster at a time). Best means fewest false_cases (newly_nondiagnosed + newly_diagnosed)
```{r Summary for three best three-symptom-combinations, each symptom from a different cluster (excluding one cluster at a time). Best means fewest false_cases, echo=FALSE, message=FALSE}
### Applying to our data
PTSD_Diagnosis_best_three_symptoms_by_cluster_min_false <- analyze_best_three_symptoms_by_cluster(simulated_ptsd_data_renamed, score_by = "false_cases")
# Summary
summary_PTSD_Diagnosis_best_three_symptoms_by_cluster_min_false <- summarize_ptsd_changes(PTSD_Diagnosis_best_three_symptoms_by_cluster_min_false$diagnosis_comparison)
# Readable summary
readable_summary_PTSD_Diagnosis_best_three_symptoms_by_cluster_min_false <- create_readable_summary(summary_PTSD_Diagnosis_best_three_symptoms_by_cluster_min_false)
# datatable(readable_summary_PTSD_Diagnosis_best_three_symptoms_by_cluster_min_false,
#           options = list(scrollX = TRUE))
```

### 5.3. Identification of the 2 most important symptoms

#### 5.3.1. Find three best two-symptom-combinations overall, regardless of which cluster they belong to.

```{r Find three best two-symptom-combinations overall, regardless of which cluster they belong to, echo=FALSE, message=FALSE}
### Function to determine the three best two-symptom-combinations overall, regardless of which cluster they belong to. 
  # Depending on the score_by select the best three-symptom-combination by minimizing the newly_nondiagnosed or by minimizing the false_cases
  # Input: Data set with all patients and 20 columns with non-binarized symptoms
analyze_best_two_symptoms <- function(data, score_by = "false_cases") {
  # Validate scoring method
  valid_scoring <- c("false_cases", "newly_nondiagnosed")
  if (!score_by %in% valid_scoring) {
    stop("score_by must be one of: ", paste(valid_scoring, collapse = ", "))
  }
  
  # Helper functions
  binarize_data <- function(data) {
    data[data <= 1] <- 0
    data[data >= 2] <- 1
    return(data == 1)
  }
  
  # Get baseline results first
  baseline_results <- analyze_ptsd_all(data)$PTSD_all
  baseline_count <- sum(baseline_results)
  
  # Binarize data
  binarized_data <- binarize_data(data)
  
  # Helper function to get diagnosis based on having all specified symptoms
  get_diagnosis <- function(data, symptoms) {
    subset_data <- data[, symptoms]
    return(apply(subset_data, 1, all))
  }
  
  # Helper function to calculate metrics
  get_metrics <- function(current_diagnoses, baseline_results) {
    newly_diagnosed <- sum(current_diagnoses & !baseline_results)
    newly_nondiagnosed <- sum(!current_diagnoses & baseline_results)
    
    list(
      true_positives = sum(current_diagnoses & baseline_results),
      true_negatives = sum(!current_diagnoses & !baseline_results),
      newly_diagnosed = newly_diagnosed,
      newly_nondiagnosed = newly_nondiagnosed,
      false_cases = newly_diagnosed + newly_nondiagnosed,
      retention_rate = sum(current_diagnoses & baseline_results) / sum(baseline_results)
    )
  }
  
  # Initialize top combinations
  top_combinations <- list(
    first = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL),
    second = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL),
    third = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL)
  )
  
  # Get all possible combinations of 2 symptoms
  all_symptoms <- 1:20
  combinations <- combn(all_symptoms, 2, simplify = FALSE)
  
  # Test each combination
  for(combination in combinations) {
    symptom_names <- paste0("symptom_", combination)
    
    # Get diagnoses and calculate metrics
    current_diagnoses <- get_diagnosis(binarized_data, symptom_names)
    metrics <- get_metrics(current_diagnoses, baseline_results)
    
    # Score based on specified method
    score <- if(score_by == "false_cases") {
      -metrics$false_cases
    } else {
      -metrics$newly_nondiagnosed
    }
    
    # Update top combinations if needed
    if(score > top_combinations$first$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- top_combinations$first
      top_combinations$first <- list(
        combination = combination,
        score = score,
        diagnoses = current_diagnoses,
        metrics = metrics
      )
    } else if(score > top_combinations$second$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- list(
        combination = combination,
        score = score,
        diagnoses = current_diagnoses,
        metrics = metrics
      )
    } else if(score > top_combinations$third$score) {
      top_combinations$third <- list(
        combination = combination,
        score = score,
        diagnoses = current_diagnoses,
        metrics = metrics
      )
    }
  }
  
  # Create comparison dataframe
  make_column_name <- function(combination) {
    sprintf("symptom_%d_%d", combination[1], combination[2])
  }
  
  comparison_df <- data.frame(
    PTSD_all = baseline_results,
    top_combinations$first$diagnoses,
    top_combinations$second$diagnoses,
    top_combinations$third$diagnoses
  )
  names(comparison_df)[2:4] <- c(
    make_column_name(top_combinations$first$combination),
    make_column_name(top_combinations$second$combination),
    make_column_name(top_combinations$third$combination)
  )
  
  return(list(
    best_symptoms = top_combinations$first$combination,
    best_symptom_names = paste0("symptom_", top_combinations$first$combination),
    best_metrics = top_combinations$first$metrics,
    second_best_symptoms = top_combinations$second$combination,
    second_symptom_names = paste0("symptom_", top_combinations$second$combination),
    second_metrics = top_combinations$second$metrics,
    third_best_symptoms = top_combinations$third$combination,
    third_symptom_names = paste0("symptom_", top_combinations$third$combination),
    third_metrics = top_combinations$third$metrics,
    original_diagnoses = sum(baseline_results),
    diagnosis_comparison = comparison_df
  ))
}
```

Summary for three best two-symptom-combinations overall, regardless of which cluster they belong to. Best means fewest newly_nondiagnosed
```{r Summary for three best two-symptom-combinations overall, regardless of which cluster they belong to. Best means fewest newly_nondiagnosed, echo=FALSE, message=FALSE}
### Applying to out data
PTSD_Diagnosis_best_two_symptoms_overall_min_false_neg <- analyze_best_two_symptoms(simulated_ptsd_data_renamed, score_by = "newly_nondiagnosed")
# Summary
summary_PTSD_Diagnosis_best_two_symptoms_overall_min_false_neg <- summarize_ptsd_changes(PTSD_Diagnosis_best_two_symptoms_overall_min_false_neg$diagnosis_comparison)
# Readable summary
readable_summary_PTSD_Diagnosis_best_two_symptoms_overall_min_false_neg <- create_readable_summary(summary_PTSD_Diagnosis_best_two_symptoms_overall_min_false_neg)
datatable(readable_summary_PTSD_Diagnosis_best_two_symptoms_overall_min_false_neg,
          options = list(scrollX = TRUE))
```
Summary for three best two-symptom-combinations overall, regardless of which cluster they belong to. Best means fewest false_cases (newly_nondiagnosed + newly_diagnosed)
```{r Summary for three best two-symptom-combinations overall, regardless of which cluster they belong to. Best means fewest false_cases, echo=FALSE, message=FALSE}
### Applying to out data
PTSD_Diagnosis_best_two_symptoms_overall_min_false <- analyze_best_two_symptoms(simulated_ptsd_data_renamed, score_by = "false_cases")
# Summary
summary_PTSD_Diagnosis_best_two_symptoms_overall_min_false <- summarize_ptsd_changes(PTSD_Diagnosis_best_two_symptoms_overall_min_false$diagnosis_comparison)
# Readable summary
readable_summary_PTSD_Diagnosis_best_two_symptoms_overall_min_false <- create_readable_summary(summary_PTSD_Diagnosis_best_two_symptoms_overall_min_false)
# datatable(readable_summary_PTSD_Diagnosis_best_two_symptoms_overall_min_false,
#           options = list(scrollX = TRUE))
```

#### 5.3.2. Find three best two-symptom-combinations from two different clusters for all possible combinations of 2 clusters

```{r Find three best two-symptom-combinations from two different clusters for all possible combinations of 2 clusters, echo=FALSE, message=FALSE}
### Function for analyzing the three best two symptom-combinations from two different clusters for all possible combinations of 2 clusters
  # Depending on the score_by select the best three-symptom-combination by minimizing the newly_nondiagnosed or by minimizing the false_cases
  # Input: Data set with all patients and 20 columns with non-binarized symptoms
analyze_best_two_symptoms_by_cluster <- function(data,  score_by = "false_cases") {
  # Validate scoring method
  valid_scoring <- c("false_cases", "newly_nondiagnosed")
  if (!score_by %in% valid_scoring) {
    stop("score_by must be one of: ", paste(valid_scoring, collapse = ", "))
  }
  
  # Define clusters
  clusters <- list(
    cluster1 = 1:5,    # Criterion A
    cluster2 = 6:7,    # Criterion B
    cluster3 = 8:14,   # Criterion C
    cluster4 = 15:20   # Criterion D
  )
  
  # Helper functions
  binarize_data <- function(data) {
    data[data <= 1] <- 0
    data[data >= 2] <- 1
    return(data == 1)
  }
  
  # Get baseline results first
  baseline_results <- analyze_ptsd_all(data)$PTSD_all
  baseline_count <- sum(baseline_results)
  
  # Binarize data
  binarized_data <- binarize_data(data)
  
  get_diagnosis <- function(data, symptoms) {
    subset_data <- data[, symptoms]
    return(apply(subset_data, 1, all))
  }
  
  get_metrics <- function(current_diagnoses, baseline_results) {
    newly_diagnosed <- sum(current_diagnoses & !baseline_results)
    newly_nondiagnosed <- sum(!current_diagnoses & baseline_results)
    
    list(
      true_positives = sum(current_diagnoses & baseline_results),
      true_negatives = sum(!current_diagnoses & !baseline_results),
      newly_diagnosed = newly_diagnosed,
      newly_nondiagnosed = newly_nondiagnosed,
      false_cases = newly_diagnosed + newly_nondiagnosed,
      retention_rate = sum(current_diagnoses & baseline_results) / sum(baseline_results)
    )
  }
  
  # Helper function to update top 3 combinations
  update_top_combinations <- function(top_combinations, current_combination, score, diagnoses, metrics, used_clusters) {
    if(score > top_combinations$first$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- top_combinations$first
      top_combinations$first <- list(
        combination = current_combination,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics,
        used_clusters = used_clusters
      )
    } else if(score > top_combinations$second$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- list(
        combination = current_combination,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics,
        used_clusters = used_clusters
      )
    } else if(score > top_combinations$third$score) {
      top_combinations$third <- list(
        combination = current_combination,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics,
        used_clusters = used_clusters
      )
    }
    return(top_combinations)
  }
  
  # Initialize results list
  results_by_cluster_pair <- list()
  counter <- 1
  
  # For each pair of clusters
  cluster_combinations <- combn(1:4, 2, simplify = FALSE)
  
  for(cluster_pair in cluster_combinations) {
    # Get the two selected clusters
    cluster1 <- clusters[[cluster_pair[1]]]
    cluster2 <- clusters[[cluster_pair[2]]]
    
   # Initialize top 3 combinations for this cluster pair
    top_combinations <- list(
      first = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL, used_clusters = cluster_pair),
      second = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL, used_clusters = cluster_pair),
      third = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL, used_clusters = cluster_pair)
    )
    
    # Generate all possible combinations from these two clusters
    combinations <- expand.grid(cluster1, cluster2)
    
    # Test each combination
    for(i in 1:nrow(combinations)) {
      current_combination <- as.numeric(combinations[i,])
      symptom_names <- paste0("symptom_", current_combination)
      
      # Get diagnoses and calculate metrics
      current_diagnoses <- get_diagnosis(binarized_data, symptom_names)
      metrics <- get_metrics(current_diagnoses, baseline_results)
      
      # Score based on specified method
      score <- if(score_by == "false_cases") {
        -metrics$false_cases
      } else {
        -metrics$newly_nondiagnosed
      }
      
      # Update top combinations
      top_combinations <- update_top_combinations(
        top_combinations,
        current_combination,
        score,
        current_diagnoses,
        metrics,
        cluster_pair
      )
    }
    
    # Store results for this cluster pair
    results_by_cluster_pair[[counter]] <- top_combinations
    counter <- counter + 1
  }
  
  # Create comparison dataframe
  comparison_df <- data.frame(PTSD_all = baseline_results)
  # Add all combinations to comparison dataframe
  for(i in 1:length(cluster_combinations)) {
    # Add first combination
    col_name <- sprintf("clusters%d_%d_symptoms_%d_%d", 
                        results_by_cluster_pair[[i]]$first$used_clusters[1],
                        results_by_cluster_pair[[i]]$first$used_clusters[2],
                        results_by_cluster_pair[[i]]$first$combination[1],
                        results_by_cluster_pair[[i]]$first$combination[2])
    comparison_df[[col_name]] <- results_by_cluster_pair[[i]]$first$diagnoses
    
    # Add second combination
    col_name <- sprintf("clusters%d_%d_second_%d_%d", 
                       results_by_cluster_pair[[i]]$second$used_clusters[1],
                       results_by_cluster_pair[[i]]$second$used_clusters[2],
                       results_by_cluster_pair[[i]]$second$combination[1],
                       results_by_cluster_pair[[i]]$second$combination[2])
    comparison_df[[col_name]] <- results_by_cluster_pair[[i]]$second$diagnoses
    
    # Add third combination
    col_name <- sprintf("clusters%d_%d_third_%d_%d", 
                       results_by_cluster_pair[[i]]$third$used_clusters[1],
                       results_by_cluster_pair[[i]]$third$used_clusters[2],
                       results_by_cluster_pair[[i]]$third$combination[1],
                       results_by_cluster_pair[[i]]$third$combination[2])
    comparison_df[[col_name]] <- results_by_cluster_pair[[i]]$third$diagnoses
  }
  
  return(list(
    results_by_cluster_pair = results_by_cluster_pair,
    original_diagnoses = sum(baseline_results),
    diagnosis_comparison = comparison_df
  ))
}
```

Summary for three best two-symptom-combinations, from two different clusters for all possible combinations of 2 clusters. Best means fewest newly_nondiagnosed
```{r Summary for three best two-symptom-combinations, from two different clusters for all possible combinations of 2 clusters. Best means fewest newly_nondiagnosed, echo=FALSE, message=FALSE}
### Applying to our data
PTSD_Diagnosis_best_two_symptoms_by_cluster_min_false_neg <- analyze_best_two_symptoms_by_cluster(simulated_ptsd_data_renamed, score_by = "newly_nondiagnosed")
# Summary
summary_PTSD_Diagnosis_best_two_symptoms_by_cluster_min_false_neg <- summarize_ptsd_changes(PTSD_Diagnosis_best_two_symptoms_by_cluster_min_false_neg$diagnosis_comparison)
# Readable summary
readable_summary_PTSD_Diagnosis_best_two_symptoms_by_cluster_min_false_neg <- create_readable_summary(summary_PTSD_Diagnosis_best_two_symptoms_by_cluster_min_false_neg)
datatable(readable_summary_PTSD_Diagnosis_best_two_symptoms_by_cluster_min_false_neg,
          options = list(scrollX = TRUE))
```

Summary for three best two-symptom-combinations, from two different clusters for all possible combinations of 2 clusters. Best means fewest false_cases (newly_nondiagnosed + newly_diagnosed)
```{r Summary for three best two-symptom-combinations, from two different clusters for all possible combinations of 2 clusters. Best means fewest false_cases, echo=FALSE, message=FALSE}
### Applying to our data
PTSD_Diagnosis_best_two_symptoms_by_cluster_min_false <- analyze_best_two_symptoms_by_cluster(simulated_ptsd_data_renamed, score_by = "false_cases")
# Summary
summary_PTSD_Diagnosis_best_two_symptoms_by_cluster_min_false <- summarize_ptsd_changes(PTSD_Diagnosis_best_two_symptoms_by_cluster_min_false$diagnosis_comparison)
# Readable summary
readable_summary_PTSD_Diagnosis_best_two_symptoms_by_cluster_min_false <- create_readable_summary(summary_PTSD_Diagnosis_best_two_symptoms_by_cluster_min_false)
# datatable(readable_summary_PTSD_Diagnosis_best_two_symptoms_by_cluster_min_false,
#           options = list(scrollX = TRUE))
```

### 5.4. Identification of the most important symptom in each cluster

```{r Identification of the most important symptom in each cluster, echo=FALSE, message=FALSE}
### Function for analyzing the best symptom from each cluster
  # Depending on the score_by select the best three-symptom-combination by minimizing the newly_nondiagnosed or by minimizing the false_cases
  # Input: Data set with all patients and 20 columns with non-binarized symptoms
analyze_best_single_symptom_by_cluster <- function(data, score_by = "false_cases") {
  # Validate scoring method
  valid_scoring <- c("false_cases", "newly_nondiagnosed")
  if (!score_by %in% valid_scoring) {
    stop("score_by must be one of: ", paste(valid_scoring, collapse = ", "))
  }
  
  # Define clusters
  clusters <- list(
    cluster1 = 1:5,    
    cluster2 = 6:7,   
    cluster3 = 8:14,   
    cluster4 = 15:20   
  )
  
  # Helper functions
  binarize_data <- function(data) {
    data[data <= 1] <- 0
    data[data >= 2] <- 1
    return(data == 1)
  }
  
  # Get baseline results first
  baseline_results <- analyze_ptsd_all(data)$PTSD_all
  baseline_count <- sum(baseline_results)
  
  # Binarize data
  binarized_data <- binarize_data(data)
  
  get_diagnosis <- function(data, symptom_index) {
    column_name <- paste0("symptom_", symptom_index)
    return(data[,column_name])
  }
  
  get_metrics <- function(current_diagnoses, baseline_results) {
    newly_diagnosed <- sum(current_diagnoses & !baseline_results)
    newly_nondiagnosed <- sum(!current_diagnoses & baseline_results)
    
    list(
      true_positives = sum(current_diagnoses & baseline_results),
      true_negatives = sum(!current_diagnoses & !baseline_results),
      newly_diagnosed = newly_diagnosed,
      newly_nondiagnosed = newly_nondiagnosed,
      false_cases = newly_diagnosed + newly_nondiagnosed,
      retention_rate = sum(current_diagnoses & baseline_results) / sum(baseline_results)
    )
  }
  
  # Helper function to update top 3 symptoms
  update_top_symptoms <- function(top_symptoms, symptom, score, diagnoses, metrics, cluster) {
    if(score > top_symptoms$first$score) {
      top_symptoms$third <- top_symptoms$second
      top_symptoms$second <- top_symptoms$first
      top_symptoms$first <- list(
        symptom = symptom,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics,
        cluster = cluster
      )
    } else if(score > top_symptoms$second$score) {
      top_symptoms$third <- top_symptoms$second
      top_symptoms$second <- list(
        symptom = symptom,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics,
        cluster = cluster
      )
    } else if(score > top_symptoms$third$score) {
      top_symptoms$third <- list(
        symptom = symptom,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics,
        cluster = cluster
      )
    }
    return(top_symptoms)
  }
  
  # Initialize results list
  results_by_cluster <- list()
  
  # For each cluster
  for(cluster_num in 1:4) {
    # Get current cluster symptoms
    current_cluster <- clusters[[cluster_num]]
    
    # Initialize top 3 symptoms for this cluster
    top_symptoms <- list(
      first = list(symptom = NULL, score = -Inf, diagnoses = NULL, metrics = NULL, cluster = cluster_num),
      second = list(symptom = NULL, score = -Inf, diagnoses = NULL, metrics = NULL, cluster = cluster_num),
      third = list(symptom = NULL, score = -Inf, diagnoses = NULL, metrics = NULL, cluster = cluster_num)
    )
    
    # Test each symptom in the cluster
    for(symptom in current_cluster) {
      # Get diagnoses and calculate metrics
      current_diagnoses <- get_diagnosis(binarized_data, symptom)
      metrics <- get_metrics(current_diagnoses, baseline_results)
      
      # Score based on chosen criterion
      score <- if(score_by == "false_cases") {
        -metrics$false_cases
      } else {
        -metrics$newly_nondiagnosed
      }
      
      # Update top symptoms
      top_symptoms <- update_top_symptoms(
        top_symptoms,
        symptom,
        score,
        current_diagnoses,
        metrics,
        cluster_num
      )
    }
    
    # Store results for this cluster
    results_by_cluster[[cluster_num]] <- top_symptoms
  }
  
  # Create comparison dataframe
  comparison_df <- data.frame(PTSD_all = baseline_results)
  # Add all symptoms to comparison dataframe
  for(i in 1:4) {
    # Add first symptom
    if(!is.null(results_by_cluster[[i]]$first$symptom)) {
    col_name <- sprintf("cluster%d_first_symptom%d", 
                        i,
                        results_by_cluster[[i]]$first$symptom)
    comparison_df[[col_name]] <- results_by_cluster[[i]]$first$diagnoses
    }
    
    # Add second symptom if exists and cluster has more than 1 symptom
    if(!is.null(results_by_cluster[[i]]$second$symptom) && length(clusters[[i]]) > 1) {
    col_name <- sprintf("cluster%d_second_symptom%d", 
                       i,
                       results_by_cluster[[i]]$second$symptom)
    comparison_df[[col_name]] <- results_by_cluster[[i]]$second$diagnoses
    }
    
    # Add third symptom if exists and cluster has more than 2 symptoms
    if(!is.null(results_by_cluster[[i]]$third$symptom) && length(clusters[[i]]) > 2) {
    col_name <- sprintf("cluster%d_third_symptom%d", 
                       i,
                       results_by_cluster[[i]]$third$symptom)
    comparison_df[[col_name]] <- results_by_cluster[[i]]$third$diagnoses
    }
  }
  
  return(list(
    results_by_cluster = results_by_cluster,
    original_diagnoses = sum(baseline_results),
    diagnosis_comparison = comparison_df
  ))
}
```

Summary for three best symptoms in each cluster. Best means fewest newly_nondiagnosed
```{r Summary for three best symptoms in each cluster. Best means fewest newly_nondiagnosed, echo=FALSE, message=FALSE}
### Applying to our data
PTSD_Diagnosis_best_single_symptom_by_cluster_min_false_neg <- analyze_best_single_symptom_by_cluster(simulated_ptsd_data_renamed, score_by = "newly_nondiagnosed")
# Summary
summary_PTSD_Diagnosis_best_single_symptom_by_cluster_min_false_neg <- summarize_ptsd_changes(PTSD_Diagnosis_best_single_symptom_by_cluster_min_false_neg$diagnosis_comparison)
# Readable summary
readable_summary_PTSD_Diagnosis_best_single_symptom_by_cluster_min_false_neg <- create_readable_summary(summary_PTSD_Diagnosis_best_single_symptom_by_cluster_min_false_neg)
datatable(readable_summary_PTSD_Diagnosis_best_single_symptom_by_cluster_min_false_neg,
          options = list(scrollX = TRUE))
```

Summary for three best symptoms in each cluster. Best means fewest false_cases (newly_nondiagnosed + newly_diagnosed)
```{r Summary for three best symptoms in each cluster. Best means fewest false_cases, echo=FALSE, message=FALSE}
### Applying to our data
PTSD_Diagnosis_best_single_symptom_by_cluster_min_false <- analyze_best_single_symptom_by_cluster(simulated_ptsd_data_renamed, score_by = "false_cases")
# Summary
summary_PTSD_Diagnosis_best_single_symptom_by_cluster_min_false <- summarize_ptsd_changes(PTSD_Diagnosis_best_single_symptom_by_cluster_min_false$diagnosis_comparison)
# Readable summary
readable_summary_PTSD_Diagnosis_best_single_symptom_by_cluster_min_false <- create_readable_summary(summary_PTSD_Diagnosis_best_single_symptom_by_cluster_min_false)
# datatable(readable_summary_PTSD_Diagnosis_best_single_symptom_by_cluster_min_false,
#           options = list(scrollX = TRUE))
```

### 5.5. Identification of the 6 most important symptoms following original diagnostic criteria

#### 5.5.1. Find three best 6-symptom-combinations overall, regardless of which cluster they belong to.

```{r Find three best 6-symptom-combinations overall, regardless of which cluster they belong to, echo=FALSE, message=FALSE}
### Function to determine the three best 6-symptom-combinations overall, regardless of which cluster they belong to.
  # Depending on the score_by select the best 6-symptom-combinations by minimizing the newly_nondiagnosed or by minimizing the false_cases
  # Input: data set with all patients and 20 columns with non-binarized symptoms
analyze_best_six_symptoms <- function(data, score_by = "false_cases") {
  # Validate scoring method
  valid_scoring <- c("false_cases", "newly_nondiagnosed")
  if (!score_by %in% valid_scoring) {
    stop("score_by must be one of: ", paste(valid_scoring, collapse = ", "))
  }
  
  # Helper functions
  binarize_data <- function(data) {
    data[data <= 1] <- 0
    data[data >= 2] <- 1
    return(data == 1) # Convert to logical
  }
  
  # Get baseline results first
  baseline_results <- analyze_ptsd_all(data)$PTSD_all
  baseline_count <- sum(baseline_results)
  
  # Binarize data
  binarized_data <- binarize_data(data)
  
  # Helper function to get diagnosis based on having all specified symptoms
  get_diagnosis <- function(data, symptoms) {
    subset_data <- data[, symptoms]
    return(apply(subset_data, 1, all))
  }
  
  # Helper function to calculate metrics
  get_metrics <- function(current_diagnoses, baseline_results) {
    newly_diagnosed <- sum(current_diagnoses & !baseline_results)
    newly_nondiagnosed <- sum(!current_diagnoses & baseline_results)
    
    list(
      true_positives = sum(current_diagnoses & baseline_results),
      true_negatives = sum(!current_diagnoses & !baseline_results),
      newly_diagnosed = newly_diagnosed,
      newly_nondiagnosed = newly_nondiagnosed,
      false_cases = newly_diagnosed + newly_nondiagnosed,
      retention_rate = sum(current_diagnoses & baseline_results) / sum(baseline_results)
    )
  }
  
  # Initialize top combinations
  top_combinations <- list(
    first = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL),
    second = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL),
    third = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL)
  )
  
  # Get all possible combinations of 6 symptoms
  all_symptoms <- 1:20
  combinations <- combn(all_symptoms, 6, simplify = FALSE)
  
  # Test each combination
  for(combination in combinations) {
    symptom_names <- paste0("symptom_", combination)
    
    # Get diagnoses and calculate metrics
    current_diagnoses <- get_diagnosis(binarized_data, symptom_names)
    metrics <- get_metrics(current_diagnoses, baseline_results)
    
    # Score based on specified method
    score <- if(score_by == "false_cases") {
      -metrics$false_cases
    } else {
      -metrics$newly_nondiagnosed
    }
    
    # Update top combinations if needed
    if(score > top_combinations$first$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- top_combinations$first
      top_combinations$first <- list(
        combination = combination,
        score = score,
        diagnoses = current_diagnoses,
        metrics = metrics
      )
    } else if(score > top_combinations$second$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- list(
        combination = combination,
        score = score,
        diagnoses = current_diagnoses,
        metrics = metrics
      )
    } else if(score > top_combinations$third$score) {
      top_combinations$third <- list(
        combination = combination,
        score = score,
        diagnoses = current_diagnoses,
        metrics = metrics
      )
    }
  }
  
  # Create comparison dataframe
  make_column_name <- function(combination) {
    sprintf("symptom_%d_%d_%d_%d_%d_%d", 
            combination[1], combination[2], combination[3],
            combination[4], combination[5], combination[6])
  }
  
  comparison_df <- data.frame(
    PTSD_all = baseline_results,
    top_combinations$first$diagnoses,
    top_combinations$second$diagnoses,
    top_combinations$third$diagnoses
  )
  names(comparison_df)[2:4] <- c(
    make_column_name(top_combinations$first$combination),
    make_column_name(top_combinations$second$combination),
    make_column_name(top_combinations$third$combination)
  )
  
  return(list(
    best_symptoms = top_combinations$first$combination,
    best_symptom_names = paste0("symptom_", top_combinations$first$combination),
    best_metrics = top_combinations$first$metrics,
    second_best_symptoms = top_combinations$second$combination,
    second_symptom_names = paste0("symptom_", top_combinations$second$combination),
    second_metrics = top_combinations$second$metrics,
    third_best_symptoms = top_combinations$third$combination,
    third_symptom_names = paste0("symptom_", top_combinations$third$combination),
    third_metrics = top_combinations$third$metrics,
    original_diagnoses = sum(baseline_results),
    diagnosis_comparison = comparison_df
  ))
}
```
Summary for top three 6-symptom-combinations overall, regardless of which cluster they belong to. Best means fewest newly_nondiagnosed
```{r Summary for top three 6-symptom-combinations overall, regardless of which cluster they belong to. Best means fewest newly_nondiagnosed, echo=FALSE, message=FALSE}
### Applying to our data
PTSD_Diagnosis_best_six_symptoms_overall_min_false_neg <- analyze_best_six_symptoms(simulated_ptsd_data_renamed, score_by = "newly_nondiagnosed")
# Summary
summary_PTSD_Diagnosis_best_six_symptoms_overall_min_false_neg <- summarize_ptsd_changes(PTSD_Diagnosis_best_six_symptoms_overall_min_false_neg$diagnosis_comparison)
# Readable summary
readable_summary_PTSD_Diagnosis_best_six_symptoms_overall_min_false_neg <- create_readable_summary(summary_PTSD_Diagnosis_best_six_symptoms_overall_min_false_neg)
datatable(readable_summary_PTSD_Diagnosis_best_six_symptoms_overall_min_false_neg,
          options = list(scrollX = TRUE))
```

Summary for top three six-symptom-combinations overall, regardless of which cluster they belong to. Best means fewest false_cases (newly_nondiagnosed + newly_diagnosed)
```{r Summary for top three six-symptom-combinations overall, regardless of which cluster they belong to. Best means fewest false_cases, echo=FALSE, message=FALSE}
### Applying to our data
PTSD_Diagnosis_best_six_symptoms_overall_min_false <- analyze_best_six_symptoms(simulated_ptsd_data_renamed, score_by = "false_cases")
# Summary
summary_PTSD_Diagnosis_best_six_symptoms_overall_min_false <- summarize_ptsd_changes(PTSD_Diagnosis_best_six_symptoms_overall_min_false$diagnosis_comparison)
# Readable summary
readable_summary_PTSD_Diagnosis_best_six_symptoms_overall_min_false <- create_readable_summary(summary_PTSD_Diagnosis_best_six_symptoms_overall_min_false)
# datatable(readable_summary_PTSD_Diagnosis_best_four_symptoms_overall_min_false,
#           options = list(scrollX = TRUE))
```

#### 5.5.2. Find three best 6-symptom-combinations following original diagnostic criteria

##### 5.5.2.1. Variant 1: Identification of the best three six-symptom-combinations following original diagnostic criteria by analyzing individual importance, selecting top candidates and testing different combinations

```{r Variant 1: Identification of the best three six-symptom-combinations following original diagnostic criteria, echo=FALSE, message=FALSE}
### Function to analyze best three best six-symptom combinations following original diagnostic criteria by 1. analyzing individual importance of symptoms, 2. selecting most important symptoms from each cluster as candidates, 3. testing different combinations of these candidates to chose best
  # Depending on the score_by select the best three-symptom-combination by minimizing the newly_nondiagnosed or by minimizing the false_cases
  # Input: Data set with all patients and 20 columns with non-binarized symptoms
variant1_analyze_best_six_symptoms_diagnostic_criteria <- function(data, score_by = "false_cases") {
  # Validate scoring method
  valid_scoring <- c("false_cases", "newly_nondiagnosed")
  if (!score_by %in% valid_scoring) {
    stop("score_by must be one of: ", paste(valid_scoring, collapse = ", "))
  }
  
  # Define clusters
  clusters <- list(
    cluster1 = 1:5,    # Criterion A - need 1
    cluster2 = 6:7,    # Criterion B - need 1
    cluster3 = 8:14,   # Criterion C - need 2
    cluster4 = 15:20   # Criterion D - need 2
  )
  
  # Helper functions
  binarize_data <- function(data) {
    data[data <= 1] <- 0
    data[data >= 2] <- 1
    return(data == 1)
  }
  
  # Get baseline results first
  baseline_results <- analyze_ptsd_all(data)$PTSD_all
  baseline_count <- sum(baseline_results)
  
  # Binarize data
  binarized_data <- binarize_data(data)
  
  # Helper function to get diagnosis based on having required symptoms from each cluster
  get_diagnosis <- function(data, symptoms_cluster1, symptoms_cluster2, 
                          symptoms_cluster3, symptoms_cluster4) {
    # Check criteria for each cluster
    cluster1_met <- data[, paste0("symptom_", symptoms_cluster1)]
    cluster2_met <- data[, paste0("symptom_", symptoms_cluster2)]
    cluster3_met <- rowSums(data[, paste0("symptom_", symptoms_cluster3)]) >= 2
    cluster4_met <- rowSums(data[, paste0("symptom_", symptoms_cluster4)]) >= 2
    
    # All criteria must be met
    return(cluster1_met & cluster2_met & cluster3_met & cluster4_met)
  }
  
  get_metrics <- function(current_diagnoses, baseline_results) {
    newly_diagnosed <- sum(current_diagnoses & !baseline_results)
    newly_nondiagnosed <- sum(!current_diagnoses & baseline_results)
    
    list(
      true_positives = sum(current_diagnoses & baseline_results),
      true_negatives = sum(!current_diagnoses & !baseline_results),
      newly_diagnosed = newly_diagnosed,
      newly_nondiagnosed = newly_nondiagnosed,
      false_cases = newly_diagnosed + newly_nondiagnosed,
      retention_rate = sum(current_diagnoses & baseline_results) / sum(baseline_results)
    )
  }
  
  # Initialize top 3 combinations
  top_combinations <- list(
    first = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL),
    second = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL),
    third = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL)
  )
  
  # Helper function to update top 3 combinations
  update_top_combinations <- function(top_combinations, current_combination, score, diagnoses, metrics) {
    if(score > top_combinations$first$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- top_combinations$first
      top_combinations$first <- list(
        combination = current_combination,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics
      )
    } else if(score > top_combinations$second$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- list(
        combination = current_combination,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics
      )
    } else if(score > top_combinations$third$score) {
      top_combinations$third <- list(
        combination = current_combination,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics
      )
    }
    return(top_combinations)
  }
  
  # Pre-select top candidates from each cluster based on individual performance
  get_top_candidates <- function(data, cluster_indices, n_candidates) { # cluster_indices z.B. 1:5, n_candidates how many top candidates are chosen
     # Subset data to only include rows where PTSD_all is TRUE
    diagnosed_data <- data[baseline_results, ]
    
    # Calculate frequencies only for diagnosed cases
    cluster_data <- diagnosed_data[, paste0("symptom_", cluster_indices)]
    symptom_freq <- colSums(cluster_data)
    
    # Get top n most frequent symptoms
    top_indices <- order(symptom_freq, decreasing=TRUE)[1:n_candidates]
    return(cluster_indices[top_indices])
}
  
  # Get top candidates from each cluster
  top_cluster1 <- get_top_candidates(binarized_data, clusters$cluster1, 3)  # Top 3 from cluster 1
  top_cluster2 <- get_top_candidates(binarized_data, clusters$cluster2, 2)  # All from cluster 2
  top_cluster3 <- get_top_candidates(binarized_data, clusters$cluster3, 4)  # Top 4 from cluster 3
  top_cluster4 <- get_top_candidates(binarized_data, clusters$cluster4, 4)  # Top 4 from cluster 4
  
  # Test combinations of top candidates
  for(symptom1 in top_cluster1) {
    for(symptom2 in top_cluster2) {
      # Get combinations of 2 symptoms from cluster 3
      cluster3_combinations <- combn(top_cluster3, 2, simplify = FALSE)
      for(symptoms3 in cluster3_combinations) {
        # Get combinations of 2 symptoms from cluster 4
        cluster4_combinations <- combn(top_cluster4, 2, simplify = FALSE)
        for(symptoms4 in cluster4_combinations) {
          # Get diagnoses for this combination
          current_diagnoses <- get_diagnosis(
            binarized_data, 
            symptom1,
            symptom2,
            symptoms3,
            symptoms4
          )
          
          # Calculate metrics
          metrics <- get_metrics(current_diagnoses, baseline_results)
          
          # Score based on specified method
          score <- if(score_by == "false_cases") {
            -metrics$false_cases
          } else {
            -metrics$newly_nondiagnosed
          }
          
          # Update top combinations instead of just best
          current_combination <- c(symptom1, symptom2, symptoms3, symptoms4)
          top_combinations <- update_top_combinations(
            top_combinations,
            current_combination,
            score,
            current_diagnoses,
            metrics
          )
        }
      }
    }
  }
  
  # Create comparison dataframe
  comparison_df <- data.frame(PTSD_all = baseline_results)
  
   # Add each of the top 3 combinations
  for(rank in c("first", "second", "third")) {
    col_name <- sprintf(
      "%s_symptoms_%d_%d_%d_%d_%d_%d",
      rank,
      top_combinations[[rank]]$combination[1],
      top_combinations[[rank]]$combination[2],
      top_combinations[[rank]]$combination[3],
      top_combinations[[rank]]$combination[4],
      top_combinations[[rank]]$combination[5],
      top_combinations[[rank]]$combination[6]
    )
    comparison_df[[col_name]] <- top_combinations[[rank]]$diagnoses
  }
  
  return(list(
    best_symptoms = top_combinations$first$combination,
    best_symptom_names = paste0("symptom_", top_combinations$first$combination),
    best_metrics = top_combinations$first$metrics,
    second_best_symptoms = top_combinations$second$combination,
    second_symptom_names = paste0("symptom_", top_combinations$second$combination),
    second_metrics = top_combinations$second$metrics,
    third_best_symptoms = top_combinations$third$combination,
    third_symptom_names = paste0("symptom_", top_combinations$third$combination),
    third_metrics = top_combinations$third$metrics,
    original_diagnoses = sum(baseline_results),
    diagnosis_comparison = comparison_df
  ))
}
```

Variant1: Summary for three best six-symptom-combinations following original diagnostic criteria. Best means fewest newly_nondiagnosed
```{r Variant1: Summary for three best six-symptom-combinations following original diagnostic criteria. Best means fewest newly_nondiagnosed, echo=FALSE, message=FALSE}
### Applying to our data
variant1_PTSD_Diagnosis_best_six_symptoms_min_false_neg <- variant1_analyze_best_six_symptoms_diagnostic_criteria(simulated_ptsd_data_renamed, score_by = "newly_nondiagnosed")
# Summary
summary_variant1_PTSD_Diagnosis_best_six_symptoms_min_false_neg <- summarize_ptsd_changes(variant1_PTSD_Diagnosis_best_six_symptoms_min_false_neg$diagnosis_comparison)
# Readable summary
readable_summary_variant1_PTSD_Diagnosis_best_six_symptoms_min_false_neg <- create_readable_summary(summary_variant1_PTSD_Diagnosis_best_six_symptoms_min_false_neg)
datatable(readable_summary_variant1_PTSD_Diagnosis_best_six_symptoms_min_false_neg,
          options = list(scrollX = TRUE))
```
Variant1: Summary for three best six-symptom-combinations following original diagnostic criteria. Best means fewest false_cases (newly_nondiagnosed + newly_diagnosed)
```{r Variant1: Summary for three best six-symptom-combinations following original diagnostic criteria. Best means fewest false_cases, echo=FALSE, message=FALSE}
### Applying to our data
variant1_PTSD_Diagnosis_best_six_symptoms_min_false <- variant1_analyze_best_six_symptoms_diagnostic_criteria(simulated_ptsd_data_renamed, score_by = "false_cases")
# Summary
summary_variant1_PTSD_Diagnosis_best_six_symptoms_min_false <- summarize_ptsd_changes(variant1_PTSD_Diagnosis_best_six_symptoms_min_false$diagnosis_comparison)
# Readable summary
readable_summary_variant1_PTSD_Diagnosis_best_six_symptoms_min_false <- create_readable_summary(summary_variant1_PTSD_Diagnosis_best_six_symptoms_min_false)
# datatable(readable_summary_variant1_PTSD_Diagnosis_best_six_symptoms_min_false,
#           options = list(scrollX = TRUE))
```

#### 5.5.2.2. Variant 2: Identification of the 6 most important symptoms following original diagnostic criteria by generating all possible combinations

```{r Variant 2: Identification of the 6 most important symptoms following original diagnostic criteria by generating all possible combinations, echo=FALSE, message=FALSE}
### Function to analyze all combinations of 6 symptoms following original diagnostic criteria by generating all possible combinations, output: for each combination and each individual true/false of the PTSD diagnosis
  # Input: Data set with all patients and 20 columns with non-binarized symptoms
variant2_analyze_best_six_symptoms_diagnostic_criteria <- function(data) {
  check_ptsd_criteria_six <- function(symptoms, selected_indices) {
    full_symptoms <- rep(0, 20)
    full_symptoms[selected_indices] <- symptoms
    criterion_1 <- any(full_symptoms[1:5] == 1)
    criterion_2 <- any(full_symptoms[6:7] == 1)
    criterion_3 <- sum(full_symptoms[8:14] == 1) >= 2
    criterion_4 <- sum(full_symptoms[15:20] == 1) >= 2
    return(criterion_1 & criterion_2 & criterion_3 & criterion_4)
  }
  
  results <- analyze_ptsd_all(data)
  valid_combinations <- list()
  
  for(a in 1:5) {
    for(b in 6:7) {
      c_symptoms <- combn(8:14, 2)
      d_symptoms <- combn(15:20, 2)
      
      for(i in 1:ncol(c_symptoms)) {
        for(j in 1:ncol(d_symptoms)) {
          
          combination <- c(a, b, c_symptoms[,i], d_symptoms[,j])
          valid_combinations[[length(valid_combinations) + 1]] <- combination
        }
      }
    }
  }
  
  for(i in seq_along(valid_combinations)) {
    selected_symptoms <- valid_combinations[[i]]
    selected_data <- data[, selected_symptoms]
    selected_data[selected_data <= 1] <- 0
    selected_data[selected_data >= 2] <- 1
    ptsd_results <- apply(selected_data, 1, check_ptsd_criteria_six, selected_indices = selected_symptoms)
    col_name <- paste0("PTSD_", paste(selected_symptoms, collapse = "_"))
    results[[col_name]] <- ptsd_results
  }
  
  return(results)
}
```

Variant 2: Summary for all six-symptom-combinations sorted by newly non-diagnosed identified with variant 2 (10 six-symptom-combinations with fewest lost diagnoses)
```{r Variant 2: Summary for all six-symptom-combinations sorted by newly non-diagnosed, echo=FALSE, message=FALSE}
### Applying to our data
variant2_PTSD_Diagnosis_best_six_symptoms <- variant2_analyze_best_six_symptoms_diagnostic_criteria(simulated_ptsd_data_renamed)
  # Summary
summary_variant2_PTSD_Diagnosis_best_six_symptoms <- summarize_ptsd_changes(variant2_PTSD_Diagnosis_best_six_symptoms)
  # Readable Summary
readable_summary_variant2_PTSD_Diagnosis_best_six_symptoms <- create_readable_summary(summary_variant2_PTSD_Diagnosis_best_six_symptoms) 
 # Sort by newly non-diagnosed 
summary_variant2_PTSD_Diagnosis_best_six_symptoms_min_false_neg <- summary_variant2_PTSD_Diagnosis_best_six_symptoms %>%
  arrange(newly_nondiagnosed)
readable_summary_variant2_PTSD_Diagnosis_best_six_symptoms_min_false_neg <- create_readable_summary(summary_variant2_PTSD_Diagnosis_best_six_symptoms_min_false_neg)
datatable(head(readable_summary_variant2_PTSD_Diagnosis_best_six_symptoms_min_false_neg, 11),
          options = list(scrollX = TRUE))
```

Variant 2: Summary for all six-symptom-combinations sorted by false_cases (newly_nondiagnosed + newly_diagnosed) identified with variant 2 (10 six-symptom-combinations with fewest false_cases)
```{r Variant 2: Summary for all six-symptom-combinations sorted by false_cases, echo=FALSE, message=FALSE}
  # Sort by newly false cases (newly diagnosed + newly nondiagnosed)
summary_variant2_PTSD_Diagnosis_best_six_symptoms_min_false <- summary_variant2_PTSD_Diagnosis_best_six_symptoms %>% 
  arrange(false_cases)
readable_summary_variant2_PTSD_Diagnosis_best_six_symptoms_min_false <- create_readable_summary(summary_variant2_PTSD_Diagnosis_best_six_symptoms_min_false)
# datatable(head(readable_summary_variant2_PTSD_Diagnosis_best_six_symptoms_min_false, 11),
#           options = list(scrollX = TRUE))
```

#### 5.5.3. Find three best 6-symptom-combinations, with at least one symptom from each cluster

```{r Find three best 6-symptom-combinations, with at least one symptom from each cluster, echo=FALSE, message=FALSE}
### Function to determine the three best 6-symptom-combinations, with at least one symptom from each cluster
  # Depending on the score_by select the best 6-symptom-combinations by minimizing the newly_nondiagnosed or by minimizing the false_cases
  # Input: data set with all patients and 20 columns with non-binarized symptoms
### Function to analyze six-symptom combinations where at least one symptom comes from each cluster and remaining symptoms can be from any cluster
analyze_best_six_symptoms_by_cluster <- function(data, score_by = "false_cases") {
  # Validate scoring method
  valid_scoring <- c("false_cases", "newly_nondiagnosed")
  if (!score_by %in% valid_scoring) {
    stop("score_by must be one of: ", paste(valid_scoring, collapse = ", "))
  }
  
  # Define clusters
  clusters <- list(
    cluster1 = 1:5,    # Criterion A
    cluster2 = 6:7,    # Criterion B
    cluster3 = 8:14,   # Criterion C
    cluster4 = 15:20   # Criterion D
  )
  
  # Helper functions
  binarize_data <- function(data) {
    data[data <= 1] <- 0
    data[data >= 2] <- 1
    return(data == 1)
  }
  
  # Get baseline results first
  baseline_results <- analyze_ptsd_all(data)$PTSD_all
  baseline_count <- sum(baseline_results)
  
  # Binarize data
  binarized_data <- binarize_data(data)
  
  # Helper function to get diagnosis based on having all specified symptoms
  get_diagnosis <- function(data, symptoms) {
    subset_data <- data[, paste0("symptom_", symptoms)]
    return(apply(subset_data, 1, all))
  }
  
  # Helper function to calculate metrics
  get_metrics <- function(current_diagnoses, baseline_results) {
    newly_diagnosed <- sum(current_diagnoses & !baseline_results)
    newly_nondiagnosed <- sum(!current_diagnoses & baseline_results)
    
    list(
      true_positives = sum(current_diagnoses & baseline_results),
      true_negatives = sum(!current_diagnoses & !baseline_results),
      newly_diagnosed = newly_diagnosed,
      newly_nondiagnosed = newly_nondiagnosed,
      false_cases = newly_diagnosed + newly_nondiagnosed,
      retention_rate = sum(current_diagnoses & baseline_results) / sum(baseline_results)
    )
  }
  
  # Initialize top combinations
  top_combinations <- list(
    first = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL),
    second = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL),
    third = list(combination = NULL, score = -Inf, diagnoses = NULL, metrics = NULL)
  )
  
  # Helper function to update top 3 combinations
  update_top_combinations <- function(top_combinations, current_combination, score, diagnoses, metrics) {
    if(score > top_combinations$first$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- top_combinations$first
      top_combinations$first <- list(
        combination = current_combination,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics
      )
    } else if(score > top_combinations$second$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- list(
        combination = current_combination,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics
      )
    } else if(score > top_combinations$third$score) {
      top_combinations$third <- list(
        combination = current_combination,
        score = score,
        diagnoses = diagnoses,
        metrics = metrics
      )
    }
    return(top_combinations)
  }
  
  # Pre-select most promising symptoms from each cluster based on individual performance
  get_top_candidates <- function(data, cluster_indices, n_candidates) {
    diagnosed_data <- data[baseline_results, ]
    cluster_data <- diagnosed_data[, paste0("symptom_", cluster_indices)]
    symptom_freq <- colSums(cluster_data)
    top_indices <- order(symptom_freq, decreasing=TRUE)[1:min(n_candidates, length(cluster_indices))]
    return(cluster_indices[top_indices])
  }
  
  # Get top candidates from each cluster
  top_cluster1 <- get_top_candidates(binarized_data, clusters$cluster1, 4)
  top_cluster2 <- get_top_candidates(binarized_data, clusters$cluster2, 2)
  top_cluster3 <- get_top_candidates(binarized_data, clusters$cluster3, 4)
  top_cluster4 <- get_top_candidates(binarized_data, clusters$cluster4, 4)
  
  # Combine all symptoms for selecting the additional ones
  all_symptoms <- 1:20
  
  # For each combination of one symptom from each cluster
  for(symptom1 in top_cluster1) {
    for(symptom2 in top_cluster2) {
      for(symptom3 in top_cluster3) {
        for(symptom4 in top_cluster4) {
          # Get base combination with one from each cluster
          base_combination <- c(symptom1, symptom2, symptom3, symptom4)
          
          # Get remaining symptoms to choose from
          remaining_symptoms <- setdiff(all_symptoms, base_combination)
          
          # Choose 2 more symptoms from remaining ones
          remaining_combinations <- combn(remaining_symptoms, 2, simplify = FALSE)
          
          for(extra_symptoms in remaining_combinations) {
            current_combination <- c(base_combination, extra_symptoms)
            
            # Get diagnoses for this combination
            current_diagnoses <- get_diagnosis(binarized_data, current_combination)
            
            # Calculate metrics
            metrics <- get_metrics(current_diagnoses, baseline_results)
            
            # Score based on specified method
            score <- if(score_by == "false_cases") {
              -metrics$false_cases
            } else {
              -metrics$newly_nondiagnosed
            }
            
            # Update top combinations
            top_combinations <- update_top_combinations(
              top_combinations,
              current_combination,
              score,
              current_diagnoses,
              metrics
            )
          }
        }
      }
    }
  }
  
  # Create comparison dataframe
  comparison_df <- data.frame(PTSD_all = baseline_results)
  
  # Add each of the top 3 combinations
  for(rank in c("first", "second", "third")) {
    col_name <- sprintf(
      "%s_symptoms_%s",
      rank,
      paste(top_combinations[[rank]]$combination, collapse = "_")
    )
    comparison_df[[col_name]] <- top_combinations[[rank]]$diagnoses
  }
  
  return(list(
    best_symptoms = top_combinations$first$combination,
    best_symptom_names = paste0("symptom_", top_combinations$first$combination),
    best_metrics = top_combinations$first$metrics,
    second_best_symptoms = top_combinations$second$combination,
    second_symptom_names = paste0("symptom_", top_combinations$second$combination),
    second_metrics = top_combinations$second$metrics,
    third_best_symptoms = top_combinations$third$combination,
    third_symptom_names = paste0("symptom_", top_combinations$third$combination),
    third_metrics = top_combinations$third$metrics,
    original_diagnoses = sum(baseline_results),
    diagnosis_comparison = comparison_df
  ))
}
```

Summary for three best six-symptom-combinations with at least one from each cluster. Best means fewest newly_nondiagnosed
```{r Summary for three best six-symptom-combinations with at least one from each cluster. Best means fewest newly_nondiagnosed, echo=FALSE, message=FALSE}
### Applying to our data
PTSD_Diagnosis_best_six_symptoms_by_cluster_min_false_neg <- analyze_best_six_symptoms_by_cluster(simulated_ptsd_data_renamed, score_by = "newly_nondiagnosed")
# Summary
summary_PTSD_Diagnosis_best_six_symptoms_by_cluster_min_false_neg <- summarize_ptsd_changes(PTSD_Diagnosis_best_six_symptoms_by_cluster_min_false_neg$diagnosis_comparison)
# Readable summary
readable_summary_PTSD_Diagnosis_best_six_symptoms_by_cluster_min_false_neg <- create_readable_summary(summary_PTSD_Diagnosis_best_six_symptoms_by_cluster_min_false_neg)
datatable(readable_summary_PTSD_Diagnosis_best_six_symptoms_by_cluster_min_false_neg,
          options = list(scrollX = TRUE))
```

Summary for three best six-symptom-combinations with at least one from each cluster. Best means fewest false_cases
```{r Summary for three best six-symptom-combinations with at least one from each cluster. Best means fewest false_cases, echo=FALSE, message=FALSE}
### Applying to our data
PTSD_Diagnosis_best_six_symptoms_by_cluster_min_false <- analyze_best_six_symptoms_by_cluster(simulated_ptsd_data_renamed, score_by = "false_cases")
# Summary
summary_PTSD_Diagnosis_best_six_symptoms_by_cluster_min_false <- summarize_ptsd_changes(PTSD_Diagnosis_best_six_symptoms_by_cluster_min_false$diagnosis_comparison)
# Readable summary
readable_summary_PTSD_Diagnosis_best_six_symptoms_by_cluster_min_false <- create_readable_summary(summary_PTSD_Diagnosis_best_six_symptoms_by_cluster_min_false)
# datatable(readable_summary_PTSD_Diagnosis_best_six_symptoms_by_cluster_min_false,
#           options = list(scrollX = TRUE))
```

## 6. Analysis of the possible combinations of binarized symptoms that fulfill the diagnosis

### 6.1. Theoretically possible combinations of binarized symptoms fulfilling diagnosis and their frequency in real-world data

```{r Theoretically possible combinations of binarized symptoms fulfilling diagnosis and their frequency in real-world data, echo=FALSE, message=FALSE}
### Create function
  # Input-data: real-world dataset fulfilling diagnostic criteria, 20 columns with binarized symptoms, each row corresponds to an individual
create_possible_combined_with_real_combinations <- function(data) {
  # Generate new dataframe (all_combinations) with all possible combinations of symptom combination (binarized)
  symptom_cols <- paste0("symptom_", 1:20) # Total number of symptoms
  symptom_values_binarized <- 0:1  # 0: absent, 1: present
  all_combinations <- expand.grid(lapply(1:20, function(i) symptom_values_binarized))
  colnames(all_combinations) <- symptom_cols
  
  # Define criteria for filtering
  criteria <- (
    rowSums(all_combinations[, 1:5]) >= 1 &          # At least 1 symptom from 1-5
      rowSums(all_combinations[, 6:7]) >= 1 &          # At least 1 symptom from 6-7
      rowSums(all_combinations[, 8:14]) >= 2 &         # At least 2 symptoms from 8-14
      rowSums(all_combinations[, 15:20]) >= 2          # At least 2 symptoms from 15-20
  )
  
  # Filter combinations based on criteria
  all_combinations_selectedPTSD <- filter(all_combinations, criteria == TRUE)
  
  # Count frequency of binarized profiles in real-world data
  binarized_data_counted <- data %>%
    group_by_all() %>%
    summarise(freq = n(), .groups = 'drop')
  
  # Merge theoretically possible combinations with their frequency in real-world data
  merged_possible_real <- left_join(all_combinations_selectedPTSD, binarized_data_counted, by = symptom_cols)
  merged_possible_real$freq <- coalesce(merged_possible_real$freq, 0)
 
   return(merged_possible_real)
  }

### Applying function to our data
possible_combined_with_real <- create_possible_combined_with_real_combinations(binarized_diagnosed)
```

### 6.2. Description of the theoretically possible and actually occurring symptom combinations fulfilling diagnosis 

```{r  Description of the theoretically possible and actually occurring symptom combinations fulfilling diagnosis, echo=FALSE, message=FALSE}
### Create function for describing the resulting dataframe
  # Input data frame: resulting data frame of the function create_possible_combined_with_real_combinations, i.e.: 21 columns (20 symptoms and one freq column). The symptoms are each binarized. As rows all theoretically possible combinations that fulfill the diagnostic criteria. The last column freq contains the number of presentations of this symptom combination in the real-world data.
create_description_possible_combined_with_real <- function(data){
  # Filter for only combinations that have actually occurred
  real_combinations <- data %>%
    filter(freq > 0)
  
  ## Description of theoretically and actually occurring combinations
  # How many individuals were diagnosed
  n_pat_diagnosed <- sum(data$freq)
  # How many combinations exist theoretically
  n_possible_combinations <- nrow(data)
  # How many combinations are represented in real-world data
  n_real_combinations <- nrow(real_combinations)
  # Percentage of combinations represented in real-world data
  percentage_real_possible <- round(n_real_combinations/ n_possible_combinations * 100, 2)
  
  # Display of some results in a dataframe
  df1 <- tibble (
    metric = c(
      "Number of individuals fulfilling the diagnosis",
      "Number of theoretically possible combinations ", 
      "Number of combinations that actually occur", 
      "Percentage of possible combinations that actually occur", 
      "How often the most common combination actually occurs",
      "Median frequency in relation to theoretically possible combinations",
      "Median frequency in relation to combinations actually occurring"
    ),
    value = c(
       n_pat_diagnosed,
       n_possible_combinations, 
       n_real_combinations, 
       sprintf("%.2f%%", percentage_real_possible),
       max(real_combinations$freq), 
       stats::median(data$freq),
       stats::median(real_combinations$freq)
    )
  )
  
  ## 10 most common combinations
  ten_most_common_real <- real_combinations %>% 
    dplyr::arrange(desc(freq)) %>% 
    dplyr::slice_head(n = 10)
  
  # Profiles (N, %) reported by less or equal than in 1% of the individuals only respecting combinations that actually occur
  one_percent_individuals = round(n_pat_diagnosed*0.01, 0)
  N_Profiles <- real_combinations %>% filter(freq<=one_percent_individuals) %>% nrow()
  Percentage_Profiles <- round((N_Profiles / nrow(real_combinations))*100, 3)
  
  # Number of individuals reporting one of the ten most common combinations
  N_individuals_ten_most_common <- real_combinations %>%
    arrange(desc(freq)) %>%
    slice_head(n = 10) %>% 
    summarise(sum(freq)) %>%
    pull()
  
  # Number of individuals reporting NOT one of the ten most common combinations
  N_individuals_not_ten_most_common <- real_combinations %>%
    arrange(desc(freq)) %>%
    slice_tail(n = -10) %>% 
    summarise(sum(freq)) %>%
    pull()
  
  # Number of individuals reporting 1% most common combinations
  one_percent_combinations <- round(nrow(real_combinations)*0.01, 0)
  N_individuals_one_perc_most_common <- real_combinations %>%
    arrange(desc(freq)) %>%
    slice_head(n = one_percent_combinations) %>% 
    summarise(sum(freq)) %>%
    pull()
  
  # Number of individuals reporting one of the 50% least common combinations
  fifty_percent_combinations <- round(nrow(real_combinations)*0.5, 0)
  N_individuals_fifty_perc_least_common <- real_combinations %>%
    arrange(desc(freq)) %>%
    slice_tail(n = fifty_percent_combinations) %>% 
    summarise(sum(freq)) %>%
    pull()

  # Median distribution
  median_distribution <- real_combinations %>%
    arrange(desc(freq)) %>%
    mutate (cum_sum = cumsum(freq)) %>% #cumulative sum of frequencies
    mutate (cum_percentage = cum_sum/sum(freq) * 100) %>%
    mutate (median_group = if_else(lag(cum_percentage, default = 0) < 50, "above", "below"))

  N_combinations_first_50percent_individuals <- sum(median_distribution$median_group == "above")
  N_combinations_remaining_50percent_individuals <- sum(median_distribution$median_group == "below")
  
  # Display of some results in a dataframe
  df2 <- tibble(
    metric = c(
      "Number of Profiles reported by 1% of the individuals only respecting combinations that actually occur", 
      "Percentage of Profiles reported by 1% of the individuals only respecting combinations that actually occur", 
      "Number of individuals reporting one of the ten most common combinations", 
      "Number of individuals reporting NOT one of the ten most common combinations",
      "Number of individuals reporting 1% most common combinations",
      "Number of individuals reporting one of the 50% least common combinations", 
      "Number of combinations needed to reach first 50% of individuals",
      "Number of combinations in remaining 50% of individuals"
    ),
    value = c(
    N_Profiles, 
    sprintf("%.2f%%", Percentage_Profiles), 
    N_individuals_ten_most_common,
    N_individuals_not_ten_most_common, 
    N_individuals_one_perc_most_common,
    N_individuals_fifty_perc_least_common,
    N_combinations_first_50percent_individuals,
    N_combinations_remaining_50percent_individuals
    )
  )
  
  ## Description of single symptoms
  symptom_cols <- setdiff(colnames(data), "freq")
  df3 <- map_dfr(symptom_cols, function(col){
    tibble(
      symptom = col,
      N_combinations = sum(real_combinations[[col]] == 1),
      Perc_combinations = sprintf("%.2f%%", (sum(real_combinations[[col]] == 1) / nrow(real_combinations)) * 100),
      N_patients = sum((real_combinations[[col]] == 1)* real_combinations$freq), 
      Perc_patients = sprintf("%.2f%%", (sum((real_combinations[[col]] == 1) * real_combinations$freq) / n_pat_diagnosed) * 100)
    )
  })

  list(
    "summary_statistics" = df1,
    "ten_most_common_combinations" =  ten_most_common_real,
    "distribution_statistics" = df2,
    "symptom_statistics" = df3
  )
}

### Applying function to our data
description_possible_combined_with_real <- create_description_possible_combined_with_real(possible_combined_with_real)
```
Overview statistics when all symptoms are considered
```{r  overview statistics when all symptoms are considered, echo=FALSE, message=FALSE}
datatable(description_possible_combined_with_real$summary_statistics,
          options = list(scrollX = TRUE))
```
Ten most common combinations when all symptoms are considered
```{r  Ten most common combinations when all symptoms are considered, echo=FALSE, message=FALSE}
datatable(description_possible_combined_with_real$ten_most_common_combinations,
          options = list(scrollX = TRUE))
```
Statistics on the distribution of patients and combinations when all symptoms are considered
```{r Statistics on the distribution of patients and combinations when all symptoms are considered, echo=FALSE, message=FALSE}
datatable(description_possible_combined_with_real$distribution_statistics,
          options = list(scrollX = TRUE))
```
Statistics on the prevalence of symptoms among patients and symptom combinations when all symptoms are considered
```{r Statistics on the prevalence of symptoms among patients and symptom combinations when all symptoms are considered, echo=FALSE, message=FALSE}
datatable(description_possible_combined_with_real$symptom_statistics,
          options = list(scrollX = TRUE))
```

### 6.3. Determine the change resulting from the omission of one or more symptoms

#### 6.3.1. Function for neglecting one or more symptoms and then theoretically possible combinations of binarized symptoms fulfilling diagnosis and their frequency in real-world data

```{r  Neglecting one or more symptoms and then theoretically possible combinations of binarized symptoms fulfilling diagnosis and their frequency in real-world data, echo=FALSE, message=FALSE}
### Create function for deleting one or more symptoms. Then theoretically possible combinations of binarized symptoms fulfilling diagnosis and their frequencies in real-world data
  # Input: data: real-world dataset fulfilling diagnostic criteria, 20 columns with binarized symptoms, symptoms_to_delete: number/vector of the symptom(s) you want to delete
create_possible_combined_with_real_symp_del <- function(data, symptoms_to_delete) {
  # Convert single number to vector if necessary
  if(!is.vector(symptoms_to_delete)) {
    symptoms_to_delete <- c(symptoms_to_delete)
  }
  
  # Checks input
  if (!all(sapply(symptoms_to_delete, is.numeric)) || 
      any(symptoms_to_delete < 1) || 
      any(symptoms_to_delete > 20) ||
      length(unique(symptoms_to_delete)) != length(symptoms_to_delete)) {
    stop("symptoms_to_delete must be unique numbers between 1 and 20")
  }
  
  # Create vector of all symptom numbers excluding the ones to delete
  remaining_symptoms <- setdiff(1:20, symptoms_to_delete)
  
  # Create new column names excluding the deleted symptoms
  symptom_cols <- paste0("symptom_", remaining_symptoms)
  
  # Rename the columns in the input data and remove deleted symptoms
  renamed_data <- data %>%
    rename_with(~ paste0("symptom_", 1:20)) %>%
    select(-paste0("symptom_", symptoms_to_delete))  # Remove the deleted symptoms
  
  # Generate new dataframe with all possible combinations
  symptom_values_binarized <- 0:1  # 0: absent, 1: present
  all_combinations <- expand.grid(
    lapply(1:length(remaining_symptoms), function(i) symptom_values_binarized)
  )
  colnames(all_combinations) <- symptom_cols
  
  # Helper function to count symptoms in a range excluding deleted symptoms
  count_in_range <- function(df, start, end) {
    range_symptoms <- intersect(remaining_symptoms, start:end) # Which remaining_symptoms fall in the range
    if(length(range_symptoms) > 0) {
      cols <- paste0("symptom_", range_symptoms)
      return(rowSums(df[, cols, drop = FALSE]))
    }
    return(0)
  }
  
  # Check if we have enough symptoms in each range for the original requirements
  check_range_requirements <- function(start, end, required) {
    range_symptoms <- intersect(remaining_symptoms, start:end)
    if(length(range_symptoms) < required) {
      stop(sprintf("Cannot delete these symptoms - range %d-%d needs at least %d symptom(s) for diagnosis", 
                  start, end, required))
    }
  }
  
  # Check if we have enough symptoms in each range
  check_range_requirements(1, 5, 1)   # Range 1-5 needs at least 1 symptom
  check_range_requirements(6, 7, 1)   # Range 6-7 needs at least 1 symptom
  check_range_requirements(8, 14, 2)  # Range 8-14 needs at least 2 symptoms
  check_range_requirements(15, 20, 2) # Range 15-20 needs at least 2 symptoms
  
  # Define criteria for filtering
  criteria <- (
    count_in_range(all_combinations, 1, 5) >= 1 &    
    count_in_range(all_combinations, 6, 7) >= 1 &    
    count_in_range(all_combinations, 8, 14) >= 2 &   
    count_in_range(all_combinations, 15, 20) >= 2 
  )
  
  # Filter combinations based on criteria
  all_combinations_selectedPTSD <- filter(
    all_combinations, 
    criteria == TRUE
  )
  
  # Count frequency of binarized profiles in real-world data
  binarized_data_counted <- renamed_data %>%
    group_by_all() %>%
    summarise(freq = n(), .groups = 'drop')
  
  # Merge theoretically possible combinations with their frequency in real-world data
  merged_possible_real <- left_join(
    all_combinations_selectedPTSD, 
    binarized_data_counted, 
    by = symptom_cols
  )
  merged_possible_real$freq <- coalesce(
    merged_possible_real$freq, 
    0
  )
  
  return(merged_possible_real)
}
```

#### 6.3.2. Function for describing the resulting data frame with one or more deleted symptoms and for comparing this data set with the data set that takes all symptoms into account. 

```{r Function for describing the resulting data frame with one or more deleted symptoms and for comparing this data set with the data set that takes all symptoms into account., echo=FALSE, message=FALSE}
### Create function for describing the resulting data frame with one or more deleted symptoms and for comparing this data set with the data set that takes all symptoms into account. 
  # Input data frames: resulting data frames of 1. the function create_possible_combined_with_real_combinations and 2. "create_possible_combined_with_real_symp_del". The first data set comprises 21 columns (20 symptoms and one column freq). The second minus the deleted symptoms. The symptoms are each binarized. The rows are all theoretically possible combinations that fulfill the diagnostic criteria. The last column freq contains the number of presentations of this symptom combination in the real data.
create_description_and_comparison_symp_del_possible_real <- function(data_all_symptoms, data_deleted_symptoms) {
  # Get description of both datasets using the original function
  desc_all <- create_description_possible_combined_with_real(data_all_symptoms)
  desc_deleted <- create_description_possible_combined_with_real(data_deleted_symptoms)
  
  # Helper function to safely convert percentage strings to numeric
  safe_percent_convert <- function(x) {
    if(is.character(x)) {
      as.numeric(gsub("[^0-9.-]", "", x))
    } else {
      as.numeric(x)
    }
  }
  
  ## Create comparison dataframes
  # Comparison 1 - Summary statistics
  comparison_df1 <- tibble(
    metric = desc_all$summary_statistics$metric,
    all_symptoms = desc_all$summary_statistics$value,
    deleted_symptoms = desc_deleted$summary_statistics$value,
    difference = mapply(function(all_val, del_val) {
        if(grepl("%", all_val)) {
          sprintf("%.2f%%", safe_percent_convert(del_val) - safe_percent_convert(all_val))
        } else {
          as.numeric(del_val) - as.numeric(all_val)
        }
      }, desc_all$summary_statistics$value,  desc_deleted$summary_statistics$value)
  )
  
  # Comparison 2 - Distribution statistics
  comparison_df2 <- tibble(
    metric = desc_all$distribution_statistics$metric,
    all_symptoms = desc_all$distribution_statistics$value,
    deleted_symptoms = desc_deleted$distribution_statistics$value,
    difference = mapply(function(all_val, del_val) {
        if(grepl("%", all_val)) {
          sprintf("%.2f%%", safe_percent_convert(del_val) - safe_percent_convert(all_val))
        } else {
          as.numeric(del_val) - as.numeric(all_val)
        }
      },  desc_all$distribution_statistics$value, desc_deleted$distribution_statistics$value)
  )

 # Comparison 3 - Symptom statistics
  all_symptoms <- unique(desc_all$symptom_statistics$symptom)

  comparison_df3 <- map_dfr(all_symptoms, function(symptom) {
    row_all <- desc_all$symptom_statistics %>% 
      filter(symptom == !!symptom)
    
    # Check if symptom exists in deleted dataset
    row_deleted <- desc_deleted$symptom_statistics %>% 
      filter(symptom == !!symptom)
   
    if(nrow(row_deleted) > 0) {
      tibble(
        symptom = symptom,
        N_combinations_all = as.character(row_all$N_combinations),
        N_combinations_deleted = as.character(row_deleted$N_combinations),
        N_combinations_diff = as.character(row_deleted$N_combinations - row_all$N_combinations),
        Perc_combinations_all = row_all$Perc_combinations,
        Perc_combinations_deleted = row_deleted$Perc_combinations,
        Perc_combinations_diff = sprintf("%.2f%%", 
          safe_percent_convert(row_deleted$Perc_combinations) - 
          safe_percent_convert(row_all$Perc_combinations)),
        N_patients_all = as.character(row_all$N_patients),
        N_patients_deleted = as.character(row_deleted$N_patients),
        N_patients_diff = as.character(row_deleted$N_patients - row_all$N_patients),
        Perc_patients_all = row_all$Perc_patients,
        Perc_patients_deleted = row_deleted$Perc_patients,
        Perc_patients_diff = sprintf("%.2f%%", 
          safe_percent_convert(row_deleted$Perc_patients) - 
          safe_percent_convert(row_all$Perc_patients))
      )
    } else {
      # For deleted symptoms
      tibble(
        symptom = symptom,
        N_combinations_all = as.character(row_all$N_combinations),
        N_combinations_deleted = "DELETED",
        N_combinations_diff = "NA",
        Perc_combinations_all = row_all$Perc_combinations,
        Perc_combinations_deleted = "DELETED",
        Perc_combinations_diff = "NA",
        N_patients_all = as.character(row_all$N_patients),
        N_patients_deleted = "DELETED",
        N_patients_diff = "NA",
        Perc_patients_all = row_all$Perc_patients,
        Perc_patients_deleted = "DELETED",
        Perc_patients_diff = "NA"
      )
    }
  })
  
  ## Return all results in a list
  list(
    "summary_statistics_comparison" = comparison_df1,
    "distribution_statistics_comparison" = comparison_df2,
    "symptom_statistics_comparison" = comparison_df3
  )
}
```

#### 6.3.3. Application of the generated functions  

##### 6.3.3.1. Deletion of symptom 1 (to illustrate function)

```{r Deletion of symptom 1, echo=FALSE, message=FALSE}
### Applying function create_possible_combined_with_real_symp_del
possible_combined_with_real_symp_del_1 <- create_possible_combined_with_real_symp_del(binarized_diagnosed, 1)

### Applying function create_description_and_comparison_symp_del_possible_real
description_and_comparison_symp_del_1_possible_real <- create_description_and_comparison_symp_del_possible_real(possible_combined_with_real, possible_combined_with_real_symp_del_1)
```
Overview statistics for deletion of symptom 1 (to illustrate function)
```{r Overview statistics for deletion of symptom 1, echo=FALSE, message=FALSE}
datatable(description_and_comparison_symp_del_1_possible_real$summary_statistics_comparison,
          options = list(scrollX = TRUE))
```
Statistics on the distribution of patients and combinations for deletion of symptom 1 (to illustrate function)
```{r Statistics on the distribution of patients and combinations for deletion of symptom 1, echo=FALSE, message=FALSE}
datatable(description_and_comparison_symp_del_1_possible_real$distribution_statistics_comparison,
          options = list(scrollX = TRUE))
```
Statistics on the prevalence of symptoms among patients and symptom combinations for deletion of symptom 1 (to illustrate function)
```{r Statistics on the prevalence of symptoms among patients and symptom combinations for deletion of symptom 1, echo=FALSE, message=FALSE}
datatable(description_and_comparison_symp_del_1_possible_real$symptom_statistics_comparison,
          options = list(scrollX = TRUE))
```

##### 6.3.3.2. Deletion of each symptom and analysis of the impact of the deleted symptom on the diagnosis

```{r Deletion of each symptom and analysis of the impact of the deleted symptom on the diagnosis, echo=FALSE, message=FALSE}
### Function to analyze impact of removing each symptom
  # Input-data: real-world dataset fulfilling diagnostic criteria, 20 columns with binarized symptoms, each row corresponds to an individual
analyze_deletion_each_symptom <- function(data) {
  # Create empty dataframes to store results
  summary_metrics <- tibble(
    scenario = character(),
    N_pat_diagnosed = numeric(),
    N_combinations_theroretical = numeric(),
    N_combinations_actual = numeric(),
    Perc_combination_actual_theoretical = numeric()
  )
  
  distribution_metrics <- tibble(
    scenario = character(),
    N_combinations_under_oneperc_pat = numeric(),
    Perc_combinations_under_oneperc_pat = numeric(),
    N_pat_of_ten_most_common_combinations = numeric(),
    N_pat_not_ten_most_common_combinations = numeric(),
    N_pat_one_perc_most_common_combinations = numeric(),
    N_pat_fifty_perc_least_common_combinations = numeric(),
    N_combinations_first_fifty_perc_of_pat = numeric(),
    N_combinations_remaining_fifty_perc_of_pat = numeric()
  )
  
  ## Get baseline metrics (with all symptoms)
  baseline_data <- create_possible_combined_with_real_combinations(data)
  baseline_desc <- create_description_possible_combined_with_real(baseline_data)
  
  # Add baseline row to summary metrics
  summary_metrics <- summary_metrics %>%
    add_row(
      scenario = "All Symptoms",
      N_pat_diagnosed = as.numeric(baseline_desc$summary_statistics$value[1]),
      N_combinations_theroretical = as.numeric(baseline_desc$summary_statistics$value[2]),
      N_combinations_actual = as.numeric(baseline_desc$summary_statistics$value[3]),
      Perc_combination_actual_theoretical = as.numeric(gsub("[^0-9.]", "", baseline_desc$summary_statistics$value[4]))/100
    )
  
  # Add baseline row to distribution metrics
  distribution_metrics <- distribution_metrics %>%
    add_row(
      scenario = "All Symptoms",
      N_combinations_under_oneperc_pat = as.numeric(baseline_desc$distribution_statistics$value[1]),
      Perc_combinations_under_oneperc_pat = as.numeric(gsub("[^0-9.]", "", baseline_desc$distribution_statistics$value[2]))/100,
      N_pat_of_ten_most_common_combinations = as.numeric(baseline_desc$distribution_statistics$value[3]),
      N_pat_not_ten_most_common_combinations = as.numeric(baseline_desc$distribution_statistics$value[4]),
      N_pat_one_perc_most_common_combinations = as.numeric(baseline_desc$distribution_statistics$value[5]),
      N_pat_fifty_perc_least_common_combinations = as.numeric(baseline_desc$distribution_statistics$value[6]),
      N_combinations_first_fifty_perc_of_pat = as.numeric(baseline_desc$distribution_statistics$value[7]),
      N_combinations_remaining_fifty_perc_of_pat = as.numeric(baseline_desc$distribution_statistics$value[8])
    )
  
  ## Analyze each symptom deletion
  for(i in 1:20) {
    # Calculate metrics without current symptom
    current_data <- create_possible_combined_with_real_symp_del(data, i)
    current_desc <- create_description_possible_combined_with_real(current_data)
    
    # Add row to summary metrics
    summary_metrics <- summary_metrics %>%
      add_row(
        scenario = paste("Without Symptom", i),
        N_pat_diagnosed = as.numeric(current_desc$summary_statistics$value[1]),
        N_combinations_theroretical = as.numeric(current_desc$summary_statistics$value[2]),
        N_combinations_actual = as.numeric(current_desc$summary_statistics$value[3]),
        Perc_combination_actual_theoretical = as.numeric(gsub("[^0-9.]", "", current_desc$summary_statistics$value[4]))/100
      )
    
     # Add row to distribution metrics
    distribution_metrics <- distribution_metrics %>%
      add_row(
        scenario = paste("Without Symptom", i),
        N_combinations_under_oneperc_pat = as.numeric(current_desc$distribution_statistics$value[1]),
        Perc_combinations_under_oneperc_pat = as.numeric(gsub("[^0-9.]", "", current_desc$distribution_statistics$value[2]))/100,
        N_pat_of_ten_most_common_combinations = as.numeric(current_desc$distribution_statistics$value[3]),
        N_pat_not_ten_most_common_combinations = as.numeric(current_desc$distribution_statistics$value[4]),
        N_pat_one_perc_most_common_combinations = as.numeric(current_desc$distribution_statistics$value[5]),
        N_pat_fifty_perc_least_common_combinations = as.numeric(current_desc$distribution_statistics$value[6]),
        N_combinations_first_fifty_perc_of_pat = as.numeric(current_desc$distribution_statistics$value[7]),
        N_combinations_remaining_fifty_perc_of_pat = as.numeric(current_desc$distribution_statistics$value[8])
      )
  }

  ## Calculate differences from baseline  
  summary_metrics <- summary_metrics %>%
    mutate(across(where(is.numeric), 
                  ~ . - first(.),
                  .names = "{col}_diff"))
  
  # Format percentage columns in summary metrics
  summary_metrics <- summary_metrics %>%
    mutate(
      Perc_combination_actual_theoretical = scales::percent(Perc_combination_actual_theoretical, accuracy = 0.01),
      Perc_combination_actual_theoretical_diff = scales::percent(Perc_combination_actual_theoretical_diff, accuracy = 0.01)
    ) %>%
  # Reorder columns to keep metrics and their differences together
    select(
      scenario,
      N_pat_diagnosed, N_pat_diagnosed_diff,
      N_combinations_theroretical, N_combinations_theroretical_diff,
      N_combinations_actual, N_combinations_actual_diff,
      Perc_combination_actual_theoretical, Perc_combination_actual_theoretical_diff
    )
  
  distribution_metrics <- distribution_metrics %>%
    mutate(
      Perc_combinations_under_oneperc_pat = scales::percent(Perc_combinations_under_oneperc_pat, accuracy = 0.01),
    )
  
  return(list(
    summary_metrics = summary_metrics,
    distribution_metrics = distribution_metrics
  ))
}

### Applying function analyze_deletion_each_symptom
analysis_del_each_symptom <- analyze_deletion_each_symptom(binarized_diagnosed)
```
Overview statistics for deletion of each symptom individually
```{r Overview statistics for deletion of each symptom individually, echo=FALSE, message=FALSE}
datatable(analysis_del_each_symptom$summary_metrics,
          options = list(scrollX = TRUE))
```
Statistics on the distribution of patients and combinations for deletion of each symptom individually
```{r Statistics on the distribution of patients and combinations for deletion of each symptom individually, echo=FALSE, message=FALSE}
datatable(analysis_del_each_symptom$distribution_metrics,
          options = list(scrollX = TRUE))
```

##### 6.3.3.3. Extension of the previous analysis: additional analysis of the impact of several simultaneously deleted symptoms on the diagnosis

```{r Extension of the previous analysis: additional analysis of the impact of several simultaneously deleted symptoms on the diagnosis, echo=FALSE, message=FALSE}
### Function to extend the previous analysis in 5.3.3.2: Analyze impact of removing multiple selected symptoms at the same time
  # Input-data: real-world dataset fulfilling diagnostic criteria, 20 columns with binarized symptoms, each row corresponds to an individual
analyze_deletion_multiple_symptoms <- function(data, symptom_combinations) {
  # First, get the single symptom deletion results
  single_deletion_results <- analyze_deletion_each_symptom(data)
  
  # Initialize new rows for summary and distribution metrics
  new_summary_rows <- tibble()
  new_distribution_rows <- tibble()
  
  # Process each combination of symptoms
  for(combo in symptom_combinations) {
    # Convert to vector if necessary
    if(!is.vector(combo)) {
      combo <- c(combo)
    }
    
    # Create scenario name
    scenario_name <- paste("Without Symptoms", paste(combo, collapse = ","))
    
    # Calculate metrics without current combination of symptoms
    current_data <- create_possible_combined_with_real_symp_del(data, combo)
    current_desc <- create_description_possible_combined_with_real(current_data)
    
    # Add row to summary metrics
    new_summary_row <- tibble(
      scenario = scenario_name,
      N_pat_diagnosed = as.numeric(current_desc$summary_statistics$value[1]),
      N_combinations_theroretical = as.numeric(current_desc$summary_statistics$value[2]),
      N_combinations_actual = as.numeric(current_desc$summary_statistics$value[3]),
      Perc_combination_actual_theoretical = as.numeric(gsub("[^0-9.]", "", current_desc$summary_statistics$value[4]))/100
    )
    
    # Calculate differences from baseline
    baseline <- single_deletion_results$summary_metrics %>% 
      filter(scenario == "All Symptoms")
    
    new_summary_row <- new_summary_row %>%
      mutate(
        N_pat_diagnosed_diff = N_pat_diagnosed - baseline$N_pat_diagnosed,
        N_combinations_theroretical_diff = N_combinations_theroretical - baseline$N_combinations_theroretical,
        N_combinations_actual_diff = N_combinations_actual - baseline$N_combinations_actual,
        Perc_combination_actual_theoretical_diff = Perc_combination_actual_theoretical - 
          as.numeric(gsub("[^0-9.]", "", baseline$Perc_combination_actual_theoretical))/100
      )
    
    # Format percentage columns
    new_summary_row <- new_summary_row %>%
      mutate(
        Perc_combination_actual_theoretical = scales::percent(Perc_combination_actual_theoretical, accuracy = 0.01),
        Perc_combination_actual_theoretical_diff = scales::percent(Perc_combination_actual_theoretical_diff, accuracy = 0.01)
      )
    
    new_summary_rows <- bind_rows(new_summary_rows, new_summary_row)
    
    # Add row to distribution metrics
    new_distribution_row <- tibble(
      scenario = scenario_name,
      N_combinations_under_oneperc_pat = as.numeric(current_desc$distribution_statistics$value[1]),
      Perc_combinations_under_oneperc_pat = as.numeric(gsub("[^0-9.]", "", current_desc$distribution_statistics$value[2]))/100,
      N_pat_of_ten_most_common_combinations = as.numeric(current_desc$distribution_statistics$value[3]),
      N_pat_not_ten_most_common_combinations = as.numeric(current_desc$distribution_statistics$value[4]),
      N_pat_one_perc_most_common_combinations = as.numeric(current_desc$distribution_statistics$value[5]),
      N_pat_fifty_perc_least_common_combinations = as.numeric(current_desc$distribution_statistics$value[6]),
      N_combinations_first_fifty_perc_of_pat = as.numeric(current_desc$distribution_statistics$value[7]),
      N_combinations_remaining_fifty_perc_of_pat = as.numeric(current_desc$distribution_statistics$value[8])
    ) %>%
    mutate(
      Perc_combinations_under_oneperc_pat = scales::percent(Perc_combinations_under_oneperc_pat, accuracy = 0.01)
    )
    
    new_distribution_rows <- bind_rows(new_distribution_rows, new_distribution_row)
  }
  
  # Combine original results with new rows
  combined_summary <- bind_rows(
    single_deletion_results$summary_metrics,
    new_summary_rows
  )
  
  combined_distribution <- bind_rows(
    single_deletion_results$distribution_metrics,
    new_distribution_rows
  )
  
  return(list(
    summary_metrics = combined_summary,
    distribution_metrics = combined_distribution
  ))
}

### Applying function analyze_deletion_each_symptom
# Define combinations of symptoms to be analyzed
  # Here I choose the 6 two symptom combinations and the 5 three symptom combinations that lead to the fewest lost diagnoses 
symptom_combinations <- list(
  c(2, 14), c(2, 9), c(3, 14), c(1, 14), c(2, 11), c(3, 9),
  c(2, 14, 18), c(2, 9, 18), c(3, 14, 18), c(1, 14, 18), c(2, 11, 18)
)
analysis_del_multiple_symptoms <- analyze_deletion_multiple_symptoms(binarized_diagnosed, symptom_combinations)
```
Overview statistics for deletion of each symptom individually and selected symptom combinations.
selected symptom combinations: 6 two-symptoms-combinations with fewest lost diagnoses and 5 three-symptoms-combinations with fewest lost diagnoses.
Here: c(2, 14), c(2, 9), c(3, 14), c(1, 14), c(2, 11), c(3, 9), c(2, 14, 18), c(2, 9, 18), c(3, 14, 18), c(1, 14, 18), c(2, 11, 18)
```{r Overview statistics for deletion of each symptom individually and selected symptom combinations, echo=FALSE, message=FALSE}
datatable(analysis_del_multiple_symptoms$summary_metrics,
          options = list(scrollX = TRUE))
```
Statistics on the distribution of patients and combinations for deletion of each symptom individually and selected symptom combinations.
selected symptom combinations: 6 two-symptoms-combinations with fewest lost diagnoses and 5 three-symptoms-combinations with fewest lost diagnoses.
Here: c(2, 14), c(2, 9), c(3, 14), c(1, 14), c(2, 11), c(3, 9), c(2, 14, 18), c(2, 9, 18), c(3, 14, 18), c(1, 14, 18), c(2, 11, 18)
```{r Statistics on the distribution of patients and combinations for deletion of each symptom individually and selected symptom combinations, echo=FALSE, message=FALSE}
datatable(analysis_del_multiple_symptoms$distribution_metrics,
          options = list(scrollX = TRUE))
```

### 6.4. Analysis of possible combinations of binarized symptoms fulfilling the diagnosis according to the original criteria, according to a selected 6-symptom combination and a selected 4-symptom combination

Theoretically possible combinations of binarized symptoms fulfilling a diagnosis and their frequency in real data according to the original diagnostic criteria, according to selected 6-symptom-combination and according to selected 4-symptom-combination
```{r  Theoretically possible combinations of binarized symptoms fulfilling a diagnosis and their frequency in real data according to the original diagnostic criteria, according to selected 6-symptom-combination and according to selected 4-symptom-combination, echo=FALSE, message=FALSE}
### Create function to generate dataframe with three frequency columns for original criteria, selectetd six-symptom-combination, selected four-symptom-combination
  # Input-data: real-world dataset fulfilling diagnostic criteria, 20 columns with binarized symptoms, each row corresponds to an individual
create_possible_combined_with_real_and_symptom_reduction <- function(data,
    four_symptoms_overall,           # Best 4 symptoms overall
    four_symptoms_by_cluster,        # 4 symptoms, one from each cluster
    six_symptoms_diagnostic,         # 6 symptoms following diagnostic criteria (1-1-2-2)
    six_symptoms_by_cluster,         # 6 symptoms with minimum one from each cluster plus two
    six_symptoms_overall) {          # Best 6 symptoms overall
  
  # Generate all possible combinations 
  symptom_cols <- paste0("symptom_", 1:20)
  symptom_values_binarized <- 0:1
  all_combinations <- expand.grid(lapply(1:20, function(i) symptom_values_binarized))
  colnames(all_combinations) <- symptom_cols
  
  # Define original PTSD criteria
  original_criteria <- (
    rowSums(all_combinations[, 1:5] == 1) >= 1 &
      rowSums(all_combinations[, 6:7] == 1) >= 1 &
      rowSums(all_combinations[, 8:14] == 1) >= 2 &
      rowSums(all_combinations[, 15:20] == 1) >= 2
  )
  
  # Filter combinations based on original criteria
  all_combinations_selectedPTSD <- filter(all_combinations, original_criteria == TRUE)
  
  # Helper function to get subset data
  get_subset_data <- function(data, symptoms) {
    subset_indices <- apply(data[, paste0("symptom_", symptoms)], 1, function(row) {
      all(row == 1)
    })
    return(data[subset_indices, ])
  }
  
  # Helper function to count frequencies
  count_frequencies <- function(subset_data, freq_name) {
    subset_data %>%
      group_by_all() %>%
      summarise(!!freq_name := n(), .groups = 'drop')
  }
  
  # Get frequencies for each approach
  original_freq <- count_frequencies(data, "freq_original")
  four_overall_freq <- count_frequencies(get_subset_data(data, four_symptoms_overall), "freq_four_overall")
  four_cluster_freq <- count_frequencies(get_subset_data(data, four_symptoms_by_cluster), "freq_four_cluster")
  six_diagnostic_freq <- count_frequencies(get_subset_data(data, six_symptoms_diagnostic), "freq_six_diagnostic")
  six_cluster_freq <- count_frequencies(get_subset_data(data, six_symptoms_by_cluster), "freq_six_cluster")
  six_overall_freq <- count_frequencies(get_subset_data(data, six_symptoms_overall), "freq_six_overall")
  
 # Merge all together
  merged_possible_real <- all_combinations_selectedPTSD %>%
    left_join(original_freq, by = symptom_cols) %>%
    left_join(four_overall_freq, by = symptom_cols) %>%
    left_join(four_cluster_freq, by = symptom_cols) %>%
    left_join(six_diagnostic_freq, by = symptom_cols) %>%
    left_join(six_cluster_freq, by = symptom_cols) %>%
    left_join(six_overall_freq, by = symptom_cols)
  
  # Replace NAs with 0
  merged_possible_real <- merged_possible_real %>%
    mutate(across(starts_with("freq_"), ~replace_na(., 0)))
  
  return(merged_possible_real)
}

### Applying function to our data
possible_combined_with_real_and_symptom_reduction <- create_possible_combined_with_real_and_symptom_reduction(
  data = binarized_diagnosed,
  four_symptoms_overall = PTSD_Diagnosis_best_four_symptoms_overall_min_false_neg$best_symptoms,
  four_symptoms_by_cluster = variant3_PTSD_Diagnosis_best_four_symptoms_by_cluster_min_false_neg$best_symptoms,
  six_symptoms_diagnostic = variant1_PTSD_Diagnosis_best_six_symptoms_min_false_neg$best_symptoms,
  six_symptoms_by_cluster = PTSD_Diagnosis_best_six_symptoms_by_cluster_min_false_neg$best_symptoms,
  six_symptoms_overall = PTSD_Diagnosis_best_six_symptoms_overall_min_false_neg$best_symptoms
)
```

Description of the theoretically possible and actually occurring symptom combinations fulfilling original diagnostic criteria, a selected 6-symptom-combination and a selected 4-symptom-combination
```{r  Description of the theoretically possible and actually occurring symptom combinations fulfilling original diagnostic criteria, a selected 6-symptom-combination and a selected 4-symptom-combination, echo=FALSE, message=FALSE}
### Create function for describing the resulting dataframe 
  # Input data frame: resulting data frame of the function create_possible_combined_with_real_and_symptom_reduction
create_description_possible_combined_with_real_and_symptom_reduction <- function(data) {
  # Summary statistics (df1)
  df1 <- tibble(
    metric = c(
      "Number of individuals fulfilling diagnosis with original criteria",
      "Number of individuals fulfilling diagnosis with four symptoms overall",
      "Number of individuals fulfilling diagnosis with four symptoms by cluster",
      "Number of individuals fulfilling diagnosis with six symptoms diagnostic",
      "Number of individuals fulfilling diagnosis with six symptoms by cluster",
      "Number of individuals fulfilling diagnosis with six symptoms overall",
     "Number of theoretically possible combinations with original criteria",
      "Number of combinations that actually occur with original criteria",
      "Number of combinations that actually occur with four symptoms overall",
      "Number of combinations that actually occur with four symptoms by cluster",
      "Number of combinations that actually occur with six symptoms diagnostic",
      "Number of combinations that actually occur with six symptoms by cluster",
      "Number of combinations that actually occur with six symptoms overall",
      "Most common combination frequency (original criteria)",
      "Most common combination frequency (four symptoms overall)",
      "Most common combination frequency (four symptoms by cluster)",
      "Most common combination frequency (six symptoms diagnostic)",
      "Most common combination frequency (six symptoms by cluster)",
      "Most common combination frequency (six symptoms overall)"
    ),
    value = c(
      sum(data$freq_original),
      sum(data$freq_four_overall),
      sum(data$freq_four_cluster),
      sum(data$freq_six_diagnostic),
      sum(data$freq_six_cluster),
      sum(data$freq_six_overall),
      nrow(data),
      sum(data$freq_original > 0),
      sum(data$freq_four_overall > 0),
      sum(data$freq_four_cluster > 0),
      sum(data$freq_six_diagnostic > 0),
      sum(data$freq_six_cluster > 0),
      sum(data$freq_six_overall > 0),
      max(data$freq_original),
      max(data$freq_four_overall),
      max(data$freq_four_cluster),
      max(data$freq_six_diagnostic),
      max(data$freq_six_cluster),
      max(data$freq_six_overall)
    )
  )
  
  ## Distribution statistics calculation
  # Helper function for distribution statistics
  calculate_distribution_stats <- function(freq_col) {
    total_patients <- sum(freq_col)
    one_percent_threshold <- total_patients * 0.01
    
    # Sort frequencies
    sorted_freq <- sort(freq_col[freq_col > 0], decreasing = TRUE)
    
    # Calculate metrics
    rare_profiles <- sum(freq_col > 0 & freq_col <= one_percent_threshold)
    n_ten_most_common <- sum(head(sorted_freq, 10))
    n_not_ten_most_common <- sum(tail(sorted_freq, -10))
    
    # Calculate median distribution
    cum_sum <- cumsum(sorted_freq)
    n_combinations_first_50 <- which(cum_sum >= total_patients/2)[1]
    n_combinations_remaining_50 <- length(sorted_freq) - n_combinations_first_50
    
    c(rare_profiles, n_ten_most_common, n_not_ten_most_common,
      n_combinations_first_50, n_combinations_remaining_50)
  }
  
  # Calculate distribution statistics for each approach
  approaches <- c("original", "four_overall", "four_cluster", 
                 "six_diagnostic", "six_cluster", "six_overall")
  
  stats_list <- lapply(approaches, function(app) {
    calculate_distribution_stats(data[[paste0("freq_", app)]])
  })
  
  # Distribution statistics (df2)
  df2 <- tibble(
    metric = rep(c(
      "Profiles reported by 1% of individuals",
      "Number of individuals in 10 most common combinations",
      "Number of individuals not in 10 most common combinations",
      "Combinations needed for first 50% of individuals",
      "Remaining combinations for other 50% of individuals"
    ), times = 6),
    approach = rep(c(
      "Original criteria",
      "Four symptoms overall",
      "Four symptoms by cluster",
      "Six symptoms diagnostic",
      "Six symptoms by cluster",
      "Six symptoms overall"
    ), each = 5),
    value = unlist(stats_list)
  )
  
  return(list(
    "summary_statistics" = df1,
    "distribution_statistics" = df2
  ))
}

### Applying function to our data
description_possible_combined_with_real_and_symptom_reduction <- create_description_possible_combined_with_real_and_symptom_reduction(possible_combined_with_real_and_symptom_reduction)
```

Overview statistics original criteria, selected six-symptom-combinations and selected four-symptom-combinations
```{r  overview statistics original criteria, selected six-symptom-combinations and selected four-symptom-combinations, echo=FALSE, message=FALSE}
datatable(description_possible_combined_with_real_and_symptom_reduction$summary_statistics,
          options = list(scrollX = TRUE))
```

Statistics on the distribution of patients and combinations for original criteria, selected six-symptom-combinations and selected four-symptom-combinations
```{r Statistics on the distribution of patients and combinations for original criteria, selected six-symptom-combinations and selected four-symptom-combinations, echo=FALSE, message=FALSE}
datatable(description_possible_combined_with_real_and_symptom_reduction$distribution_statistics,
          options = list(scrollX = TRUE))
```

## x. Session Info

```{r Session Info, echo=FALSE, message=FALSE}
sessionInfo()
```

## Additional Chunks

### Former 3. Binarizing & create datax
```{r PCL Binarizing, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
## selecting only the "TRUE" Diagnosis
simulated_ptsd_total_selected <- simulated_ptsd_total %>%
  filter(PTSD_Diagnosis=="TRUE")

## Binarization
# set cut-off
cut_off <- 1 #will be used with <= // 

# rename variables to "Q1-QN" for binarization
n_cols <- ncol(simulated_ptsd_total_selected)
simulated_ptsd_selected_rename <- simulated_ptsd_total_selected %>%
  rename_with(~ paste0("Q", seq_len(n_cols - 2)), 1:(n_cols -2 ))

# Binarize
data_binarized <- simulated_ptsd_selected_rename
for (i in 1:(ncol(data_binarized)-2)){
  orig <- paste("q", i, sep = "")
  bin <- paste("Q", i, sep = "")
  data_binarized[orig] <- dplyr::case_when(data_binarized[bin]<= cut_off ~ 0, data_binarized[bin]>cut_off ~ 1)  #0 = "Symptom absent", 1 = "Symptom present"
}

# Create new data frame
data2 <- data_binarized %>% 
  select(PTSD_Diagnosis:ncol(data_binarized)) %>% 
  select(-PTSD_Diagnosis) %>% 
  tibble()

## Count frequency of profiles
data2_counted <- plyr::count(data2[, ])

# Create sum score of endorsed symptoms
data2_counted <- data2_counted %>% 
  mutate(total_bin = rowSums(data2_counted)-freq)

## Create datax 
# Create full dataset
datax <- dplyr::left_join(data_binarized, data2_counted)


## Create dataframes for PsychPower Package 
data_PTSD_PsychPower <- simulated_ptsd_total_selected %>%
  select (symptom_1:symptom_20)
#binarize
data_binarized_PsychPower <- PsychPower::binarize(data_PTSD_PsychPower, cut_off = 1)
#frequency
data_frequency_PsychPower <- PsychPower::pheno_frequency(
  data_binarized_PsychPower, target_columns = tidyselect::starts_with("v_bin"))
```

### Former 4. Description of combinations

Most common combination, its frequency & median frequency

```{r PCL Describe Number, eval=FALSE, message=FALSE, include=FALSE}
desc_pheno_PsychPower <- PsychPower::describe_pheno(data_frequency_PsychPower, frequency = "freq") 
desc_pheno_PsychPower
```

10 Most common Phenotype

```{r PCL Describe 10 most, eval=FALSE, message=FALSE, include=FALSE}
PsychPower::common_pheno(data_frequency_PsychPower, frequency = "freq", n_phenotypes = 10)
```

Profiles (N, %) reported by less or equal than in 1% of the individuals
          
```{r PCL Having less then 1 percent, eval=FALSE, message=FALSE, include=FALSE}
one_percent_sample = round(nrow(data_binarized_PsychPower)*0.01, 0)
data2_counted %>% filter(freq<=one_percent_sample) %>% nrow()
          
round((data2_counted %>% filter(freq<=one_percent_sample) %>% nrow()) / nrow(data2_counted), 3)
```
          
Individuals reporting one of the ten most common combinations
          
```{r PCL Having 10 most, eval=FALSE, message=FALSE, include=FALSE}
data2_counted <- data2_counted %>% 
  arrange(desc(freq))
data2_counted[1:10,] %>% summarise(sum(freq))
```
          
Individuals reporting NOT one of the ten most common combinations
          
```{r PCL  Not Having 10 most, eval=FALSE, message=FALSE, include=FALSE}
data2_counted[11:nrow(data2_counted),] %>% summarise(sum(freq))
```
          
Individuals reporting 1% most common combinations
          
```{r PCL Describe 1 percent, eval=FALSE, message=FALSE, include=FALSE}
one_percent =  round(nrow(data2_counted)*0.01, 0)
          
data2_counted[1:one_percent,] %>% summarise(sum(freq))
```

Individuals reporting one of the 50% least common combinations
          
```{r PCL Describe 50 percent, eval=FALSE, message=FALSE, include=FALSE}
fifty_percent =  round(nrow(data2_counted)*0.5, 0)
          
data2_counted[fifty_percent:nrow(data2_counted),] %>% summarise(sum(freq))
          
```
          
Plot 100 most common profiles
          
```{r Plot 100 most common, eval=FALSE, message=FALSE, include=FALSE}
fig1 <- PsychPower::plot_pheno(data_frequency_PsychPower, frequency = "freq",
                                         n_phenotypes = 100, color = "grey26")
fig1
```          
  
### Former 5.3.1. Function for neglecting one or more symptoms and then theoretically possible combinations of binarized symptoms fulfilling diagnosis and their frequency in real-world data
With helper function to  adjust required symptoms proportionally

```{r Former Neglecting one or more symptoms and then theoretically possible combinations of binarized symptoms fulfilling diagnosis and their frequency in real-world data, eval=FALSE, message=FALSE, include=FALSE}
### Create function for deleting one or more symptoms. Then theoretically possible combinations of binarized symptoms fulfilling diagnosis and their frequencies in real-world data
# Input: data: real-world dataset fulfilling diagnostic criteria, 20 columns with binarized symptoms, symptoms_to_delete: number/vector of the symptom(s) you want to delete
create_possible_combined_with_real_symp_del <- function(data, symptoms_to_delete) {
  # Convert single number to vector if necessary
  if(!is.vector(symptoms_to_delete)) {
    symptoms_to_delete <- c(symptoms_to_delete)
  }
  
  # Checks input
  if (!all(sapply(symptoms_to_delete, is.numeric)) || 
      any(symptoms_to_delete < 1) || 
      any(symptoms_to_delete > 20) ||
      length(unique(symptoms_to_delete)) != length(symptoms_to_delete)) {
    stop("symptoms_to_delete must be unique numbers between 1 and 20")
  }
  
  # Create vector of all symptom numbers excluding the ones to delete
  remaining_symptoms <- setdiff(1:20, symptoms_to_delete)
  
  # Ensure we have enough symptoms left for the criteria
  if(length(remaining_symptoms) < 6) {
    stop("Cannot delete that many symptoms - need enough symptoms to maintain criteria requirements")
  }
  
  # Create new column names excluding the deleted symptoms
  symptom_cols <- paste0("symptom_", remaining_symptoms)
  
  ## Rename the columns in the input data and remove deleted symptoms
  renamed_data <- data %>%
    rename_with(~ paste0("symptom_", 1:20)) %>%
    select(-paste0("symptom_", symptoms_to_delete))  # Remove the deleted symptoms
  
  # Generate new dataframe with all possible combinations
  symptom_values_binarized <- 0:1  # 0: absent, 1: present
  all_combinations <- expand.grid(
    lapply(1:length(remaining_symptoms), function(i) symptom_values_binarized)
  )
  colnames(all_combinations) <- symptom_cols
  
  # Helper function to count symptoms in a range excluding deleted symptoms
  count_in_range <- function(df, start, end) {
    range_symptoms <- intersect(remaining_symptoms, start:end) # Which remaining_symptoms fall in the range
    if(length(range_symptoms) > 0) {
      cols <- paste0("symptom_", range_symptoms)
      return(rowSums(df[, cols, drop = FALSE]))
    }
    return(0)
  }
  
  # Helper function to calculate required symptoms based on remaining symptoms in range 
  #(original_required: how many symptoms in this range must be fulfilled in the original criteria?)
  get_required_symptoms <- function(start, end, original_required) {
    range_symptoms <- intersect(remaining_symptoms, start:end)
    if(length(range_symptoms) == 0) {
      stop(sprintf("Cannot delete all symptoms from range %d-%d as at least %d symptom(s) from this range are required for diagnosis", 
                   start, end, original_required))
    }
    # Adjust required symptoms proportionally, rounding up
    new_required <- ceiling(original_required * length(range_symptoms) / length(start:end))
    return(new_required)
  }
  
  # Calculate adjusted requirements based on remaining symptoms
  req_1_5 <- get_required_symptoms(1, 5, 1)
  req_6_7 <- get_required_symptoms(6, 7, 1)
  req_8_14 <- get_required_symptoms(8, 14, 2)
  req_15_20 <- get_required_symptoms(15, 20, 2)
  
  # Define criteria for filtering with adjusted ranges and requirements
  criteria <- (
    count_in_range(all_combinations, 1, 5) >= req_1_5 &
      count_in_range(all_combinations, 6, 7) >= req_6_7 &
      count_in_range(all_combinations, 8, 14) >= req_8_14 &
      count_in_range(all_combinations, 15, 20) >= req_15_20
  )
  
  # Filter combinations based on criteria
  all_combinations_selectedPTSD <- filter(
    all_combinations, 
    criteria == TRUE
  )
  
  # Count frequency of binarized profiles in real-world data
  binarized_data_counted <- renamed_data %>%
    group_by_all() %>%
    summarise(freq = n(), .groups = 'drop')
  
  # Merge theoretically possible combinations with their frequency in real-world data
  merged_possible_real <- left_join(
    all_combinations_selectedPTSD, 
    binarized_data_counted, 
    by = symptom_cols
  )
  merged_possible_real$freq <- coalesce(
    merged_possible_real$freq, 
    0
  )
  
  return(merged_possible_real)
}
```

### Former 6. Theoretically possible combinations of nonbinarized symptoms fulfilling diagnosis and their frequency in real-world data

```{r Theoretically possible combinations of nonbinarized symptoms fulfilling diagnosis and their frequency in real-world data, eval=FALSE, message=FALSE, include=FALSE}
### DOES NOT WORK, BECAUSE MEMORY LIMIT IS REACHED, I am searching for alternative options (the use of chunks to handle the memory limit does not work either)
### Function for Dataframe with all possible combinations of nonbinarized symptoms fulfilling PTSD and their frequencies in real-world data
create_possible_combinedwith_real_nonbinarized <- function(data) {
  # Generate new dataframe (all_combinations_nonbinarized) with all possible combinations of symptom values
  symptom_cols_nonbinarized <- paste0("symptom_", 1:20)
  symptom_values_nonbinarized <- 0:4
  all_combinations_nonbinarized <- expand.grid(lapply(1:20, function(i) symptom_values_nonbinarized))
  colnames(all_combinations_nonbinarized) <- symptom_cols_nonbinarized
  
  # Calculate the total score for each row
  all_combinations_nonbinarized$total <- rowSums(all_combinations_nonbinarized)
  
  # Define criteria for filtering
  criteria_1 <- all_combinations_nonbinarized$total >= 33
  criteria_2 <- (
    rowSums(all_combinations_nonbinarized[, 1:5] >= 2) >= 1 &
      rowSums(all_combinations_nonbinarized[, 6:7] >= 2) >= 1 &
      rowSums(all_combinations_nonbinarized[, 8:14] >= 2) >= 2 &
      rowSums(all_combinations_nonbinarized[, 15:20] >= 2) >= 2
  )
  
  # Filter combinations based on criteria
  all_combinations_nonbinarized_selectedPTSD <- all_combinations_nonbinarized[criteria_1 | criteria_2, ]
  
  # Count frequency of nonbinarized profiles in real-world data
  # First ensure only using symptom columns
  simulated_ptsd_total_selected <- data[, symptom_cols_nonbinarized]
  
  simulated_ptsd_total_selected_counted <- simulated_ptsd_total_selected %>%
    group_by_all() %>%
    summarise(freq = n(), .groups = 'drop')
  
  # Merge with the frequency data
  merged_possible_real_nonbinarized <- left_join(all_combinations_nonbinarized_selectedPTSD, simulated_ptsd_total_selected_counted, by = symptom_cols_nonbinarized)
  merged_possible_real_nonbinarized$freq <- coalesce(merged_possible_real_nonbinarized$freq, 0)
  
  return(merged_possible_real_nonbinarized)
}

## Applying function to our data
possible_combinedwith_real_nonbinarized <- create_possible_combinedwith_real_nonbinarized(simulated_ptsd_total_selected)

```

### Former 7. Symptomanalysis, conditional probabilities, pairwise consideration, symptomgroup analysis

```{r Symptomanalysis, conditional probabilities, pairwise consideration, symptomgroup analysis, eval=FALSE, message=FALSE, echo=FALSE, include=FALSE}
### Create function for Analysis of individual symptom frequencies, conditional probabilities, pairwise considerations, chi-square tests, symptom groups analysis
# Input data frames: 20 columns with non-binarized symptoms, all patients included before application of diagnostic criteria
analyze_symptoms_condprob_pairs <- function(data) {
  ## Preparation of inputdata
  # Input validation
  if (!is.data.frame(data) || ncol(data) < 20) {
    stop("Input must be a data frame with at least 20 symptom columns")
  }
  
  # Function to binarize symptom scores with cutoff of 1 in orginial dataframe
  binarize_symptoms <- function(data, cutoff = 1) {
    result <- as.data.frame(lapply(data, function(x) {
      as.numeric(x > cutoff)
    }))
    return(result)
  }
  # Apply to our data
  binarized_all <- binarize_symptoms(data)
  
  # Function to check if diagnostic criteria are met for binarized data (without consideration of totalscore)
  check_diagnosis <- function(data) {
    criteria1 <- rowSums(data[, 1:5] > 0) >= 1     # At least 1 from 1-5
    criteria2 <- rowSums(data[, 6:7] > 0) >= 1     # At least 1 from 6-7
    criteria3 <- rowSums(data[, 8:14] > 0) >= 2    # At least 2 from 8-14
    criteria4 <- rowSums(data[, 15:20] > 0) >= 2   # At least 2 from 15-20
    
    criteria1 & criteria2 & criteria3 & criteria4
  }
  # Add diagnosis column
  binarized_all$PTSD_Diagnosis <- check_diagnosis(binarized_all)
  
  ## Create list to store results
  results <- list()
  
  ## 1. Individual symptom frequencies
  results$symptom_frequencies <- list(
    # 1.1 Overall frequencies
    freq_overall = colMeans(binarized_all[, 1:20]),
    # 1.2 Frequencies in diagnosed patients
    freq_diagnosed = colMeans(binarized_all[binarized_all$PTSD_Diagnosis == TRUE, 1:20]),
    # 1.3 Frequencies in non-diagnosed patients
    freq_nondiagnosed = colMeans(binarized_all[binarized_all$PTSD_Diagnosis == FALSE, 1:20])
  )
  
  ## 2. Conditional probabilities
  # 2.1 P(Diagnosis | Symptom present)
  results$conditional_probs <- lapply(1:20, function(i) {
    prob_diag_if_present = sum(binarized_all$PTSD_Diagnosis == TRUE & binarized_all[,i] == 1) / max(sum(binarized_all[,i] == 1), 1)
    # 2.2 P(Diagnosis | Symptom absent)
    prob_diag_if_absent <- sum(binarized_all$PTSD_Diagnosis == FALSE & binarized_all[,i] == 0) /  max(sum(binarized_all[,i] == 0), 1)
    # 2.3 Relative risk
    relative_risk <- if(prob_diag_if_absent > 0) {
      prob_diag_if_present / prob_diag_if_absent
    } else {
      NA
    }
    list(
      prob_diag_if_present = prob_diag_if_present,
      prob_diag_if_absent = prob_diag_if_absent,
      relative_risk = relative_risk
    )
  })
  
  ## 3. Pairwise considerations
  results$correlations <- list(
    # 3.1 Overall correlation matrix
    overall_cor = cor(binarized_all[, 1:20]),
    # 3.2 Correlation matrix for diagnosed patients
    diagnosed_cor = cor(binarized_all[binarized_all$PTSD_Diagnosis == TRUE, 1:20])
  )
  # 3.3.-3.6. 
  n <- 20
  # 3.3. Observed joint probabilities overall and for diagnosed patients
  joint_prob_overall <- matrix(0, n, n)
  joint_prob_diagnosed <- matrix(0, n, n)
  
  for(i in 1:n) {
    for(j in 1:n) {
      joint_prob_overall[i,j] <- mean(binarized_all[,i] & binarized_all[,j])
      joint_prob_diagnosed[i,j] <- mean(binarized_all[binarized_all$PTSD_Diagnosis == TRUE,i] & binarized_all[binarized_all$PTSD_Diagnosis == TRUE,j])
    }
  }
  
  # 3.4. Expected under independence overall and for diagnosed patients
  expected_prob_overall <- outer(results$symptom_frequencies$freq_overall, results$symptom_frequencies$freq_overall)
  expected_prob_diagnosed <- outer(results$symptom_frequencies$freq_diagnosed, results$symptom_frequencies$freq_diagnosed)
  
  # 3.5. Ratio observed and expected overall
  ratio_overall <- ifelse(expected_prob_overall > 0, joint_prob_overall/expected_prob_overall, NA)
  # 3.6. Comparison observed and expected for diagnosed patients
  ratio_diagnosed <- ifelse(expected_prob_diagnosed > 0, joint_prob_diagnosed/expected_prob_diagnosed, NA)
  # 3.7. Difference observed vs expected overall
  diff_prob_overall <- joint_prob_overall - expected_prob_overall
  # 3.8. Difference observed vs expected for diagnosed patients
  diff_prob_diagnosed <- joint_prob_diagnosed -  expected_prob_diagnosed
  
  # Store 3.3 -3.8. in a list
  results$symptom_relationships <- list(
    observed = list(
      overall = joint_prob_overall,
      diagnosed = joint_prob_diagnosed
    ),
    expected = list(
      overall = expected_prob_overall,
      diagnosed = expected_prob_diagnosed
    ),
    ratios = list(
      overall = ratio_overall,
      diagnosed = ratio_diagnosed
    ),
    differences = list(
      overall = diff_prob_overall,
      diagnosed = diff_prob_diagnosed
    )
  )
  
  ## 4. Symptom groups analysis
  # 4.1 Probability of diagnosis when group criteria are met
  group_criteria <- list(
    group1 = rowSums(binarized_all[,1:5]) >= 1,
    group2 = rowSums(binarized_all[,6:7]) >= 1,
    group3 = rowSums(binarized_all[,8:14]) >= 2,
    group4 = rowSums(binarized_all[,15:20]) >= 2
  )
  
  results$group_prob_diagnosis <- lapply(group_criteria, function(criteria) {
    sum(binarized_all$PTSD_Diagnosis == TRUE & criteria) / sum(criteria)
  })
  return(results)
}

### Applying function to our data
symptomanalysis_condprob_pairs <- analyze_symptoms_condprob_pairs(simulated_ptsd_data)
```

### Former 8. Calculation of the probability of different symptom combinations with binarized symptoms that meet the diagnostic criteria using different models

```{r Calculation of the probability of different symptom combinations with binarized symptoms that meet the diagnostic criteria using different models, eval=FALSE, message=FALSE, include=FALSE}
### PROBABLY NOT POSSIBLE, AS THERE IS TOO MUCH DATA; NOT EVEN WHEN PERFORMING ONE METHOD INDIVIDUALLY EXCEPT FOR THE INDEPENDENCE METHOD
# Required packages
library(copula)
library(mvtnorm)
library(glmnet)
library(IsingSampler)

### Creation of a function to calculate the probability of different combinations of symptoms with binarized symptoms meeting the diagnostic criteria using different models (Independence Method, Observed Frequencies, Gaussian Copula, Multivariate Normal, Log-linear Model, Ising Model)
# Input: 20 columns with non-binarized symptoms, all patients included before application of diagnostic criteria
calculate_sympcombprob <- function(raw_data) {
  # Helper function to binarize symptoms
  binarize_symptoms <- function(data, cutoff = 1) {
    result <- as.data.frame(lapply(data, function(x) {
      as.numeric(x > cutoff)
    }))
    return(result)
  }
  
  ## Step 1: Binarize the input data
  binarized_all <- binarize_symptoms(raw_data)
  
  ## Step 2: Apply criteria to create valid_binarized_all
  criteria_met_data <- (
    rowSums(binarized_all[, 1:5]) >= 1 &      # At least 1 symptom from 1-5
      rowSums(binarized_all[, 6:7]) >= 1 &      # At least 1 symptom from 6-7
      rowSums(binarized_all[, 8:14]) >= 2 &     # At least 2 symptoms from 8-14
      rowSums(binarized_all[, 15:20]) >= 2      # At least 2 symptoms from 15-20
  )
  
  # Filter binarized data based on criteria
  valid_binarized_all <- binarized_all[criteria_met_data, ]
  
  ## Step 3: Calculate covariance matrix from binarized data
  covariance_matrix <- cov(valid_binarized_all)
  
  ## Step 4: Calculate symptom frequencies
  symptom_frequencies <- colMeans(valid_binarized_all)
  
  ## Step 5: Generate all possible symptom combinations
  symptom_cols <- paste0("symptom_", 1:20)
  symptom_values <- 0:1
  all_combinations <- expand.grid(lapply(1:20, function(i) symptom_values))
  colnames(all_combinations) <- symptom_cols
  
  ## Step 6: Apply criteria
  criteria_met <- (
    rowSums(all_combinations[, 1:5]) >= 1 &      # At least 1 symptom from 1-5
      rowSums(all_combinations[, 6:7]) >= 1 &      # At least 1 symptom from 6-7
      rowSums(all_combinations[, 8:14]) >= 2 &     # At least 2 symptoms from 8-14
      rowSums(all_combinations[, 15:20]) >= 2      # At least 2 symptoms from 15-20
  )
  
  # Filter combinations based on criteria
  valid_combinations <- all_combinations[criteria_met, ]
  
  ## Step 7: Calculate probabilities for valid combinations with different methods
  # 7.1. Method 1: Independence Method
  independence_prob <- function(combination) { #combination = single row
    prob <- 1
    for(i in 1:length(combination)) { #for each position in this combination
      if(combination[i] == 1) {
        prob <- prob * symptom_frequencies[i]
      } else {
        prob <- prob * (1 - symptom_frequencies[i])
      }
    }
    return(prob)
  }
  valid_combinations$prob_independence <- apply(valid_combinations, 1, independence_prob)
  
  
  # 7.2. Method 2: Observed Frequencies
  observed_combinations <- apply(valid_binarized_all, 1, paste, collapse = "") #each row into a single string
  get_observed_freq <- function(combination) {
    pattern <- paste(combination, collapse = "") #convert combination to a string pattern
    freq <- sum(observed_combinations == pattern) / length(observed_combinations) #how many times this exact pattern appears, divide by total number of observations
    return(max(freq, 1e-10)) # Add small constant to avoid zero probabilities
  }
  valid_combinations$prob_observed <- apply(valid_combinations, 1, get_observed_freq)
  
  # 7.3. Method 3: Gaussian Copula
  gaussian_copula_prob <- function(combination, covariance_matrix) {
    # Create normal copula object
    normal_copula <- normalCopula(param = P2p(covariance_matrix),
                                  dim = ncol(covariance_matrix),
                                  dispstr = "un")
    
    # Convert binary values to uniform margins
    u <- ifelse(as.numeric(combination) == 1, 0.999, 0.001)  # Avoid exact 0s and 1s
    
    # Calculate probability using copula
    prob <- pCopula(u, normal_copula)
    return(max(prob, 1e-10))
  }
  
  valid_combinations$prob_copula <- apply(valid_combinations, 1,
                                          function(x) gaussian_copula_prob(x, covariance_matrix))
  
  # 7.4. Method 4: Multivariate Normal
  # Convert covariance matrix to correlation matrix
  correlation_matrix <- cov2cor(covariance_matrix)
  
  # Initialize probability vector
  n_combinations <- nrow(valid_combinations)
  mvn_probabilities <- numeric(n_combinations)
  
  # Calculate MVN probability for each combination
  for(i in 1:n_combinations) {
    current_combination <- as.numeric(valid_combinations[i,])
    
    # Use actual frequencies as means
    means <- symptom_frequencies
    
    # Set bounds based on presence/absence of symptoms
    lower_bounds <- ifelse(current_combination == 1, symptom_frequencies, -Inf)
    upper_bounds <- ifelse(current_combination == 1, Inf, symptom_frequencies)
    
    # Calculate probability
    prob <- try({
      mvtnorm::pmvnorm(
        lower = lower_bounds,
        upper = upper_bounds,
        mean = means,
        sigma = correlation_matrix
      )
    }, silent = TRUE)
    
    if(!inherits(prob, "try-error")) {
      mvn_probabilities[i] <- prob[1]
    } else {
      mvn_probabilities[i] <- NA
    }
  }
  
  # Add MVN probabilities to results
  valid_combinations$prob_mvn <- mvn_probabilities
  
  # 7.5. Method 5: Log-linear Model # ERROR BECAUSE NO VARIATION
  X <- as.matrix(valid_binarized_all) #Convert binarized data to a matrix
  X[is.na(X)] <- 0 #Replace NAs with 0
  y <- rep(1, nrow(X)) #Create vector of 1's with same length as number of patients for constant response for density estimation
  loglinear_model <- cv.glmnet(X, y, family = "poisson")
  
  # Define probability function
  loglinear_prob <- function(combination) {
    # Ensure combination is properly formatted
    combination_matrix <- matrix(as.numeric(combination), nrow = 1)
    prob <- exp(predict(loglinear_model, newx = combination_matrix, s = "lambda.min"))
    return(max(prob[1], 1e-10))
  }
  valid_combinations$prob_loglinear <- apply(valid_combinations, 1, loglinear_prob)
  
  # 7.6. Method 6: Ising Model
  # Fit Ising model
  ising_data <- as.matrix(valid_binarized_all)
  # Convert to numeric and ensure binary
  ising_data <- matrix(as.numeric(ising_data), nrow = nrow(ising_data))
  ising_data[is.na(ising_data)] <- 0
  
  ising_model <- EstimateIsing(ising_data, method = "pl") # Pseudo-likelihood estimation
  
  ising_prob <- function(combination) {
    # Ensure combination is numeric binary
    combination <- as.numeric(combination)
    # Calculate energy of the configuration
    energy <- CalculateEnergy(combination, ising_model$graph)
    # Convert energy to probability
    prob <- exp(-energy) / ising_model$Z # Z is the partition function
    return(max(prob, 1e-10))
  }
  valid_combinations$prob_ising <- apply(valid_combinations, 1, ising_prob)
  
  # Normalize probabilities to sum to 1 for each method
  normalize_probs <- function(probs) {
    return(probs / sum(probs, na.rm = TRUE))
  }
  
  prob_columns <- grep("^prob_", colnames(valid_combinations), value = TRUE)
  valid_combinations[prob_columns] <- lapply(valid_combinations[prob_columns], normalize_probs)
  
  return(valid_combinations)
}

### Applying function to our data
sympcomb_prob_binarized <- calculate_sympcombprob(simulated_ptsd_data)
```
