---
title: "Assess Simulated PTSD Symptoms"
author:
- "<h5 style=\"font-style:italic\"> Laura Weidmann"
- "<h5 style=\"font-style:italic\"> Tobias R. Spiller"
date: "<h5 style=\"font-style:roman\"> `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
  pdf_document:
    toc: true
    toc_depth: '5'
subtitle: Version 0.0.2
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)
```

```{r Load Libraries, message=FALSE, warning=TRUE, include=FALSE}
# Data handling
library(tidyverse)
library(readr)
library(data.table)

# Demographics
library(table1)
library(gtsummary)

# Analysis
#library("devtools")
#devtools::install_github("orduek/PsychPower")
library(PsychPower)
```

## 1. Import and prepare data

```{r PCL import data, message=FALSE, warning=TRUE, include=FALSE}
# Import data
simulated_ptsd_data <- read_csv("Data/simulated_skewed_ptsd_data.csv")
```

## 2. Sample descriptive

#### 2.1 Total score & Diagnosis

```{r PCL Table 1, echo=FALSE, message=FALSE}
# Create PCL total score
simulated_ptsd_total <- simulated_ptsd_data %>% 
  mutate(total = rowSums(select(., symptom_1:symptom_20)))

# # Create Diagnosis
#simulated_ptsd_total <- simulated_ptsd_total %>%
  #mutate(
    #PTSD_Diagnosis = ifelse(
      #(symptom_1 > 2 | symptom_2 > 2 | symptom_3 > 2 | symptom_4 > 2 | symptom_5 > 2) &
      #(symptom_6 > 2 | symptom_7 > 2) &
      #(symptom_8 > 2 | symptom_9 > 2 | symptom_10 > 2 | symptom_11 > 2 | symptom_12 > 2 | symptom_13 > 2 | symptom_14 #> 2) &
      #(symptom_15 > 2 | symptom_16 > 2 | symptom_17 > 2 | symptom_18 > 2 | symptom_19 > 2 | symptom_20 > 2),
      #TRUE, FALSE
    #)
  #)

# Create function that implements PTSD diagnostic criteria
  create_ptsd_diagnosis <- function(data) {
    
    # Create new column initially set to FALSE
    data$PTSD_Diagnosis <- FALSE
    
    # Check total score criterion
    total_criterion <- data$total >= 33
    
    # Check symptom criteria
    criterion_B <- rowSums(data[, paste0("symptom_", 1:5)] >= 2) >= 1  # At least 1 from symptoms 1-5
    criterion_C <- rowSums(data[, paste0("symptom_", 6:7)] >= 2) >= 1  # At least 1 from symptoms 6-7
    criterion_D <- rowSums(data[, paste0("symptom_", 8:14)] >= 2) >= 2 # At least 2 from symptoms 8-14
    criterion_E <- rowSums(data[, paste0("symptom_", 15:20)] >= 2) >= 2 # At least 2 from symptoms 15-20
    
    # All symptom criteria met
    all_symptom_criteria <- criterion_B & criterion_C & criterion_D & criterion_E
    
    # Combined criteria: Either total >= 33 OR all symptom criteria met OR both conditions together
    data$PTSD_Diagnosis <- total_criterion | 
      all_symptom_criteria |
      (total_criterion & all_symptom_criteria)
    
    return(data)
  }

#Apply the function to our dataframe
  simulated_ptsd_total <- create_ptsd_diagnosis (simulated_ptsd_total)

#Summarize
simulated_ptsd_total %>%
  summarise(
    mean_total = mean(total),
    sd_total = sd(total),
    n_diagnosed = sum(PTSD_Diagnosis)
  )
```

#### 2.2 Cronbach's alpha

Selected sample

```{r PCL cronbach, echo=FALSE, message=FALSE}
cronbach <- psych::alpha(subset(simulated_ptsd_total, select = (-total)))
cronbach$total
```

### 2.3 Summary of items & histogram

```{r PCL summary, echo=FALSE, message=FALSE}
summary(simulated_ptsd_total)
hist(simulated_ptsd_total$total)
```

## 3. Binarizing & create datax

Symptoms rated 2 or 3 will be considered as present. [not shown]

```{r PCL Binarizing, echo=FALSE, message=FALSE, warning=FALSE}
#selecting only the "TRUE" Diagnosis
simulated_ptsd_total_selected <- simulated_ptsd_total %>%
  filter(PTSD_Diagnosis=="TRUE")

## set cut-off
cut_off <- 1 #will be used with <= // 

##rename variables to "Q1-QN" for binarization
n_cols <- ncol(simulated_ptsd_total)
simulated_ptsd_selected_rename <- simulated_ptsd_total_selected %>%
  rename_with(~ paste0("Q", seq_len(n_cols - 2)), 1:(n_cols -2 ))

## Binarize
data_binarized <- simulated_ptsd_selected_rename

for (i in 1:(ncol(data_binarized)-2)){
  orig <- paste("q", i, sep = "")
  bin <- paste("Q", i, sep = "")
  data_binarized[orig] <- dplyr::case_when(data_binarized[bin]<= cut_off ~ 0, data_binarized[bin]>cut_off ~ 1)  #0 = "Symptom absent", 1 = "Symptom present"
}

# Create new data frame
data2 <- data_binarized %>% 
  select(PTSD_Diagnosis:ncol(data_binarized)) %>% 
  select(-PTSD_Diagnosis) %>% 
  tibble()

## Count frequency of profiles
data2_counted <- plyr::count(data2[, ]) 

# Create sum score of endorsed symptoms
data2_counted <- data2_counted %>% 
  mutate(total_bin = rowSums(data2_counted)-freq)

## Create datax 
# Create full dataset
datax <- dplyr::left_join(data_binarized, data2_counted)


## Create dataframes for PsychPower Package 
data_PTSD_PsychPower <- simulated_ptsd_total_selected %>%  
  	select (symptom_1:symptom_20)
#binarize
data_binarized_PsychPower <- PsychPower::binarize(data_PTSD_PsychPower, cut_off = 1)
#frequency
data_frequency_PsychPower <- PsychPower::pheno_frequency(
  data_binarized_PsychPower, target_columns = tidyselect::starts_with("v_bin"))
```

## 4. Description of combinations

Most common combination, its frequency & median frequency

```{r PCL Describe Number, echo=FALSE, message=FALSE}
desc_pheno_PsychPower <- PsychPower::describe_pheno(data_frequency_PsychPower, frequency = "freq") 
desc_pheno_PsychPower
```

10 Most common Phenotype

```{r PCL Describe 10 most, echo=FALSE, message=FALSE}
PsychPower::common_pheno(data_frequency_PsychPower, frequency = "freq", n_phenotypes = 10)
```

Profiles (N, %) reported by less or equal than in 1% of the individuals

```{r PCL Having less then 1 percent, echo=FALSE, message=FALSE}
one_percent_sample = round(nrow(data_binarized_PsychPower)*0.01, 0)
data2_counted %>% filter(freq<=one_percent_sample) %>% nrow()

round((data2_counted %>% filter(freq<=one_percent_sample) %>% nrow()) / nrow(data2_counted), 3)
```

Individuals reporting one of the ten most common combinations

```{r PCL Having 10 most, echo=FALSE, message=FALSE}
data2_counted <- data2_counted %>% 
  arrange(desc(freq))
data2_counted[1:10,] %>% summarise(sum(freq))
```

Individuals reporting NOT one of the ten most common combinations

```{r PCL  Not Having 10 most, echo=FALSE, message=FALSE}
data2_counted[11:nrow(data2_counted),] %>% summarise(sum(freq))
```

Individuals reporting 1% most common combinations

```{r PCL Describe 1 percent, echo=FALSE, message=FALSE}
one_percent =  round(nrow(data2_counted)*0.01, 0)

data2_counted[1:one_percent,] %>% summarise(sum(freq))
```

Individuals reporting one of the 50% least common combinations

```{r PCL Describe 50 percent, echo=FALSE, message=FALSE}
fifty_percent =  round(nrow(data2_counted)*0.5, 0)

data2_counted[fifty_percent:nrow(data2_counted),] %>% summarise(sum(freq))

```

Plot 100 most common profiles

```{r Plot 100 most common, echo=FALSE, message=FALSE}
fig1 <- PsychPower::plot_pheno(data_frequency_PsychPower, frequency = "freq",
                               n_phenotypes = 100, color = "grey26")
fig1
```

## 5. Analysis of the possible combinations of binarized symptoms that fulfill the diagnosis (without consideration of the total score criterion)
#### 5.1. Theoretically possible combinations of binarized symptoms fulfilling diagnosis (without consideration of the total score criterion) and their frequency in real-world data
```{r Theoretically possible combinations of binarized symptoms fulfilling diagnosis (without consideration of the total score criterion) and their frequency in real-world data, echo=FALSE, message=FALSE}
 ### Create function
  # Input-data: real-world dataset fulfilling diagnostic criteria, 20 columns with binarized symptoms
create_possible_combinedwith_real_binarized_without_total <- function(data) {
  # First rename the columns in the input data (real-world) to match the expected format
  renamed_data <- data %>%
    rename_with(~ paste0("symptom_", 1:20))
  
  # Generate new dataframe (all_combinations_binarized_without_total) with all possible combinations of symptom combination (binarized) without respecting the criteria of totalscore
  symptom_cols_binarized_without_total <- paste0("symptom_", 1:20) # Total number of symptoms
  symptom_values_binarized_without_total <- 0:1  # 0: absent, 1: present
  all_combinations_binarized_without_total <- expand.grid(lapply(1:20, function(i) symptom_values_binarized_without_total))
  colnames(all_combinations_binarized_without_total) <- symptom_cols_binarized_without_total
  
  # Define criteria for filtering
  criteria_1 <- (
    rowSums(all_combinations_binarized_without_total[, 1:5]) >= 1 &          # At least 1 symptom from 1-5
      rowSums(all_combinations_binarized_without_total[, 6:7]) >= 1 &          # At least 1 symptom from 6-7
      rowSums(all_combinations_binarized_without_total[, 8:14]) >= 2 &         # At least 2 symptoms from 8-14
      rowSums(all_combinations_binarized_without_total[, 15:20]) >= 2          # At least 2 symptoms from 15-20
  )
  
  # Filter combinations based on criteria
  all_combinations_binarized_without_total_selectedPTSD <- filter(all_combinations_binarized_without_total, criteria_1==TRUE)
  
  # Count frequency of binarized profiles in real-world data
  binarized_data_counted <- renamed_data %>%
    group_by_all() %>%
    summarise(freq = n(), .groups = 'drop')
  
  # Merge theoretically possible combinations with their frequency in real-world data
  merged_binarized_possible_real_without_total <- left_join(all_combinations_binarized_without_total_selectedPTSD, binarized_data_counted, by = symptom_cols_binarized_without_total)
  merged_binarized_possible_real_without_total$freq <- coalesce(merged_binarized_possible_real_without_total$freq, 0)
 
   return(merged_binarized_possible_real_without_total)
  }
  
  ### Applying function to our data
possible_combinedwith_real_binarized_without_total <- create_possible_combinedwith_real_binarized_without_total(data2)
```
#### 5.2 Description of the theoretically possible and actually occurring symptom combinations fulfilling diagnosis (without consideration of the total score criterion)
```{r  Description of the theoretically possible and actually occurring symptom combinations fulfilling diagnosis (without consideration of the total score criterion), echo=FALSE, message=FALSE}
  ### Create function for describing the resulting dataframe
  # Input data frame: resulting data frame of the function “create_possible_combinedwith_real_binarized_without_total”, i.e.: 21 columns (20 symptoms and one “freq” column). The symptoms are each binarized. As rows all theoretically possible combinations that fulfill the diagnostic criteria. The last column “freq” contains the number of presentations of this symptom combination in the real-world data.
description_merged_binarized_possible_real_without_total <- function(data){
  # Filter for only combinations that have actually occurred
  real_combinations <- data %>%
    filter(freq > 0)
  
  ## Description of theoretically and actually occurring combinations
  # How many individuals were diagnosed
  n_pat_diagnosed <- sum(data$freq)
  # How many combinations exist theoretically
  n_possible_combinations <- nrow(data)
  # How many combinations are represented in real-world data
  n_real_combinations <- nrow(real_combinations)
  # Percentage of combinations represented in real-world data
  percentage_real_possible <- round(n_real_combinations/ n_possible_combinations * 100, 2)
  
  # Display of some results in a matrix
  matrix_data1 <- c(
    n_pat_diagnosed,
    n_possible_combinations, 
    n_real_combinations, 
    percentage_real_possible,
    max(real_combinations$freq), 
    stats::median(data$freq),
    stats::median(real_combinations$freq))
  
  description_matrix1 <- matrix(matrix_data1, ncol = 1, nrow = 7)
  rownames(description_matrix1) <- c("Number of individuals fulfilling the diagnosis", 
				     "Number of theoretically possible combinations ", 
				     "Number of combinations that actually occur", 
                                     "Percentage of possible combinations that actually occur", 
                                     "How often the most common combination actually occurs",
                                     "Median frequency in relation to theoretically possible combinations",
                                     "Median frequency in relation to combinations actually occurring")
  description_matrix1[3,1] <- sprintf("%.2f%%", percentage_real_possible)
  
  # 10 most common combinations
  ten_most_common_real <- real_combinations %>% 
    dplyr::arrange(desc(freq)) %>% 
    dplyr::slice_head(n = 10)
  
  # Profiles (N, %) reported by less or equal than in 1% of the individuals only respecting combinations that actually occur
  one_percent_individuals = round(n_pat_diagnosed*0.01, 0)
  N_Profiles <- real_combinations %>% filter(freq<=one_percent_individuals) %>% nrow()
  Percentage_Profiles <- round((N_Profiles / nrow(real_combinations))*100, 3)
  
  # Number of individuals reporting one of the ten most common combinations
  N_individuals_ten_most_common <- real_combinations %>%
    arrange(desc(freq)) %>%
    slice_head(n = 10) %>% 
    summarise(sum(freq)) %>%
    pull()
  
  # Number of individuals reporting NOT one of the ten most common combinations
  N_individuals_not_ten_most_common <- real_combinations %>%
    arrange(desc(freq)) %>%
    slice_tail(n = -10) %>% 
    summarise(sum(freq)) %>%
    pull()
  
  # Number of individuals reporting 1% most common combinations
  one_percent_combinations <- round(nrow(real_combinations)*0.01, 0)
  N_individuals_one_perc_most_common <- real_combinations %>%
    arrange(desc(freq)) %>%
    slice_head(n = one_percent_combinations) %>% 
    summarise(sum(freq)) %>%
    pull()
  
  # Number of individuals reporting one of the 50% least common combinations
  fifty_percent_combinations <- round(nrow(real_combinations)*0.5, 0)
  N_individuals_fifty_perc_least_common <- real_combinations %>%
    arrange(desc(freq)) %>%
    slice_tail(n = fifty_percent_combinations) %>% 
    summarise(sum(freq)) %>%
    pull()
  
  # Display of some results in a matrix
  matrix_data2 <- c(
    N_Profiles, 
    Percentage_Profiles, 
    N_individuals_ten_most_common,
    N_individuals_not_ten_most_common, 
    N_individuals_one_perc_most_common,
    N_individuals_fifty_perc_least_common)
  
  description_matrix2 <- matrix(matrix_data2, ncol = 1, nrow = 6)
  rownames(description_matrix2) <- c("Number of Profiles reported by less or equal than in 1% of the individuals only respecting combinations that actually occur", 
                                     "Percentage of Profiles reported by less or equal than in 1% of the individuals only respecting combinations that actually occur", 
                                     "Number of individuals reporting one of the ten most common combinations", 
                                     "Number of individuals reporting NOT one of the ten most common combinations",
                                     "Number of individuals reporting 1% most common combinations",
                                     "Number of individuals reporting one of the 50% least common combinations")
  description_matrix2[2,1] <- sprintf("%.2f%%", Percentage_Profiles)
  
  ## Description of single symptoms
  symptom_cols <- setdiff(colnames(data), "freq")
  n_symptoms <- length(symptom_cols)
  symp_prevalence_comb <- numeric(n_symptoms)
  symp_absolute_comb <- numeric(n_symptoms)
  symp_prevalence_pat <- numeric(n_symptoms)
  symp_absolute_pat <- numeric(n_symptoms)
  
  for (i in 1:n_symptoms) {
    symp_absolute_comb[i] <- sum(real_combinations[,i] == 1)   # Absolute prevalence of a symptom in different actually occurring combinations
    symp_prevalence_comb[i] <- (symp_absolute_comb[i]/nrow(real_combinations)) * 100 # Percentage prevalence of a symptom in different actually occurring combinations
    symp_absolute_pat[i] <- sum((real_combinations[,i] == 1)* real_combinations$freq) # Absolute prevalence of a symptom in patients
    symp_prevalence_pat[i] <- (symp_absolute_pat[i]/n_pat_diagnosed) * 100 # Percentage prevalence of a symptom in patients
  }
  
  # Display of some results in a matrix
  matrix_data3 <- c(
    symp_absolute_comb, 
    symp_prevalence_comb, 
    symp_absolute_pat, 
    symp_prevalence_pat
  )
  
  description_matrix3 <- matrix(matrix_data3, ncol = 4, byrow = FALSE)
  rownames(description_matrix3) <- colnames(data) [1:(ncol(data)-1)]     
  colnames(description_matrix3) <- c("Number of combinations actually occurring that include this symptom", 
                                     "Percentage of combinations actually occurring that include this symptom",
                                     "Number of patients in the data set showing this symptom",
                                     "Percentage of patients in the data set showing this symptom") 
  description_matrix3[,2] <- sprintf("%.2f%%", symp_prevalence_comb)
  description_matrix3[,4] <- sprintf("%.2f%%", symp_prevalence_pat)
  
  list(
    "Matrix1" = description_matrix1,
    "Matrix2" = description_matrix2,
    "Matrix3" = description_matrix3
  )
}

   ### Applying function to our data
description_merged_binarized_possible_real_without_total(possible_combinedwith_real_binarized_without_total)
```
#### 5.3. Determine the change resulting from the omission of one or more symptoms
##### 5.3.1. Function for neglecting one or more symptoms and then theoretically possible combinations of binarized symptoms fulfilling diagnosis (without considering the total score criterion) and their frequency in real-world data
```{r  Neglecting one or more symptoms and then theoretically possible combinations of binarized symptoms fulfilling diagnosis (without considering the total score criterion) and their frequency in real-world data, echo=FALSE, message=FALSE}
  ### Create function for deleting one or more symptoms. Then theoretically possible combinations of binarized symptoms fulfilling diagnosis (without consideration of the total score criterion) and their frequencies in real-world data
  # Input: data: real-world dataset fulfilling diagnostic criteria, 20 columns with binarized symptoms, symptoms_to_delete: number of the symptom(s) you want to delete
create_possible_combined_with_real_binarized_without_total_symp_del <- function(data, symptoms_to_delete) {
  # Convert single number to vector if necessary
  if(!is.vector(symptoms_to_delete)) {
    symptoms_to_delete <- c(symptoms_to_delete)
  }
  
  # Checks input
  if (!all(sapply(symptoms_to_delete, is.numeric)) || 
      any(symptoms_to_delete < 1) || 
      any(symptoms_to_delete > 20) ||
      length(unique(symptoms_to_delete)) != length(symptoms_to_delete)) {
    stop("symptoms_to_delete must be unique numbers between 1 and 20")
  }
  
  # Create vector of all symptom numbers excluding the ones to delete
  remaining_symptoms <- setdiff(1:20, symptoms_to_delete)
  
  # Ensure we have enough symptoms left for the criteria
  if(length(remaining_symptoms) < 6) {
    stop("Cannot delete that many symptoms - need enough symptoms to maintain criteria requirements")
  }
  
  # Create new column names excluding the deleted symptoms
  symptom_cols_binarized_without_total <- paste0("symptom_", remaining_symptoms)
  
  ## Rename the columns in the input data and remove deleted symptoms
  renamed_data <- data %>%
    rename_with(~ paste0("symptom_", 1:20)) %>%
    select(-paste0("symptom_", symptoms_to_delete))  # Remove the deleted symptoms
  
  # Generate new dataframe with all possible combinations
  symptom_values_binarized_without_total <- 0:1  # 0: absent, 1: present
  all_combinations_binarized_without_total <- expand.grid(
    lapply(1:length(remaining_symptoms), function(i) symptom_values_binarized_without_total)
  )
  colnames(all_combinations_binarized_without_total) <- symptom_cols_binarized_without_total
  
  # Helper function to count symptoms in a range excluding deleted symptoms
  count_in_range <- function(df, start, end) {
    range_symptoms <- intersect(remaining_symptoms, start:end) # Which remaining_symptoms fall in the range
    if(length(range_symptoms) > 0) {
      cols <- paste0("symptom_", range_symptoms)
      return(rowSums(df[, cols, drop = FALSE]))
    }
    return(0)
  }
  
  # Helper function to calculate required symptoms based on remaining symptoms in range 
  #(original_required: how many symptoms in this range must be fulfilled in the original criteria?)
  get_required_symptoms <- function(start, end, original_required) {
    range_symptoms <- intersect(remaining_symptoms, start:end)
    # Adjust required symptoms proportionally, rounding up
    new_required <- ceiling(original_required * length(range_symptoms) / length(start:end))
    return(new_required)
  }
  
  # Calculate adjusted requirements based on remaining symptoms
  req_1_5 <- get_required_symptoms(1, 5, 1)
  req_6_7 <- get_required_symptoms(6, 7, 1)
  req_8_14 <- get_required_symptoms(8, 14, 2)
  req_15_20 <- get_required_symptoms(15, 20, 2)
  
  # Define criteria for filtering with adjusted ranges and requirements
  criteria_1 <- (
    count_in_range(all_combinations_binarized_without_total, 1, 5) >= req_1_5 &
      count_in_range(all_combinations_binarized_without_total, 6, 7) >= req_6_7 &
      count_in_range(all_combinations_binarized_without_total, 8, 14) >= req_8_14 &
      count_in_range(all_combinations_binarized_without_total, 15, 20) >= req_15_20
  )
  
  # Filter combinations based on criteria
  all_combinations_binarized_without_total_selectedPTSD <- filter(
    all_combinations_binarized_without_total, 
    criteria_1 == TRUE
  )
  
  # Count frequency of binarized profiles in real-world data
  binarized_data_counted <- renamed_data %>%
    group_by_all() %>%
    summarise(freq = n(), .groups = 'drop')
  
  # Merge theoretically possible combinations with their frequency in real-world data
  merged_binarized_possible_real_without_total <- left_join(
    all_combinations_binarized_without_total_selectedPTSD, 
    binarized_data_counted, 
    by = symptom_cols_binarized_without_total
  )
  merged_binarized_possible_real_without_total$freq <- coalesce(
    merged_binarized_possible_real_without_total$freq, 
    0
  )
  
  return(merged_binarized_possible_real_without_total)
}
```
##### 5.3.2. Function for describing the resulting data frame with one or more deleted symptoms and for comparing this data set with the data set that takes all symptoms into account. 
```{r Function for describing the resulting data frame with one or more deleted symptoms and for comparing this data set with the data set that takes all symptoms into account., echo=FALSE, message=FALSE}
  ### Create function for describing the resulting data frame with one or more deleted symptoms and for comparing this data set with the data set that takes all symptoms into account. 
  # Input data frames: resulting data frames of 1. the function “create_possible_combinedwith_real_binarized_without_total” and 2. "create_possible_combinedwith_real_binarized_without_total_symp_del". The first data set comprises 21 columns (20 symptoms and one column “freq”). The second minus the deleted symptoms. The symptoms are each binarized. The rows are all theoretically possible combinations that fulfill the diagnostic criteria. The last column “freq” contains the number of presentations of this symptom combination in the real data.
description_and_comparison_symp_del_merged_binarized_possible_real_without_total <- function(data_all_symptoms, data_deleted_symptoms) {
  # Get description of both datasets using the original function
  desc_all <- description_merged_binarized_possible_real_without_total(data_all_symptoms)
  desc_deleted <- description_merged_binarized_possible_real_without_total(data_deleted_symptoms)
  
  # Helper function to safely convert percentage strings to numeric
  safe_percent_convert <- function(x) {
    if(is.character(x)) {
      as.numeric(gsub("[^0-9.-]", "", x))
    } else {
      as.numeric(x)
    }
  }
  
  ## Create comparison matrices
  # Matrix 1 comparison - Basic statistics
  comparison_matrix1 <- matrix(
    c(desc_all$Matrix1[,1], desc_deleted$Matrix1[,1], 
      mapply(function(all_val, del_val) {
        if(grepl("%", all_val)) {
          sprintf("%.2f%%", safe_percent_convert(del_val) - safe_percent_convert(all_val))
        } else {
          as.numeric(del_val) - as.numeric(all_val)
        }
      }, desc_all$Matrix1[,1], desc_deleted$Matrix1[,1])),
    ncol = 3
  )
  rownames(comparison_matrix1) <- rownames(desc_all$Matrix1)
  colnames(comparison_matrix1) <- c("All Symptoms", "Deleted Symptoms", "Difference")
  
  # Matrix 2 comparison - Statistic on the combinations
  comparison_matrix2 <- matrix(
    c(desc_all$Matrix2[,1], desc_deleted$Matrix2[,1],
      mapply(function(all_val, del_val) {
        if(grepl("%", all_val)) {
          sprintf("%.2f%%", safe_percent_convert(del_val) - safe_percent_convert(all_val))
        } else {
          as.numeric(del_val) - as.numeric(all_val)
        }
      }, desc_all$Matrix2[,1], desc_deleted$Matrix2[,1])),
    ncol = 3
  )
  rownames(comparison_matrix2) <- rownames(desc_all$Matrix2)
  colnames(comparison_matrix2) <- c("All Symptoms", "Deleted Symptoms", "Difference")
  
  # Matrix 3 comparison
  common_symptoms <- intersect(
    rownames(desc_all$Matrix3),
    rownames(desc_deleted$Matrix3)
  )
  
  comparison_matrix3 <- matrix(
    nrow = length(common_symptoms),
    ncol = 12
  )
  rownames(comparison_matrix3) <- common_symptoms
  colnames(comparison_matrix3) <- c(
    "N combinations actually occurring that include this symptom (All)", 
    "N combinations actually occurring that include this symptom(Deleted)", 
    "N combinations actually occurring that include this symptom Difference",
    "% combinations actually occurring that include this symptom (All)", 
    "% combinations actually occurring that include this symptom (Deleted)", 
    "% combinations actually occurring that include this symptom Difference",
    "N patients in the data set showing this symptom (All)", 
    "N patients in the data set showing this symptom (Deleted)",
    "N patients in the data set showing this symptom Difference",
    "% patients in the data set showing this symptom (All)",
    "% patients in the data set showing this symptom (Deleted)",
    "% patients in the data set showing this symptom Difference"
  )
  
  for(symptom in common_symptoms) {
    row_all <- desc_all$Matrix3[symptom,]
    row_deleted <- desc_deleted$Matrix3[symptom,]
    
    # Convert values safely
    values <- c(
      safe_percent_convert(row_all[1]), 
      safe_percent_convert(row_deleted[1]),
      safe_percent_convert(row_deleted[1]) - safe_percent_convert(row_all[1]),
      safe_percent_convert(row_all[2]),
      safe_percent_convert(row_deleted[2]),
      safe_percent_convert(row_deleted[2]) - safe_percent_convert(row_all[2]),
      safe_percent_convert(row_all[3]), 
      safe_percent_convert(row_deleted[3]),
      safe_percent_convert(row_deleted[3]) - safe_percent_convert(row_all[3]),
      safe_percent_convert(row_all[4]),
      safe_percent_convert(row_deleted[4]),
      safe_percent_convert(row_deleted[4]) - safe_percent_convert(row_all[4])
    )
    
    comparison_matrix3[symptom,] <- values
  }
  
  # Format percentages in matrices (only for percentage columns)
  percent_cols <- c(4,5,6,10,11,12)
  for(col in percent_cols) {
    comparison_matrix3[,col] <- sprintf("%.2f%%", as.numeric(comparison_matrix3[,col]))
  }
  
  ## Return all results in a list
  list(
    "Comparison_Matrix1" = comparison_matrix1,
    "Comparison_Matrix2" = comparison_matrix2,
    "Comparison_Matrix3" = comparison_matrix3
  )
}

```
##### 5.3.3. Application of the generated functions  
###### 5.3.3.1. Deletion of symptom 1
```{r Deletion of symptom 3, echo=FALSE, message=FALSE}
  ## Applying function “create_possible_combinedwith_real_binarized_without_total_symp_del“
possible_combinedwith_real_binarized_without_total_symp_del_1 <- create_possible_combined_with_real_binarized_without_total_symp_del(data2, 1)

  ## Applying function “description_and_comparison_symp_del_merged_binarized_possible_real_without_total“
description_and_comparison_symp_del_merged_binarized_possible_real_without_total(possible_combinedwith_real_binarized_without_total, possible_combinedwith_real_binarized_without_total_symp_del_1)
```
## 6. Theoretically possible combinations of nonbinarized symptoms fulfilling diagnosis and their frequency in real-world data
```{r Theoretically possible combinations of nonbinarized symptoms fulfilling diagnosis and their frequency in real-world data, echo=FALSE, message=FALSE}
# ### DOES NOT WORK, BECAUSE MEMORY LIMIT IS REACHED, I am searching for alternative options (the use of chunks to handle the memory limit does not work either)
# ### Function for Dataframe with all possible combinations of nonbinarized symptoms fulfilling PTSD and their frequencies in real-world data
# create_possible_combinedwith_real_nonbinarized <- function(data) {
#   # Generate new dataframe (all_combinations_nonbinarized) with all possible combinations of symptom values
#   symptom_cols_nonbinarized <- paste0("symptom_", 1:20)
#   symptom_values_nonbinarized <- 0:4
#   all_combinations_nonbinarized <- expand.grid(lapply(1:20, function(i) symptom_values_nonbinarized))
#   colnames(all_combinations_nonbinarized) <- symptom_cols_nonbinarized
# 
#   # Calculate the total score for each row
#   all_combinations_nonbinarized$total <- rowSums(all_combinations_nonbinarized)
# 
#   # Define criteria for filtering
#   criteria_1 <- all_combinations_nonbinarized$total >= 33
#   criteria_2 <- (
#     rowSums(all_combinations_nonbinarized[, 1:5] >= 2) >= 1 &
#       rowSums(all_combinations_nonbinarized[, 6:7] >= 2) >= 1 &
#       rowSums(all_combinations_nonbinarized[, 8:14] >= 2) >= 2 &
#       rowSums(all_combinations_nonbinarized[, 15:20] >= 2) >= 2
#   )
# 
#   # Filter combinations based on criteria
#   all_combinations_nonbinarized_selectedPTSD <- all_combinations_nonbinarized[criteria_1 | criteria_2, ]
# 
#   # Count frequency of nonbinarized profiles in real-world data
#     # First ensure only using symptom columns
#     simulated_ptsd_total_selected <- data[, symptom_cols_nonbinarized]
# 
#     simulated_ptsd_total_selected_counted <- simulated_ptsd_total_selected %>%
#     group_by_all() %>%
#     summarise(freq = n(), .groups = 'drop')
# 
#   # Merge with the frequency data
#   merged_possible_real_nonbinarized <- left_join(all_combinations_nonbinarized_selectedPTSD, simulated_ptsd_total_selected_counted, by = symptom_cols_nonbinarized)
#   merged_possible_real_nonbinarized$freq <- coalesce(merged_possible_real_nonbinarized$freq, 0)
# 
#   return(merged_possible_real_nonbinarized)
# }
# 
#   ## Applying function to our data
#   possible_combinedwith_real_nonbinarized <- create_possible_combinedwith_real_nonbinarized(simulated_ptsd_total_selected)

```
## x. Session Info

```{r Session Info, echo=FALSE, message=FALSE}
sessionInfo()
```
