---
title: "Assess Simulated PTSD Symptoms"
author:
- "<h5 style=\"font-style:italic\"> Laura Weidmann"
- "<h5 style=\"font-style:italic\"> Tobias R. Spiller"
date: "<h5 style=\"font-style:roman\"> `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
  pdf_document:
    toc: true
    toc_depth: '5'
subtitle: Version 0.0.2
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)
```

```{r Load Libraries, message=FALSE, warning=TRUE, include=FALSE}
# Data handling
library(tidyverse)
library(readr)
library(data.table)

# Demographics
library(table1)
library(gtsummary)

# Analysis
#library("devtools")
#devtools::install_github("orduek/PsychPower")
library(PsychPower)
```

## 1. Import and prepare data

```{r PCL import data, message=FALSE, warning=TRUE, include=FALSE}
# Import data
simulated_ptsd_data <- read_csv("Data/simulated_skewed_ptsd_data.csv")
```

## 2. Sample descriptive

#### 2.1 Total score & Diagnosis

```{r PCL Table 1, echo=FALSE, message=FALSE}
# Create PCL total score
simulated_ptsd_total <- simulated_ptsd_data %>% 
  mutate(total = rowSums(select(., symptom_1:symptom_20)))

# # Create Diagnosis
#simulated_ptsd_total <- simulated_ptsd_total %>%
  #mutate(
    #PTSD_Diagnosis = ifelse(
      #(symptom_1 > 2 | symptom_2 > 2 | symptom_3 > 2 | symptom_4 > 2 | symptom_5 > 2) &
      #(symptom_6 > 2 | symptom_7 > 2) &
      #(symptom_8 > 2 | symptom_9 > 2 | symptom_10 > 2 | symptom_11 > 2 | symptom_12 > 2 | symptom_13 > 2 | symptom_14 #> 2) &
      #(symptom_15 > 2 | symptom_16 > 2 | symptom_17 > 2 | symptom_18 > 2 | symptom_19 > 2 | symptom_20 > 2),
      #TRUE, FALSE
    #)
  #)

# Create function that implements PTSD diagnostic criteria
  create_ptsd_diagnosis <- function(data) {
    
    # Create new column initially set to FALSE
    data$PTSD_Diagnosis <- FALSE
    
    # Check total score criterion
    total_criterion <- data$total >= 33
    
    # Check symptom criteria
    criterion_B <- rowSums(data[, paste0("symptom_", 1:5)] >= 2) >= 1  # At least 1 from symptoms 1-5
    criterion_C <- rowSums(data[, paste0("symptom_", 6:7)] >= 2) >= 1  # At least 1 from symptoms 6-7
    criterion_D <- rowSums(data[, paste0("symptom_", 8:14)] >= 2) >= 2 # At least 2 from symptoms 8-14
    criterion_E <- rowSums(data[, paste0("symptom_", 15:20)] >= 2) >= 2 # At least 2 from symptoms 15-20
    
    # All symptom criteria met
    all_symptom_criteria <- criterion_B & criterion_C & criterion_D & criterion_E
    
    # Combined criteria: Either total >= 33 OR all symptom criteria met OR both conditions together
    data$PTSD_Diagnosis <- total_criterion | 
      all_symptom_criteria |
      (total_criterion & all_symptom_criteria)
    
    return(data)
  }

#Apply the function to our dataframe
  simulated_ptsd_total <- create_ptsd_diagnosis (simulated_ptsd_total)

#Summarize
simulated_ptsd_total %>%
  summarise(
    mean_total = mean(total),
    sd_total = sd(total),
    n_diagnosed = sum(PTSD_Diagnosis)
  )
```

#### 2.2 Cronbach's alpha

Selected sample

```{r PCL cronbach, echo=FALSE, message=FALSE}
cronbach <- psych::alpha(subset(simulated_ptsd_total, select = (-total)))
cronbach$total
```

### 2.3 Summary of items & histogram

```{r PCL summary, echo=FALSE, message=FALSE}
summary(simulated_ptsd_total)
hist(simulated_ptsd_total$total)
```

## 3. Binarizing & create datax

Symptoms rated 2 or 3 will be considered as present. [not shown]

```{r PCL Binarizing, echo=FALSE, message=FALSE, warning=FALSE}
#selecting only the "TRUE" Diagnosis
simulated_ptsd_total_selected <- simulated_ptsd_total %>%
  filter(PTSD_Diagnosis=="TRUE")

## set cut-off
cut_off <- 1 #will be used with <= // 

##rename variables to "Q1-QN" for binarization
n_cols <- ncol(simulated_ptsd_total)
simulated_ptsd_selected_rename <- simulated_ptsd_total_selected %>%
  rename_with(~ paste0("Q", seq_len(n_cols - 2)), 1:(n_cols -2 ))

## Binarize
data_binarized <- simulated_ptsd_selected_rename

for (i in 1:(ncol(data_binarized)-2)){
  orig <- paste("q", i, sep = "")
  bin <- paste("Q", i, sep = "")
  data_binarized[orig] <- dplyr::case_when(data_binarized[bin]<= cut_off ~ 0, data_binarized[bin]>cut_off ~ 1)  #0 = "Symptom absent", 1 = "Symptom present"
}

# Create new data frame
data2 <- data_binarized %>% 
  select(PTSD_Diagnosis:ncol(data_binarized)) %>% 
  select(-PTSD_Diagnosis) %>% 
  tibble()

## Count frequency of profiles
data2_counted <- plyr::count(data2[, ]) 

# Create sum score of endorsed symptoms
data2_counted <- data2_counted %>% 
  mutate(total_bin = rowSums(data2_counted)-freq)

## Create datax 
# Create full dataset
datax <- dplyr::left_join(data_binarized, data2_counted)


## Create dataframes for PsychPower Package 
data_PTSD_PsychPower <- simulated_ptsd_total_selected %>%  
  	select (symptom_1:symptom_20)
#binarize
data_binarized_PsychPower <- PsychPower::binarize(data_PTSD_PsychPower, cut_off = 1)
#frequency
data_frequency_PsychPower <- PsychPower::pheno_frequency(
  data_binarized_PsychPower, target_columns = tidyselect::starts_with("v_bin"))
```

## 4. Description of combinations

Most common combination, its frequency & median frequency

```{r PCL Describe Number, echo=FALSE, message=FALSE}
desc_pheno_PsychPower <- PsychPower::describe_pheno(data_frequency_PsychPower, frequency = "freq") 
desc_pheno_PsychPower
```

10 Most common Phenotype

```{r PCL Describe 10 most, echo=FALSE, message=FALSE}
PsychPower::common_pheno(data_frequency_PsychPower, frequency = "freq", n_phenotypes = 10)
```

Profiles (N, %) reported by less or equal than in 1% of the individuals

```{r PCL Having less then 1 percent, echo=FALSE, message=FALSE}
one_percent_sample = round(nrow(data_binarized_PsychPower)*0.01, 0)
data2_counted %>% filter(freq<=one_percent_sample) %>% nrow()

round((data2_counted %>% filter(freq<=one_percent_sample) %>% nrow()) / nrow(data2_counted), 3)
```

Individuals reporting one of the ten most common combinations

```{r PCL Having 10 most, echo=FALSE, message=FALSE}
data2_counted <- data2_counted %>% 
  arrange(desc(freq))
data2_counted[1:10,] %>% summarise(sum(freq))
```

Individuals reporting NOT one of the ten most common combinations

```{r PCL  Not Having 10 most, echo=FALSE, message=FALSE}
data2_counted[11:nrow(data2_counted),] %>% summarise(sum(freq))
```

Individuals reporting 1% most common combinations

```{r PCL Describe 1 percent, echo=FALSE, message=FALSE}
one_percent =  round(nrow(data2_counted)*0.01, 0)

data2_counted[1:one_percent,] %>% summarise(sum(freq))
```

Individuals reporting one of the 50% least common combinations

```{r PCL Describe 50 percent, echo=FALSE, message=FALSE}
fifty_percent =  round(nrow(data2_counted)*0.5, 0)

data2_counted[fifty_percent:nrow(data2_counted),] %>% summarise(sum(freq))

```

Plot 100 most common profiles

```{r Plot 100 most common, echo=FALSE, message=FALSE}
fig1 <- PsychPower::plot_pheno(data_frequency_PsychPower, frequency = "freq",
                               n_phenotypes = 100, color = "grey26")
fig1
```

## 5. Analysis of the possible combinations of binarized symptoms that fulfill the diagnosis (without consideration of the total score criterion)
#### 5.1. Theoretically possible combinations of binarized symptoms fulfilling diagnosis (without consideration of the total score criterion) and their frequency in real-world data
```{r Theoretically possible combinations of binarized symptoms fulfilling diagnosis (without consideration of the total score criterion) and their frequency in real-world data, echo=FALSE, message=FALSE}
 ### Create function
  # Input-data: real-world dataset fulfilling diagnostic criteria, 20 columns with binarized symptoms
create_possible_combinedwith_real_binarized_without_total <- function(data) {
  # First rename the columns in the input data (real-world) to match the expected format
  renamed_data <- data %>%
    rename_with(~ paste0("symptom_", 1:20))
  
  # Generate new dataframe (all_combinations_binarized_without_total) with all possible combinations of symptom combination (binarized) without respecting the criteria of totalscore
  symptom_cols_binarized_without_total <- paste0("symptom_", 1:20) # Total number of symptoms
  symptom_values_binarized_without_total <- 0:1  # 0: absent, 1: present
  all_combinations_binarized_without_total <- expand.grid(lapply(1:20, function(i) symptom_values_binarized_without_total))
  colnames(all_combinations_binarized_without_total) <- symptom_cols_binarized_without_total
  
  # Define criteria for filtering
  criteria_1 <- (
    rowSums(all_combinations_binarized_without_total[, 1:5]) >= 1 &          # At least 1 symptom from 1-5
      rowSums(all_combinations_binarized_without_total[, 6:7]) >= 1 &          # At least 1 symptom from 6-7
      rowSums(all_combinations_binarized_without_total[, 8:14]) >= 2 &         # At least 2 symptoms from 8-14
      rowSums(all_combinations_binarized_without_total[, 15:20]) >= 2          # At least 2 symptoms from 15-20
  )
  
  # Filter combinations based on criteria
  all_combinations_binarized_without_total_selectedPTSD <- filter(all_combinations_binarized_without_total, criteria_1==TRUE)
  
  # Count frequency of binarized profiles in real-world data
  binarized_data_counted <- renamed_data %>%
    group_by_all() %>%
    summarise(freq = n(), .groups = 'drop')
  
  # Merge theoretically possible combinations with their frequency in real-world data
  merged_binarized_possible_real_without_total <- left_join(all_combinations_binarized_without_total_selectedPTSD, binarized_data_counted, by = symptom_cols_binarized_without_total)
  merged_binarized_possible_real_without_total$freq <- coalesce(merged_binarized_possible_real_without_total$freq, 0)
 
   return(merged_binarized_possible_real_without_total)
  }
  
  ### Applying function to our data
possible_combinedwith_real_binarized_without_total <- create_possible_combinedwith_real_binarized_without_total(data2)
```
#### 5.2 Description of the theoretically possible and actually occurring symptom combinations fulfilling diagnosis (without consideration of the total score criterion)
```{r  Description of the theoretically possible and actually occurring symptom combinations fulfilling diagnosis (without consideration of the total score criterion), echo=FALSE, message=FALSE}
  ### Create function for describing the resulting dataframe
  # Input data frame: resulting data frame of the function “create_possible_combinedwith_real_binarized_without_total”, i.e.: 21 columns (20 symptoms and one “freq” column). The symptoms are each binarized. As rows all theoretically possible combinations that fulfill the diagnostic criteria. The last column “freq” contains the number of presentations of this symptom combination in the real-world data.
description_merged_binarized_possible_real_without_total <- function(data){
  # Filter for only combinations that have actually occurred
  real_combinations <- data %>%
    filter(freq > 0)
  
  ## Description of theoretically and actually occurring combinations
  # How many individuals were diagnosed
  n_pat_diagnosed <- sum(data$freq)
  # How many combinations exist theoretically
  n_possible_combinations <- nrow(data)
  # How many combinations are represented in real-world data
  n_real_combinations <- nrow(real_combinations)
  # Percentage of combinations represented in real-world data
  percentage_real_possible <- round(n_real_combinations/ n_possible_combinations * 100, 2)
  
  # Display of some results in a matrix
  matrix_data1 <- c(
    n_pat_diagnosed,
    n_possible_combinations, 
    n_real_combinations, 
    percentage_real_possible,
    max(real_combinations$freq), 
    stats::median(data$freq),
    stats::median(real_combinations$freq))
  
  description_matrix1 <- matrix(matrix_data1, ncol = 1, nrow = 7)
  rownames(description_matrix1) <- c("Number of individuals fulfilling the diagnosis", 
				     "Number of theoretically possible combinations ", 
				     "Number of combinations that actually occur", 
                                     "Percentage of possible combinations that actually occur", 
                                     "How often the most common combination actually occurs",
                                     "Median frequency in relation to theoretically possible combinations",
                                     "Median frequency in relation to combinations actually occurring")
  description_matrix1[4,1] <- sprintf("%.2f%%", percentage_real_possible)
  
  # 10 most common combinations
  ten_most_common_real <- real_combinations %>% 
    dplyr::arrange(desc(freq)) %>% 
    dplyr::slice_head(n = 10)
  
  # Profiles (N, %) reported by less or equal than in 1% of the individuals only respecting combinations that actually occur
  one_percent_individuals = round(n_pat_diagnosed*0.01, 0)
  N_Profiles <- real_combinations %>% filter(freq<=one_percent_individuals) %>% nrow()
  Percentage_Profiles <- round((N_Profiles / nrow(real_combinations))*100, 3)
  
  # Number of individuals reporting one of the ten most common combinations
  N_individuals_ten_most_common <- real_combinations %>%
    arrange(desc(freq)) %>%
    slice_head(n = 10) %>% 
    summarise(sum(freq)) %>%
    pull()
  
  # Number of individuals reporting NOT one of the ten most common combinations
  N_individuals_not_ten_most_common <- real_combinations %>%
    arrange(desc(freq)) %>%
    slice_tail(n = -10) %>% 
    summarise(sum(freq)) %>%
    pull()
  
  # Number of individuals reporting 1% most common combinations
  one_percent_combinations <- round(nrow(real_combinations)*0.01, 0)
  N_individuals_one_perc_most_common <- real_combinations %>%
    arrange(desc(freq)) %>%
    slice_head(n = one_percent_combinations) %>% 
    summarise(sum(freq)) %>%
    pull()
  
  # Number of individuals reporting one of the 50% least common combinations
  fifty_percent_combinations <- round(nrow(real_combinations)*0.5, 0)
  N_individuals_fifty_perc_least_common <- real_combinations %>%
    arrange(desc(freq)) %>%
    slice_tail(n = fifty_percent_combinations) %>% 
    summarise(sum(freq)) %>%
    pull()
  
  # Display of some results in a matrix
  matrix_data2 <- c(
    N_Profiles, 
    Percentage_Profiles, 
    N_individuals_ten_most_common,
    N_individuals_not_ten_most_common, 
    N_individuals_one_perc_most_common,
    N_individuals_fifty_perc_least_common)
  
  description_matrix2 <- matrix(matrix_data2, ncol = 1, nrow = 6)
  rownames(description_matrix2) <- c("Number of Profiles reported by less or equal than in 1% of the individuals only respecting combinations that actually occur", 
                                     "Percentage of Profiles reported by less or equal than in 1% of the individuals only respecting combinations that actually occur", 
                                     "Number of individuals reporting one of the ten most common combinations", 
                                     "Number of individuals reporting NOT one of the ten most common combinations",
                                     "Number of individuals reporting 1% most common combinations",
                                     "Number of individuals reporting one of the 50% least common combinations")
  description_matrix2[2,1] <- sprintf("%.2f%%", Percentage_Profiles)
  
  ## Description of single symptoms
  symptom_cols <- setdiff(colnames(data), "freq")
  n_symptoms <- length(symptom_cols)
  symp_prevalence_comb <- numeric(n_symptoms)
  symp_absolute_comb <- numeric(n_symptoms)
  symp_prevalence_pat <- numeric(n_symptoms)
  symp_absolute_pat <- numeric(n_symptoms)
  
  for (i in 1:n_symptoms) {
    symp_absolute_comb[i] <- sum(real_combinations[,i] == 1)   # Absolute prevalence of a symptom in different actually occurring combinations
    symp_prevalence_comb[i] <- (symp_absolute_comb[i]/nrow(real_combinations)) * 100 # Percentage prevalence of a symptom in different actually occurring combinations
    symp_absolute_pat[i] <- sum((real_combinations[,i] == 1)* real_combinations$freq) # Absolute prevalence of a symptom in patients
    symp_prevalence_pat[i] <- (symp_absolute_pat[i]/n_pat_diagnosed) * 100 # Percentage prevalence of a symptom in patients
  }
  
  # Display of some results in a matrix
  matrix_data3 <- c(
    symp_absolute_comb, 
    symp_prevalence_comb, 
    symp_absolute_pat, 
    symp_prevalence_pat
  )
  
  description_matrix3 <- matrix(matrix_data3, ncol = 4, byrow = FALSE)
  rownames(description_matrix3) <- colnames(data) [1:(ncol(data)-1)]     
  colnames(description_matrix3) <- c("Number of combinations actually occurring that include this symptom", 
                                     "Percentage of combinations actually occurring that include this symptom",
                                     "Number of patients in the data set showing this symptom",
                                     "Percentage of patients in the data set showing this symptom") 
  description_matrix3[,2] <- sprintf("%.2f%%", symp_prevalence_comb)
  description_matrix3[,4] <- sprintf("%.2f%%", symp_prevalence_pat)
  
  list(
    "Matrix1" = description_matrix1,
    "Matrix2" = description_matrix2,
    "Matrix3" = description_matrix3
  )
}

   ### Applying function to our data
description_merged_binarized_possible_real_without_total(possible_combinedwith_real_binarized_without_total)
```
#### 5.3. Determine the change resulting from the omission of one or more symptoms
##### 5.3.1. Function for neglecting one or more symptoms and then theoretically possible combinations of binarized symptoms fulfilling diagnosis (without considering the total score criterion) and their frequency in real-world data
```{r  Neglecting one or more symptoms and then theoretically possible combinations of binarized symptoms fulfilling diagnosis (without considering the total score criterion) and their frequency in real-world data, echo=FALSE, message=FALSE}
  ### Create function for deleting one or more symptoms. Then theoretically possible combinations of binarized symptoms fulfilling diagnosis (without consideration of the total score criterion) and their frequencies in real-world data
  # Input: data: real-world dataset fulfilling diagnostic criteria, 20 columns with binarized symptoms, symptoms_to_delete: number of the symptom(s) you want to delete
create_possible_combined_with_real_binarized_without_total_symp_del <- function(data, symptoms_to_delete) {
  # Convert single number to vector if necessary
  if(!is.vector(symptoms_to_delete)) {
    symptoms_to_delete <- c(symptoms_to_delete)
  }
  
  # Checks input
  if (!all(sapply(symptoms_to_delete, is.numeric)) || 
      any(symptoms_to_delete < 1) || 
      any(symptoms_to_delete > 20) ||
      length(unique(symptoms_to_delete)) != length(symptoms_to_delete)) {
    stop("symptoms_to_delete must be unique numbers between 1 and 20")
  }
  
  # Create vector of all symptom numbers excluding the ones to delete
  remaining_symptoms <- setdiff(1:20, symptoms_to_delete)
  
  # Ensure we have enough symptoms left for the criteria
  if(length(remaining_symptoms) < 6) {
    stop("Cannot delete that many symptoms - need enough symptoms to maintain criteria requirements")
  }
  
  # Create new column names excluding the deleted symptoms
  symptom_cols_binarized_without_total <- paste0("symptom_", remaining_symptoms)
  
  ## Rename the columns in the input data and remove deleted symptoms
  renamed_data <- data %>%
    rename_with(~ paste0("symptom_", 1:20)) %>%
    select(-paste0("symptom_", symptoms_to_delete))  # Remove the deleted symptoms
  
  # Generate new dataframe with all possible combinations
  symptom_values_binarized_without_total <- 0:1  # 0: absent, 1: present
  all_combinations_binarized_without_total <- expand.grid(
    lapply(1:length(remaining_symptoms), function(i) symptom_values_binarized_without_total)
  )
  colnames(all_combinations_binarized_without_total) <- symptom_cols_binarized_without_total
  
  # Helper function to count symptoms in a range excluding deleted symptoms
  count_in_range <- function(df, start, end) {
    range_symptoms <- intersect(remaining_symptoms, start:end) # Which remaining_symptoms fall in the range
    if(length(range_symptoms) > 0) {
      cols <- paste0("symptom_", range_symptoms)
      return(rowSums(df[, cols, drop = FALSE]))
    }
    return(0)
  }
  
  # Helper function to calculate required symptoms based on remaining symptoms in range 
  #(original_required: how many symptoms in this range must be fulfilled in the original criteria?)
  get_required_symptoms <- function(start, end, original_required) {
    range_symptoms <- intersect(remaining_symptoms, start:end)
    if(length(range_symptoms) == 0) {
      stop(sprintf("Cannot delete all symptoms from range %d-%d as at least %d symptom(s) from this range are required for diagnosis", 
                  start, end, original_required))
    }
    # Adjust required symptoms proportionally, rounding up
    new_required <- ceiling(original_required * length(range_symptoms) / length(start:end))
    return(new_required)
  }
  
  # Calculate adjusted requirements based on remaining symptoms
  req_1_5 <- get_required_symptoms(1, 5, 1)
  req_6_7 <- get_required_symptoms(6, 7, 1)
  req_8_14 <- get_required_symptoms(8, 14, 2)
  req_15_20 <- get_required_symptoms(15, 20, 2)
  
  # Define criteria for filtering with adjusted ranges and requirements
  criteria_1 <- (
    count_in_range(all_combinations_binarized_without_total, 1, 5) >= req_1_5 &
      count_in_range(all_combinations_binarized_without_total, 6, 7) >= req_6_7 &
      count_in_range(all_combinations_binarized_without_total, 8, 14) >= req_8_14 &
      count_in_range(all_combinations_binarized_without_total, 15, 20) >= req_15_20
  )
  
  # Filter combinations based on criteria
  all_combinations_binarized_without_total_selectedPTSD <- filter(
    all_combinations_binarized_without_total, 
    criteria_1 == TRUE
  )
  
  # Count frequency of binarized profiles in real-world data
  binarized_data_counted <- renamed_data %>%
    group_by_all() %>%
    summarise(freq = n(), .groups = 'drop')
  
  # Merge theoretically possible combinations with their frequency in real-world data
  merged_binarized_possible_real_without_total <- left_join(
    all_combinations_binarized_without_total_selectedPTSD, 
    binarized_data_counted, 
    by = symptom_cols_binarized_without_total
  )
  merged_binarized_possible_real_without_total$freq <- coalesce(
    merged_binarized_possible_real_without_total$freq, 
    0
  )
  
  return(merged_binarized_possible_real_without_total)
}
```
##### 5.3.2. Function for describing the resulting data frame with one or more deleted symptoms and for comparing this data set with the data set that takes all symptoms into account. 
```{r Function for describing the resulting data frame with one or more deleted symptoms and for comparing this data set with the data set that takes all symptoms into account., echo=FALSE, message=FALSE}
  ### Create function for describing the resulting data frame with one or more deleted symptoms and for comparing this data set with the data set that takes all symptoms into account. 
  # Input data frames: resulting data frames of 1. the function “create_possible_combinedwith_real_binarized_without_total” and 2. "create_possible_combinedwith_real_binarized_without_total_symp_del". The first data set comprises 21 columns (20 symptoms and one column “freq”). The second minus the deleted symptoms. The symptoms are each binarized. The rows are all theoretically possible combinations that fulfill the diagnostic criteria. The last column “freq” contains the number of presentations of this symptom combination in the real data.
description_and_comparison_symp_del_merged_binarized_possible_real_without_total <- function(data_all_symptoms, data_deleted_symptoms) {
  # Get description of both datasets using the original function
  desc_all <- description_merged_binarized_possible_real_without_total(data_all_symptoms)
  desc_deleted <- description_merged_binarized_possible_real_without_total(data_deleted_symptoms)
  
  # Helper function to safely convert percentage strings to numeric
  safe_percent_convert <- function(x) {
    if(is.character(x)) {
      as.numeric(gsub("[^0-9.-]", "", x))
    } else {
      as.numeric(x)
    }
  }
  
  ## Create comparison matrices
  # Matrix 1 comparison - Basic statistics
  comparison_matrix1 <- matrix(
    c(desc_all$Matrix1[,1], desc_deleted$Matrix1[,1], 
      mapply(function(all_val, del_val) {
        if(grepl("%", all_val)) {
          sprintf("%.2f%%", safe_percent_convert(del_val) - safe_percent_convert(all_val))
        } else {
          as.numeric(del_val) - as.numeric(all_val)
        }
      }, desc_all$Matrix1[,1], desc_deleted$Matrix1[,1])),
    ncol = 3
  )
  rownames(comparison_matrix1) <- rownames(desc_all$Matrix1)
  colnames(comparison_matrix1) <- c("All Symptoms", "Deleted Symptoms", "Difference")
  
  # Matrix 2 comparison - Statistic on the combinations
  comparison_matrix2 <- matrix(
    c(desc_all$Matrix2[,1], desc_deleted$Matrix2[,1],
      mapply(function(all_val, del_val) {
        if(grepl("%", all_val)) {
          sprintf("%.2f%%", safe_percent_convert(del_val) - safe_percent_convert(all_val))
        } else {
          as.numeric(del_val) - as.numeric(all_val)
        }
      }, desc_all$Matrix2[,1], desc_deleted$Matrix2[,1])),
    ncol = 3
  )
  rownames(comparison_matrix2) <- rownames(desc_all$Matrix2)
  colnames(comparison_matrix2) <- c("All Symptoms", "Deleted Symptoms", "Difference")
  
  # Matrix 3 comparison
  all_symptoms <- rownames(desc_all$Matrix3)
  
  comparison_matrix3 <- matrix(
    nrow = length(all_symptoms),
    ncol = 12
  )
  rownames(comparison_matrix3) <- all_symptoms
  colnames(comparison_matrix3) <- c(
    "N combinations actually occurring that include this symptom (All)", 
    "N combinations actually occurring that include this symptom(Deleted)", 
    "N combinations actually occurring that include this symptom Difference",
    "% combinations actually occurring that include this symptom (All)", 
    "% combinations actually occurring that include this symptom (Deleted)", 
    "% combinations actually occurring that include this symptom Difference",
    "N patients in the data set showing this symptom (All)", 
    "N patients in the data set showing this symptom (Deleted)",
    "N patients in the data set showing this symptom Difference",
    "% patients in the data set showing this symptom (All)",
    "% patients in the data set showing this symptom (Deleted)",
    "% patients in the data set showing this symptom Difference"
  )
  
  for(symptom in all_symptoms) {
    row_all <- desc_all$Matrix3[symptom,]
    
      # Check if symptom exists in deleted dataset
    if(symptom %in% rownames(desc_deleted$Matrix3)) {
      row_deleted <- desc_deleted$Matrix3[symptom,]

    # Convert values safely
    values <- c(
      safe_percent_convert(row_all[1]), 
      safe_percent_convert(row_deleted[1]),
      safe_percent_convert(row_deleted[1]) - safe_percent_convert(row_all[1]),
      safe_percent_convert(row_all[2]),
      safe_percent_convert(row_deleted[2]),
      safe_percent_convert(row_deleted[2]) - safe_percent_convert(row_all[2]),
      safe_percent_convert(row_all[3]), 
      safe_percent_convert(row_deleted[3]),
      safe_percent_convert(row_deleted[3]) - safe_percent_convert(row_all[3]),
      safe_percent_convert(row_all[4]),
      safe_percent_convert(row_deleted[4]),
      safe_percent_convert(row_deleted[4]) - safe_percent_convert(row_all[4])
    )
    } else {
      # For deleted symptoms, show original values and "DELETED" for deleted dataset, "NA" for difference
      values <- c(
        safe_percent_convert(row_all[1]), 
        "DELETED",
        "NA",
        safe_percent_convert(row_all[2]),
        "DELETED",
        "NA",
        safe_percent_convert(row_all[3]), 
        "DELETED",
        "NA",
        safe_percent_convert(row_all[4]),
        "DELETED",
        "NA"
      )
    }
    
    comparison_matrix3[symptom,] <- values
  }
  
  # Format percentages in matrices (only for percentage columns and for non-deleted symptoms)
  percent_cols <- c(4,5,6,10,11,12)
  for(col in percent_cols) {
    numeric_rows <- !comparison_matrix3[,col] %in% c("DELETED", "NA")
    comparison_matrix3[numeric_rows,col] <- sprintf("%.2f%%", as.numeric(comparison_matrix3[numeric_rows,col]))
  }
  
  ## Return all results in a list
  list(
    "Comparison_Matrix1" = comparison_matrix1,
    "Comparison_Matrix2" = comparison_matrix2,
    "Comparison_Matrix3" = comparison_matrix3
  )
}

```
##### 5.3.3. Application of the generated functions  
###### 5.3.3.1. Deletion of symptom 1
```{r Deletion of symptom 1, echo=FALSE, message=FALSE}
  ## Applying function “create_possible_combinedwith_real_binarized_without_total_symp_del“
possible_combinedwith_real_binarized_without_total_symp_del_1 <- create_possible_combined_with_real_binarized_without_total_symp_del(data2, 1)

  ## Applying function “description_and_comparison_symp_del_merged_binarized_possible_real_without_total“
description_and_comparison_symp_del_merged_binarized_possible_real_without_total(possible_combinedwith_real_binarized_without_total, possible_combinedwith_real_binarized_without_total_symp_del_1)
```
###### 5.3.3.2. Deletion of each symptom and analysis of the impact of the deleted symptom on the diagnosis
```{r Deletion of each symptom and analysis of the impact of the deleted symptom on the diagnosis, echo=FALSE, message=FALSE}
  ## Function to analyze impact of removing each symptom
  # Input-data: real-world dataset fulfilling diagnostic criteria, 20 columns with binarized symptoms, each row corresponds to an individual
analyze_deletion_each_symptom <- function(data) {
  # Store results
  results <- data.frame(
    symptom = 1:20,
    n_diagnoses = numeric(20),
    diff_diagnoses = numeric(20),
    n_theoretical_combinations = numeric(20),
    diff_theoretical_combinations = numeric(20),
    n_actual_combinations = numeric(20),
    diff_actual_combinations = numeric(20),
    perc_possible_combinations = numeric(20),
    diff_perc_possible_combinations = numeric(20)
  )
  
  # Get baseline metrics (with all symptoms)
  baseline <- create_possible_combinedwith_real_binarized_without_total(data)
  baseline_desc <- description_merged_binarized_possible_real_without_total(baseline)
  
  # Store baseline values
  baseline_diagnoses <- as.numeric(baseline_desc$Matrix1[1,1])
  baseline_theoretical <- as.numeric(baseline_desc$Matrix1[2,1])
  baseline_actual <- as.numeric(baseline_desc$Matrix1[3,1])
  baseline_percentage <- as.numeric(gsub("[^0-9.]", "", baseline_desc$Matrix1[4,1]))
  
  # Analyze each symptom
  for(i in 1:20) {
    # Calculate metrics without current symptom
    current <- create_possible_combined_with_real_binarized_without_total_symp_del(data2, i)
    current_desc <- description_merged_binarized_possible_real_without_total(current)
    
    # Store results
    results$n_diagnoses[i] <- as.numeric(current_desc$Matrix1[1,1])
    results$diff_diagnoses[i] <- results$n_diagnoses[i] - baseline_diagnoses
    
    results$n_theoretical_combinations[i] <- as.numeric(current_desc$Matrix1[2,1])
    results$diff_theoretical_combinations[i] <- results$n_theoretical_combinations[i] - baseline_theoretical
    
    results$n_actual_combinations[i] <- as.numeric(current_desc$Matrix1[3,1])
    results$diff_actual_combinations[i] <- results$n_actual_combinations[i] - baseline_actual
    
    results$perc_possible_combinations[i] <- as.numeric(gsub("[^0-9.]", "", current_desc$Matrix1[4,1]))
    results$diff_perc_possible_combinations[i] <- results$perc_possible_combinations[i] - baseline_percentage
  }
  
  # Add baseline row
  baseline_row <- data.frame(
    symptom = "no deletion",
    n_diagnoses = baseline_diagnoses,
    diff_diagnoses = 0,
    n_theoretical_combinations = baseline_theoretical,
    diff_theoretical_combinations = 0,
    n_actual_combinations = baseline_actual,
    diff_actual_combinations = 0,
    perc_possible_combinations = baseline_percentage,
    diff_perc_possible_combinations = 0
  )
  
  results <- rbind(baseline_row, results)
  
  # Order results by impact on diagnoses
  results <- results[order(abs(results$diff_diagnoses)), ]
  
  return(results)
}

  ## Applying function “analyze_deletion_each_symptom“
analysis_del_each_symptom <- analyze_deletion_each_symptom(data2)
```
## 6. Theoretically possible combinations of nonbinarized symptoms fulfilling diagnosis and their frequency in real-world data
```{r Theoretically possible combinations of nonbinarized symptoms fulfilling diagnosis and their frequency in real-world data, echo=FALSE, message=FALSE}
# ### DOES NOT WORK, BECAUSE MEMORY LIMIT IS REACHED, I am searching for alternative options (the use of chunks to handle the memory limit does not work either)
# ### Function for Dataframe with all possible combinations of nonbinarized symptoms fulfilling PTSD and their frequencies in real-world data
# create_possible_combinedwith_real_nonbinarized <- function(data) {
#   # Generate new dataframe (all_combinations_nonbinarized) with all possible combinations of symptom values
#   symptom_cols_nonbinarized <- paste0("symptom_", 1:20)
#   symptom_values_nonbinarized <- 0:4
#   all_combinations_nonbinarized <- expand.grid(lapply(1:20, function(i) symptom_values_nonbinarized))
#   colnames(all_combinations_nonbinarized) <- symptom_cols_nonbinarized
# 
#   # Calculate the total score for each row
#   all_combinations_nonbinarized$total <- rowSums(all_combinations_nonbinarized)
# 
#   # Define criteria for filtering
#   criteria_1 <- all_combinations_nonbinarized$total >= 33
#   criteria_2 <- (
#     rowSums(all_combinations_nonbinarized[, 1:5] >= 2) >= 1 &
#       rowSums(all_combinations_nonbinarized[, 6:7] >= 2) >= 1 &
#       rowSums(all_combinations_nonbinarized[, 8:14] >= 2) >= 2 &
#       rowSums(all_combinations_nonbinarized[, 15:20] >= 2) >= 2
#   )
# 
#   # Filter combinations based on criteria
#   all_combinations_nonbinarized_selectedPTSD <- all_combinations_nonbinarized[criteria_1 | criteria_2, ]
# 
#   # Count frequency of nonbinarized profiles in real-world data
#     # First ensure only using symptom columns
#     simulated_ptsd_total_selected <- data[, symptom_cols_nonbinarized]
# 
#     simulated_ptsd_total_selected_counted <- simulated_ptsd_total_selected %>%
#     group_by_all() %>%
#     summarise(freq = n(), .groups = 'drop')
# 
#   # Merge with the frequency data
#   merged_possible_real_nonbinarized <- left_join(all_combinations_nonbinarized_selectedPTSD, simulated_ptsd_total_selected_counted, by = symptom_cols_nonbinarized)
#   merged_possible_real_nonbinarized$freq <- coalesce(merged_possible_real_nonbinarized$freq, 0)
# 
#   return(merged_possible_real_nonbinarized)
# }
# 
#   ## Applying function to our data
#   possible_combinedwith_real_nonbinarized <- create_possible_combinedwith_real_nonbinarized(simulated_ptsd_total_selected)

```
## 7. Symptomanalysis, conditional probabilities, pairwise consideration, symptomgroup analysis
```{r Symptomanalysis, conditional probabilities, pairwise consideration, symptomgroup analysis, echo=FALSE, message=FALSE}
  ### Create function for Analysis of individual symptom frequencies, conditional probabilities, pairwise considerations, chi-square tests, symptom groups analysis
  # Input data frames: 20 columns with non-binarized symptoms, all patients included before application of diagnostic criteria
analyze_symptoms_condprob_pairs <- function(data) {
  ## Preparation of inputdata
  # Input validation
  if (!is.data.frame(data) || ncol(data) < 20) {
    stop("Input must be a data frame with at least 20 symptom columns")
  }
  
   # Function to binarize symptom scores with cutoff of 1 in orginial dataframe
  binarize_symptoms <- function(data, cutoff = 1) {
    result <- as.data.frame(lapply(data, function(x) {
      as.numeric(x > cutoff)
    }))
    return(result)
  }
  # Apply to our data
  binarized_all <- binarize_symptoms(data)
  
  # Function to check if diagnostic criteria are met for binarized data (without consideration of totalscore)
  check_diagnosis <- function(data) {
    criteria1 <- rowSums(data[, 1:5] > 0) >= 1     # At least 1 from 1-5
    criteria2 <- rowSums(data[, 6:7] > 0) >= 1     # At least 1 from 6-7
    criteria3 <- rowSums(data[, 8:14] > 0) >= 2    # At least 2 from 8-14
    criteria4 <- rowSums(data[, 15:20] > 0) >= 2   # At least 2 from 15-20
    
    criteria1 & criteria2 & criteria3 & criteria4
  }
  # Add diagnosis column
  binarized_all$PTSD_Diagnosis <- check_diagnosis(binarized_all)
  
  ## Create list to store results
  results <- list()
  
  ## 1. Individual symptom frequencies
  results$symptom_frequencies <- list(
  # 1.1 Overall frequencies
    freq_overall = colMeans(binarized_all[, 1:20]),
  # 1.2 Frequencies in diagnosed patients
    freq_diagnosed = colMeans(binarized_all[binarized_all$PTSD_Diagnosis == TRUE, 1:20]),
  # 1.3 Frequencies in non-diagnosed patients
    freq_nondiagnosed = colMeans(binarized_all[binarized_all$PTSD_Diagnosis == FALSE, 1:20])
  )
  
  ## 2. Conditional probabilities
  # 2.1 P(Diagnosis | Symptom present)
  results$conditional_probs <- lapply(1:20, function(i) {
      prob_diag_if_present = sum(binarized_all$PTSD_Diagnosis == TRUE & binarized_all[,i] == 1) / max(sum(binarized_all[,i] == 1), 1)
  # 2.2 P(Diagnosis | Symptom absent)
      prob_diag_if_absent <- sum(binarized_all$PTSD_Diagnosis == FALSE & binarized_all[,i] == 0) /  max(sum(binarized_all[,i] == 0), 1)
  # 2.3 Relative risk
      relative_risk <- if(prob_diag_if_absent > 0) {
        prob_diag_if_present / prob_diag_if_absent
      } else {
        NA
      }
      list(
        prob_diag_if_present = prob_diag_if_present,
        prob_diag_if_absent = prob_diag_if_absent,
        relative_risk = relative_risk
      )
  })
  
  ## 3. Pairwise considerations
  results$correlations <- list(
  # 3.1 Overall correlation matrix
    overall_cor = cor(binarized_all[, 1:20]),
  # 3.2 Correlation matrix for diagnosed patients
    diagnosed_cor = cor(binarized_all[binarized_all$PTSD_Diagnosis == TRUE, 1:20])
  )
  # 3.3.-3.6. 
  n <- 20
  # 3.3. Observed joint probabilities overall and for diagnosed patients
  joint_prob_overall <- matrix(0, n, n)
  joint_prob_diagnosed <- matrix(0, n, n)
  
  for(i in 1:n) {
    for(j in 1:n) {
      joint_prob_overall[i,j] <- mean(binarized_all[,i] & binarized_all[,j])
      joint_prob_diagnosed[i,j] <- mean(binarized_all[binarized_all$PTSD_Diagnosis == TRUE,i] & binarized_all[binarized_all$PTSD_Diagnosis == TRUE,j])
    }
  }
 
  # 3.4. Expected under independence overall and for diagnosed patients
  expected_prob_overall <- outer(results$symptom_frequencies$freq_overall, results$symptom_frequencies$freq_overall)
  expected_prob_diagnosed <- outer(results$symptom_frequencies$freq_diagnosed, results$symptom_frequencies$freq_diagnosed)
 
  # 3.5. Ratio observed and expected overall
  ratio_overall <- ifelse(expected_prob_overall > 0, joint_prob_overall/expected_prob_overall, NA)
  # 3.6. Comparison observed and expected for diagnosed patients
  ratio_diagnosed <- ifelse(expected_prob_diagnosed > 0, joint_prob_diagnosed/expected_prob_diagnosed, NA)
  # 3.7. Difference observed vs expected overall
  diff_prob_overall <- joint_prob_overall - expected_prob_overall
  # 3.8. Difference observed vs expected for diagnosed patients
  diff_prob_diagnosed <- joint_prob_diagnosed -  expected_prob_diagnosed
  
  # Store 3.3 -3.8. in a list
  results$symptom_relationships <- list(
    observed = list(
      overall = joint_prob_overall,
      diagnosed = joint_prob_diagnosed
    ),
    expected = list(
      overall = expected_prob_overall,
      diagnosed = expected_prob_diagnosed
    ),
    ratios = list(
      overall = ratio_overall,
      diagnosed = ratio_diagnosed
    ),
   differences = list(
      overall = diff_prob_overall,
      diagnosed = diff_prob_diagnosed
    )
  )
 
  ## 4. Symptom groups analysis
  # 4.1 Probability of diagnosis when group criteria are met
  group_criteria <- list(
    group1 = rowSums(binarized_all[,1:5]) >= 1,
    group2 = rowSums(binarized_all[,6:7]) >= 1,
    group3 = rowSums(binarized_all[,8:14]) >= 2,
    group4 = rowSums(binarized_all[,15:20]) >= 2
  )
  
  results$group_prob_diagnosis <- lapply(group_criteria, function(criteria) {
    sum(binarized_all$PTSD_Diagnosis == TRUE & criteria) / sum(criteria)
  })
  return(results)
}

  ### Applying function to our data
symptomanalysis_condprob_pairs <- analyze_symptoms_condprob_pairs(simulated_ptsd_data)
```
## 8. Calculation of the probability of different symptom combinations with binarized symptoms that meet the diagnostic criteria using different models
```{r Calculation of the probability of different symptom combinations with binarized symptoms that meet the diagnostic criteria using different models, echo=FALSE, message=FALSE}
#   ### PROBABLY NOT POSSIBLE, AS THERE IS TOO MUCH DATA; NOT EVEN WHEN PERFORMING ONE METHOD INDIVIDUALLY EXCEPT FOR THE INDEPENDENCE METHOD
#  # Required packages
# library(copula)
# library(mvtnorm)
# library(glmnet)
# library(IsingSampler)
# 
# ### Creation of a function to calculate the probability of different combinations of symptoms with binarized symptoms meeting the diagnostic criteria using different models (Independence Method, Observed Frequencies, Gaussian Copula, Multivariate Normal, Log-linear Model, Ising Model)
# # Input: 20 columns with non-binarized symptoms, all patients included before application of diagnostic criteria
# calculate_sympcombprob <- function(raw_data) {
#   # Helper function to binarize symptoms
#   binarize_symptoms <- function(data, cutoff = 1) {
#     result <- as.data.frame(lapply(data, function(x) {
#       as.numeric(x > cutoff)
#     }))
#     return(result)
#   }
#   
#   ## Step 1: Binarize the input data
#   binarized_all <- binarize_symptoms(raw_data)
#   
#   ## Step 2: Apply criteria to create valid_binarized_all
#   criteria_met_data <- (
#     rowSums(binarized_all[, 1:5]) >= 1 &      # At least 1 symptom from 1-5
#       rowSums(binarized_all[, 6:7]) >= 1 &      # At least 1 symptom from 6-7
#       rowSums(binarized_all[, 8:14]) >= 2 &     # At least 2 symptoms from 8-14
#       rowSums(binarized_all[, 15:20]) >= 2      # At least 2 symptoms from 15-20
#   )
#   
#   # Filter binarized data based on criteria
#   valid_binarized_all <- binarized_all[criteria_met_data, ]
#   
#   ## Step 3: Calculate covariance matrix from binarized data
#   covariance_matrix <- cov(valid_binarized_all)
#   
#   ## Step 4: Calculate symptom frequencies
#   symptom_frequencies <- colMeans(valid_binarized_all)
#   
#   ## Step 5: Generate all possible symptom combinations
#   symptom_cols <- paste0("symptom_", 1:20)
#   symptom_values <- 0:1
#   all_combinations <- expand.grid(lapply(1:20, function(i) symptom_values))
#   colnames(all_combinations) <- symptom_cols
#   
#   ## Step 6: Apply criteria
#   criteria_met <- (
#     rowSums(all_combinations[, 1:5]) >= 1 &      # At least 1 symptom from 1-5
#       rowSums(all_combinations[, 6:7]) >= 1 &      # At least 1 symptom from 6-7
#       rowSums(all_combinations[, 8:14]) >= 2 &     # At least 2 symptoms from 8-14
#       rowSums(all_combinations[, 15:20]) >= 2      # At least 2 symptoms from 15-20
#   )
#   
#   # Filter combinations based on criteria
#   valid_combinations <- all_combinations[criteria_met, ]
#   
#   ## Step 7: Calculate probabilities for valid combinations with different methods
#   # 7.1. Method 1: Independence Method
#   independence_prob <- function(combination) { #combination = single row
#     prob <- 1
#     for(i in 1:length(combination)) { #for each position in this combination
#       if(combination[i] == 1) {
#         prob <- prob * symptom_frequencies[i]
#       } else {
#         prob <- prob * (1 - symptom_frequencies[i])
#       }
#     }
#     return(prob)
#   }
#   valid_combinations$prob_independence <- apply(valid_combinations, 1, independence_prob)
# 
# 
#   # 7.2. Method 2: Observed Frequencies
#   observed_combinations <- apply(valid_binarized_all, 1, paste, collapse = "") #each row into a single string
#   get_observed_freq <- function(combination) {
#     pattern <- paste(combination, collapse = "") #convert combination to a string pattern
#     freq <- sum(observed_combinations == pattern) / length(observed_combinations) #how many times this exact pattern appears, divide by total number of observations
#     return(max(freq, 1e-10)) # Add small constant to avoid zero probabilities
#   }
#   valid_combinations$prob_observed <- apply(valid_combinations, 1, get_observed_freq)
# 
#   # 7.3. Method 3: Gaussian Copula
#   gaussian_copula_prob <- function(combination, covariance_matrix) {
#     # Create normal copula object
#     normal_copula <- normalCopula(param = P2p(covariance_matrix),
#                                   dim = ncol(covariance_matrix),
#                                   dispstr = "un")
# 
#     # Convert binary values to uniform margins
#     u <- ifelse(as.numeric(combination) == 1, 0.999, 0.001)  # Avoid exact 0s and 1s
# 
#     # Calculate probability using copula
#     prob <- pCopula(u, normal_copula)
#     return(max(prob, 1e-10))
#   }
# 
#   valid_combinations$prob_copula <- apply(valid_combinations, 1,
#                                           function(x) gaussian_copula_prob(x, covariance_matrix))
# 
#   # 7.4. Method 4: Multivariate Normal
#   # Convert covariance matrix to correlation matrix
#   correlation_matrix <- cov2cor(covariance_matrix)
# 
#   # Initialize probability vector
#   n_combinations <- nrow(valid_combinations)
#   mvn_probabilities <- numeric(n_combinations)
# 
#   # Calculate MVN probability for each combination
#   for(i in 1:n_combinations) {
#     current_combination <- as.numeric(valid_combinations[i,])
# 
#     # Use actual frequencies as means
#     means <- symptom_frequencies
# 
#     # Set bounds based on presence/absence of symptoms
#     lower_bounds <- ifelse(current_combination == 1, symptom_frequencies, -Inf)
#     upper_bounds <- ifelse(current_combination == 1, Inf, symptom_frequencies)
# 
#     # Calculate probability
#     prob <- try({
#       mvtnorm::pmvnorm(
#         lower = lower_bounds,
#         upper = upper_bounds,
#         mean = means,
#         sigma = correlation_matrix
#       )
#     }, silent = TRUE)
# 
#     if(!inherits(prob, "try-error")) {
#       mvn_probabilities[i] <- prob[1]
#     } else {
#       mvn_probabilities[i] <- NA
#     }
#   }
# 
#   # Add MVN probabilities to results
#   valid_combinations$prob_mvn <- mvn_probabilities
# 
#   # 7.5. Method 5: Log-linear Model # ERROR BECAUSE NO VARIATION
#   X <- as.matrix(valid_binarized_all) #Convert binarized data to a matrix
#   X[is.na(X)] <- 0 #Replace NAs with 0
#   y <- rep(1, nrow(X)) #Create vector of 1's with same length as number of patients for constant response for density estimation
#   loglinear_model <- cv.glmnet(X, y, family = "poisson")
# 
#   # Define probability function
#   loglinear_prob <- function(combination) {
#     # Ensure combination is properly formatted
#     combination_matrix <- matrix(as.numeric(combination), nrow = 1)
#     prob <- exp(predict(loglinear_model, newx = combination_matrix, s = "lambda.min"))
#     return(max(prob[1], 1e-10))
#   }
#   valid_combinations$prob_loglinear <- apply(valid_combinations, 1, loglinear_prob)
# 
#   # 7.6. Method 6: Ising Model 
#   # Fit Ising model
#   ising_data <- as.matrix(valid_binarized_all)
#   # Convert to numeric and ensure binary
#   ising_data <- matrix(as.numeric(ising_data), nrow = nrow(ising_data))
#   ising_data[is.na(ising_data)] <- 0
#   
#   ising_model <- EstimateIsing(ising_data, method = "pl") # Pseudo-likelihood estimation
#   
#   ising_prob <- function(combination) {
#     # Ensure combination is numeric binary
#     combination <- as.numeric(combination)
#     # Calculate energy of the configuration
#     energy <- CalculateEnergy(combination, ising_model$graph)
#     # Convert energy to probability
#     prob <- exp(-energy) / ising_model$Z # Z is the partition function
#     return(max(prob, 1e-10))
#   }
#   valid_combinations$prob_ising <- apply(valid_combinations, 1, ising_prob)
#   
#   # Normalize probabilities to sum to 1 for each method
#   normalize_probs <- function(probs) {
#     return(probs / sum(probs, na.rm = TRUE))
#   }
#   
#   prob_columns <- grep("^prob_", colnames(valid_combinations), value = TRUE)
#   valid_combinations[prob_columns] <- lapply(valid_combinations[prob_columns], normalize_probs)
#   
#   return(valid_combinations)
# }
# 
# ### Applying function to our data
# sympcomb_prob_binarized <- calculate_sympcombprob(simulated_ptsd_data)
```
## x. Session Info

```{r Session Info, echo=FALSE, message=FALSE}
sessionInfo()
```
