---
title: "Assess Simulated PTSD Symptoms"
author:
- "<h5 style=\"font-style:italic\"> Laura Weidmann"
- "<h5 style=\"font-style:italic\"> Tobias R. Spiller"
date: "<h5 style=\"font-style:roman\"> `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
  pdf_document:
    toc: true
    toc_depth: '5'
subtitle: Version 0.0.6
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)
```

```{r Load Libraries, message=FALSE, warning=TRUE, include=FALSE}
# Data handling
library(tidyverse)
library(data.table)
library(DT)
library(gtsummary)
library(psych)
```

## 1. Import and prepare data

```{r Import data, message=FALSE, warning=TRUE, include=FALSE}
# Import data
simulated_ptsd_data <- read_csv("Data/simulated_skewed_ptsd_data.csv")

# Renaming the columns in the input data (20 columns) to match the expected format
simulated_ptsd_data_renamed <- simulated_ptsd_data %>%
  rename_with(~ paste0("symptom_", 1:20))
```

## 2. Sample descriptive

### 2.1. Total score & Diagnosis

```{r Total score & Diagnosis, echo=FALSE, message=FALSE}
# Create PCL total score
simulated_ptsd_total <- simulated_ptsd_data_renamed %>% 
  mutate(total = rowSums(select(., symptom_1:symptom_20)))

# Function to determine original PTSD diagnosis with non-binarized symptom scores
create_ptsd_diagnosis_nonbinarized <- function(data) {
   criteria <- list(
    A = rowSums(data[, paste0("symptom_", 1:5)] >= 2) >= 1,
    B = rowSums(data[, paste0("symptom_", 6:7)] >= 2) >= 1,
    C = rowSums(data[, paste0("symptom_", 8:14)] >= 2) >= 2,
    D = rowSums(data[, paste0("symptom_", 15:20)] >= 2) >= 2
 )

    data$PTSD_Diagnosis <- Reduce(`&`, criteria)
    return(data)
  }

# Apply the function to our dataframe
simulated_ptsd_total <- create_ptsd_diagnosis_nonbinarized (simulated_ptsd_total)

# Summarize
simulated_ptsd_total %>%
  summarise(
    mean_total = mean(total),
    sd_total = sd(total),
    n_diagnosed = sum(PTSD_Diagnosis)
  )
```

### 2.2. Cronbach's alpha

Selected sample

```{r PCL cronbach, echo=FALSE, message=FALSE}
cronbach <- psych::alpha(subset(simulated_ptsd_total, select = (-total)))
cronbach$total
```

## 3. Minimal, polythetic set of PTSD symptoms: the three best 6 symptom combinations for which the diagnosis is fulfilled with at least 4/6

### 3.1. Helper Functions

```{r Helper Functions, echo=FALSE, message=FALSE}
# Function to binarize symptom scores (Input: dataframe with raw symptom scores, Output: dataframe with binarized scores (0/1))
binarize_data <- function(data) {
  # Binarize values (0,1 -> 0; 2,3,4 -> 1)
  data[data <= 1] <- 0
  data[data >= 2] <- 1
  return(data)
}

# Function to determine original PTSD diagnosis with binarized symptom scores (Input: dataframe with raw symptom scores of 20 symptoms; output: dataframe with binarized symptom scores and column indicating whether original diagnostic criteria are met)
create_ptsd_diagnosis_binarized <- function(data) {
  check_ptsd_criteria <- function(symptoms) {
    criterion_1 <- any(symptoms[1:5] == 1)
    criterion_2 <- any(symptoms[6:7] == 1)
    criterion_3 <- sum(symptoms[8:14] == 1) >= 2
    criterion_4 <- sum(symptoms[15:20] == 1) >= 2
    return(criterion_1 & criterion_2 & criterion_3 & criterion_4)
  }
  
  # Binarize data
  binarized_data <- binarize_data(data)
  # Check PTSD criteria for each row
  ptsd_results <- apply(binarized_data, 1, check_ptsd_criteria)

  return(data.frame(PTSD_all = ptsd_results))
}

# Function to summarize PTSD diagnosis changes (Input: dataframe with columns showing whether the diagnosis is fulfilled under certain diagnostic criteria)
summarize_ptsd_changes <- function(data) {
  # Initialize results dataframe
  summary_stats <- data.frame(
    column = names(data),
    diagnosed = colSums(data),
    non_diagnosed = colSums(!data),
    stringsAsFactors = FALSE
  )
  
  # Calculate changes compared to PTSD_all
  baseline <- data$PTSD_all
  # For each column
  for(col in names(data)) {
    current <- data[[col]]

    newly_diagnosed <- sum(!baseline & current)
    newly_nondiagnosed <- sum(baseline & !current)
    true_positive <- sum(baseline & current)
    true_negative <- sum(!baseline & !current)

    summary_stats$newly_diagnosed[summary_stats$column == col] <- newly_diagnosed
    summary_stats$newly_nondiagnosed[summary_stats$column == col] <- newly_nondiagnosed
    summary_stats$true_positive[summary_stats$column == col] <- true_positive
    summary_stats$true_negative[summary_stats$column == col] <- true_negative
    summary_stats$true_cases[summary_stats$column == col] <- true_positive + true_negative
    summary_stats$false_cases[summary_stats$column == col] <- newly_diagnosed + newly_nondiagnosed
  }
  
  # Calculate percentages
  total_cases <- nrow(data)
  summary_stats$diagnosed_percent <- round(summary_stats$diagnosed / total_cases * 100, 2)
  summary_stats$non_diagnosed_percent <- round(summary_stats$non_diagnosed / total_cases * 100, 2)
  
  return(summary_stats)
}

# Function to create readable summary (Input: resulting data frame of the function "summarize_ptsd_changes", columns: which symptom deletions should be displayed)
create_readable_summary <- function(summary_stats, columns = NULL) {
  if(is.null(columns)) {
    summary_subset <- summary_stats
  } else {
    summary_subset <- summary_stats[summary_stats$column %in% columns,]
  }
  
  readable_summary <- data.frame(
    Scenario = summary_subset$column,
    `Total Diagnosed` = paste0(summary_subset$diagnosed, 
                               " (", summary_subset$diagnosed_percent, "%)"),
    `Total Non-Diagnosed` = paste0(summary_subset$non_diagnosed,
                                   " (", summary_subset$non_diagnosed_percent, "%)"),
    `True Positive` = summary_subset$true_positive,
    `True Negative` = summary_subset$true_negative,
    `Newly Diagnosed` = summary_subset$newly_diagnosed,
    `Newly Non-Diagnosed` = summary_subset$newly_nondiagnosed,
    `True Cases` = summary_subset$true_cases,
    `False Cases` = summary_subset$false_cases
  )
  return(readable_summary)
}

# Shared function to find best combinations
find_best_combinations <- function(combinations, binarized_data, baseline_results, score_by, get_diagnosis_fn) {
  top_combinations <- list(
    first = list(combination = NULL, score = -Inf, diagnoses = NULL),
    second = list(combination = NULL, score = -Inf, diagnoses = NULL),
    third = list(combination = NULL, score = -Inf, diagnoses = NULL)
  )
  
  for(combination in combinations) {
    current_diagnoses <- get_diagnosis_fn(binarized_data, combination)
    
    newly_diagnosed <- sum(!baseline_results & current_diagnoses)
    newly_nondiagnosed <- sum(baseline_results & !current_diagnoses)
    
    score <- if(score_by == "false_cases") {
      -(newly_diagnosed + newly_nondiagnosed)
    } else {
      -newly_nondiagnosed
    }
    
    if(score > top_combinations$first$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- top_combinations$first
      top_combinations$first <- list(
        combination = combination,
        score = score,
        diagnoses = current_diagnoses
      )
    } else if(score > top_combinations$second$score) {
      top_combinations$third <- top_combinations$second
      top_combinations$second <- list(
        combination = combination,
        score = score,
        diagnoses = current_diagnoses
      )
    } else if(score > top_combinations$third$score) {
      top_combinations$third <- list(
        combination = combination,
        score = score,
        diagnoses = current_diagnoses
      )
    }
  }
  
  return(top_combinations)
}
```

### 3.2. Three best six-symptom combinations, of which at least 4 must be present to fulfill the diagnosis, regardless of the cluster

```{r Three best six-symptom combinations, of which at least 4 must be present to fulfill the diagnosis, regardless of the cluster, echo=FALSE, message=FALSE}
### Function to determine the three best six-symptom-combinations, where at least 4 need to be present for diagnosis, regardless of which cluster they belong to.
  # Input: data set with all patients and 20 columns with non-binarized symptoms
  # Depending on the score_by select the best six-symptom-combinations by minimizing the newly_nondiagnosed or by minimizing the false_cases
  # Output: List containing best combinations and diagnosis comparison
analyze_best_six_symptoms_four_required <- function(data, score_by = "false_cases") {
  # Validate scoring method
  valid_scoring <- c("false_cases", "newly_nondiagnosed")
  if (!score_by %in% valid_scoring) {
    stop("score_by must be one of: ", paste(valid_scoring, collapse = ", "))
  }

  # Get baseline results and binarize data
  baseline_results <- create_ptsd_diagnosis_binarized(data)$PTSD_all
  binarized_data <- binarize_data(data)
  
  # Helper function for diagnosis
  get_diagnosis <- function(data, symptoms) {
    subset_data <- data[, paste0("symptom_", symptoms)]
    return(rowSums(subset_data) >= 4)  # At least 4 symptoms must be present
  }

  # Generate all possible combinations of 6 symptoms and find best ones
  all_symptoms <- 1:20
  combinations <- combn(all_symptoms, 6, simplify = FALSE)

  top_combinations <- find_best_combinations(combinations, binarized_data, baseline_results, score_by, get_diagnosis)
 
  # Create comparison dataframe
  comparison_df <- data.frame(
    PTSD_all = baseline_results,
    sapply(1:3, function(i) top_combinations[[i]]$diagnoses)
  )
  names(comparison_df)[2:4] <- sapply(1:3, function(i) {
    paste0("symptom_", paste(top_combinations[[i]]$combination, collapse = "_"))
  })
  
  return(list(
    best_symptoms = lapply(1:3, function(i) top_combinations[[i]]$combination),
    diagnosis_comparison = comparison_df
  ))
}
```
Summary for three best six-symptom-combinations, of which at least 4 must be present to fulfill the diagnosis, regardless of the cluster. Best means fewest newly_nondiagnosed
```{r Summary for three best six-symptom-combinations, of which at least 4 must be present to fulfill the diagnosis, regardless of the cluster. Best means fewest newly_nondiagnosed, echo=FALSE, message=FALSE}
### Applying to our data
PTSD_Diagnosis_best_six_four_required_min_false_neg <- analyze_best_six_symptoms_four_required(simulated_ptsd_data_renamed, score_by = "newly_nondiagnosed")

# Readable summary
datatable(create_readable_summary(summarize_ptsd_changes(PTSD_Diagnosis_best_six_four_required_min_false_neg$diagnosis_comparison)),
          options = list(scrollX = TRUE))
```

### 3.3. Three best six-symptom combinations, of which at least 4 must be present to fulfill the diagnosis, one from each cluster

```{r Three best six-symptom combinations, of which at least 4 must be present to fulfill the diagnosis, one from each cluster, echo=FALSE, message=FALSE}
### Function to determine the three best six-symptom-combinations, where at least 4 need to be present for diagnosis, one from each cluster by testing all valid (at least one symptom from each cluster) combinations
  # Input: data set with all patients and 20 columns with non-binarized symptoms
  # Depending on the score_by select the best six-symptom-combinations by minimizing the newly_nondiagnosed or by minimizing the false_cases
  # Output: List containing best combinations and diagnosis comparison
analyze_best_six_symptoms_four_required_clusters <- function(data, score_by = "false_cases") {
  # Validate scoring method
  valid_scoring <- c("false_cases", "newly_nondiagnosed")
  if (!score_by %in% valid_scoring) {
    stop("score_by must be one of: ", paste(valid_scoring, collapse = ", "))
  }
  
  # Get baseline results and binarize data
  baseline_results <- create_ptsd_diagnosis_binarized(data)$PTSD_all
  binarized_data <- as.matrix(binarize_data(data))
  
  # Define clusters
  clusters <- list(
    cluster1 = 1:5,    
    cluster2 = 6:7,    
    cluster3 = 8:14,   
    cluster4 = 15:20   
  )
  
  # Create lookup array for faster cluster membership checking
  cluster_lookup <- integer(20)
  for(i in seq_along(clusters)) {
    cluster_lookup[clusters[[i]]] <- i
  }

  # Fast cluster representation check using lookup
  check_cluster_representation <- function(symptoms) {
    length(unique(cluster_lookup[symptoms])) == 4
  }
  
  # Helper function for diagnosis
  get_diagnosis <- function(data, symptoms) {
    subset_data <- data[, symptoms, drop = FALSE]
    symptom_counts <- rowSums(subset_data)
    sufficient_rows <- which(symptom_counts >= 4)
    
    result <- logical(nrow(data))
    
    if(length(sufficient_rows) > 0) {
      for(i in sufficient_rows) {
        present_symptoms <- symptoms[subset_data[i,] == 1]
        if(length(present_symptoms) >= 4) {
          result[i] <- check_cluster_representation(present_symptoms)
        }
      }
    }
    
    return(result)
  }
  
  # Generate valid combinations efficiently
  valid_combinations <- vector("list", 1000)  # Pre-allocate
  combination_count <- 0
  
  for(s1 in clusters$cluster1) {
    for(s2 in clusters$cluster2) {
      for(s3 in clusters$cluster3) {
        for(s4 in clusters$cluster4) {
          base <- c(s1, s2, s3, s4)
          remaining <- setdiff(1:20, base)
          pairs <- combn(remaining, 2, simplify = FALSE)
          
          for(pair in pairs) {
            combination_count <- combination_count + 1
            if(combination_count > length(valid_combinations)) {
              length(valid_combinations) <- length(valid_combinations) * 2
            }
            valid_combinations[[combination_count]] <- sort(c(base, pair))
          }
        }
      }
    }
  }
  
  valid_combinations <- valid_combinations[1:combination_count]
  valid_combinations <- unique(valid_combinations)

  # Find best combinations
  top_combinations <- find_best_combinations(
    valid_combinations, 
    binarized_data, 
    baseline_results, 
    score_by, 
    get_diagnosis
  )
  
  # Create comparison dataframe
  comparison_df <- data.frame(
    PTSD_all = baseline_results,
   sapply(1:3, function(i) top_combinations[[i]]$diagnoses)
  )
  names(comparison_df)[2:4] <- sapply(1:3, function(i) {
    paste0("symptom_", paste(top_combinations[[i]]$combination, collapse = "_"))
  })
  
  return(list(
    best_symptoms = lapply(1:3, function(i) top_combinations[[i]]$combination),
    diagnosis_comparison = comparison_df
  ))
}
```
Summary for three best six-symptom-combinations, of which at least 4 must be present to fulfill the diagnosis, one from each cluster. Best means fewest newly_nondiagnosed
```{r Summary for three best six-symptom-combinations, of which at least 4 must be present to fulfill the diagnosis, one from each cluster. Best means fewest newly_nondiagnosed, echo=FALSE, message=FALSE}
### Applying to our data
PTSD_Diagnosis_best_six_four_required_clusters_min_false_neg <- analyze_best_six_symptoms_four_required_clusters(simulated_ptsd_data_renamed, score_by = "newly_nondiagnosed")

# Readable summary
datatable(create_readable_summary(summarize_ptsd_changes(PTSD_Diagnosis_best_six_four_required_clusters_min_false_neg$diagnosis_comparison)),
          options = list(scrollX = TRUE))
```

## 4. Session Info

```{r Session Info, echo=FALSE, message=FALSE}
sessionInfo()
```
